### 2013-09-23 annotated_bibliography.tex
changed the author name

% This is a sample document using the University of Minnesota, Morris, Computer Science
% Senior Seminar modification of the ACM sig-alternate style to generate a simple annotated
% bibliography. The idea is that this document is fairly short, consisting of a brief description
% of your sources and how you intend to use them (or not). Most of the ``content'' of the
% generated document comes from the bibliography file, including the notes field which will
% provide the annotations.

% See https://github.com/UMM-CSci/Senior_seminar_templates for more info and to make
% suggestions and corrections.

\documentclass{sig-alternate}

\begin{document}

% --- Author Metadata here ---
%%% REMEMBER TO CHANGE THE SEMESTER AND YEAR
\conferenceinfo{UMM CSci Senior Seminar Conference, December 2013}{Morris, MN}

\title{Something about NP-complete problems}

\numberofauthors{1}

\author{
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
\alignauthor
Emma Ireland\\
	\affaddr{Division of Science and Mathematics}\\
	\affaddr{University of Minnesota, Morris}\\
	\affaddr{Morris, Minnesota, USA 56267}\\
	\email{irela065@morris.umn.edu}
}

\maketitle

\begin{abstract}
This paper discusses new results in NP-complete problems and the use of distributed
networks to solve certain partial cases of NP-complete problems.
\end{abstract}

% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Delphi theory}

\keywords{ACM proceedings, \LaTeX, text tagging}

\section{Introduction}
I will focus on using the approach Blah for solving partial cases of
NP-complete problems on distributed networks.

I plan to use the following sources:
\begin{itemize}
\item I expect~\cite{OM:2008} to be one of my main sources, and I'm still looking for another two ``core'' papers to build on.
\item I may use~\cite{Brun:2008} for comparison. 
\item I'll use~\cite{Aaronson:2005, wiki:np-complete} and possibly selected chapters of~\cite{Garey:1979} as background. 
\end{itemize}

As mentioned above I need two other ``core'' papers, and I'm still looking for good examples that I can use to explain the

I was initially considering algorithms on compete graphs as a possible topic, and looked over~\cite{winkler1984isometric, dobkin1987delaunay, folkman1970graphs} before I settled on my current topic.

% The following two commands are all you need to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
% annotated_bibliography.bib is the name of the BibTex file containing 
% all the bibliography entries for this example. Note that you *don't* include the .bib ending
% in the \bibliography command.
\bibliography{annotated_bibliography}

% You must have a ".bib" file and remember to run:
%     pdflatex bibtex pdflatex pdflatex
% in order to see all the citation references correctly.

\end{document}

### 2013-09-26 annotated_bibliography.tex
Worked on my annotated bibliography.

@@
 
-\title{Something about NP-complete problems}
+\title{Intrusion Detection with Genetic Algorithms and Fuzzy Logic}
 
@@
 \begin{abstract}
-This paper discusses new results in NP-complete problems and the use of distributed
-networks to solve certain partial cases of NP-complete problems.
+This paper discusses the use of artificial intelligence in intrusion detection systems.
 \end{abstract}
@@
 % A category with the (minimum) three required fields
-\category{H.4}{Information Systems Applications}{Miscellaneous}
+%\category{H.4}{Information Systems Applications}{Miscellaneous}
 %A category including the fourth, optional field follows...
-\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
+%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
 
-\terms{Delphi theory}
+\terms{Intrusion detection}
 
-\keywords{ACM proceedings, \LaTeX, text tagging}
+\keywords{Genetic algorithms, Fuzzy logic}
 
 \section{Introduction}
-I will focus on using the approach Blah for solving partial cases of
-NP-complete problems on distributed networks.
+I will be looking at how genetic algorithms and fuzzy logic are used to train intrusion detection systems about possible threats.
 
@@
 \begin{itemize}
-\item I expect~\cite{OM:2008} to be one of my main sources, and I'm still looking for another two ``core'' papers to build on.
-\item I may use~\cite{Brun:2008} for comparison. 
-\item I'll use~\cite{Aaronson:2005, wiki:np-complete} and possibly selected chapters of~\cite{Garey:1979} as background. 
+\item I expect~\cite{DBLP:journals/corr/abs-1204-1336} and~\cite{6496342} to be two of my main sources.
+\item I may use~\cite{bc1_ecindm} for my third source or for background. 
+\item I'll use~\cite{Liao201316, DBLP:journals/corr/abs-1204-6416, bapuji2012soft} as background. 
 \end{itemize}
 
-As mentioned above I need two other ``core'' papers, and I'm still looking for good examples that I can use to explain the
-
-I was initially considering algorithms on compete graphs as a possible topic, and looked over~\cite{winkler1984isometric, dobkin1987delaunay, folkman1970graphs} before I settled on my current topic.
+I won't be using~\cite{Sangkatsanee20112227}, because it doesn't talk about genetic algorithms or fuzzy logic.
 

### 2013-10-06 paper.tex
Added some sections to my paper.

--- 2013-09-26 annotated_bibliography.tex
@@
-% This is a sample document using the University of Minnesota, Morris, Computer Science
-% Senior Seminar modification of the ACM sig-alternate style to generate a simple annotated
-% bibliography. The idea is that this document is fairly short, consisting of a brief description
-% of your sources and how you intend to use them (or not). Most of the ``content'' of the
-% generated document comes from the bibliography file, including the notes field which will
-% provide the annotations.
+\documentclass{sig-alternate}
+\usepackage{color}
 
-% See https://github.com/UMM-CSci/Senior_seminar_templates for more info and to make
-% suggestions and corrections.
-
-\documentclass{sig-alternate}
+%%%% User-defined macros
+\newcommand{\lam}{\lambda}
+\newcommand{\mycomment}[1]{\textcolor{red}{#1}}
+%%%%% Uncomment the following line and comment out the previous one
+%%%%% to remove all comments
+%%%%% NOTE: comments still occupy a line even if invisible;
+%%%%% Don't write them as a separate paragraph
+%\newcommand{\mycomment}[1]{}
 
@@
 \maketitle
+\begin{abstract}
 
-\begin{abstract}
-This paper discusses the use of artificial intelligence in intrusion detection systems.
 \end{abstract}
@@
 
-\terms{Intrusion detection}
+%\terms{Delphi theory}
 
-\keywords{Genetic algorithms, Fuzzy logic}
+\keywords{Intrusion detection, genetic algorithm, fuzzy logic}
 
 \section{Introduction}
-I will be looking at how genetic algorithms and fuzzy logic are used to train intrusion detection systems about possible threats.
 
-I plan to use the following sources:
-\begin{itemize}
-\item I expect~\cite{DBLP:journals/corr/abs-1204-1336} and~\cite{6496342} to be two of my main sources.
-\item I may use~\cite{bc1_ecindm} for my third source or for background. 
-\item I'll use~\cite{Liao201316, DBLP:journals/corr/abs-1204-6416, bapuji2012soft} as background. 
-\end{itemize}
+\section{Background}
 
-I won't be using~\cite{Sangkatsanee20112227}, because it doesn't talk about genetic algorithms or fuzzy logic.
+\subsection{Types of Intrusion Detection Systems}
 
-% The following two commands are all you need to
+\subsection{Types of Networking Attacks}
+
+\subsection{Detection Methodologies}
+
+\subsection{Genetic Algorithm}
+
+\subsection{Fuzzy Logic}
+
+\subsection{KDD99 Data Set}
+
+\section{Fuzzy Genetic Algorithm Implementation}
+
+\subsection{RLD09 Data Set}
+
+\subsection{Algorithm Overview}
+
+\subsection{Experimental Design and Results}
+
+\section{Genetic Algorithm Implementation}
+
+\subsection{Algorithm Overview}
+
+\subsection{Experimental Design and Results}
+
+\section{Conclusions}
+
+\section{Acknowledgments}
+
+% The following two commands are all you need in the
+% initial runs of your .tex file to
 % produce the bibliography for the citations in your paper.
 \bibliographystyle{abbrv}
-% annotated_bibliography.bib is the name of the BibTex file containing 
-% all the bibliography entries for this example. Note that you *don't* include the .bib ending
-% in the \bibliography command.
+% sample_paper.bib is the name of the BibTex file containing the
+% bibliography entries. Note that you *don't* include the .bib ending here.
 \bibliography{annotated_bibliography}  
-
-% You must have a ".bib" file and remember to run:
-%     pdflatex bibtex pdflatex pdflatex
-% in order to see all the citation references correctly.
+% You must have a proper ".bib" file
+%  and remember to run:
+% latex bibtex latex latex
+% to resolve all references
 

### 2013-10-09 paper.tex
Wrote a section on the fuzzy genetic algorithm implementation.

@@
 \section{Fuzzy Genetic Algorithm Implementation}
+The focus of the research is on detecting new or unknown types of attacks in a network.
+The intrusion detection system is able to identify normal network activity as well as classify different attack types using a fuzzy genetic algorithm.
+This kind of algorithm is able to learn new attacks, and has a high detection rate.
+The system used is evaluated in terms of detection speed, detection rate, and false alarm rate.
 
 \subsection{RLD09 Data Set}
+The KDD99 dataset is 14 years old, and newer attack types are not included in it because of its age.
+Because of this, the authors of~\cite{6496342, 6559603} created their own data set, RLD09.
+To create the data set, the authors captured network data from the Computer Engineering Department at King Mongkut's University of Technology Thonburi, in Bangkok, Thailand.
+The data has around ten million preprocessed data packets.
+It has 17 different types of attacks, as well as normal network activity.
+The attacks can be divided into denial of service attacks and probe attacks.
+A packet sniffer was used to get information about TCP, UDP, and ICMP headers from protocol packets.
+Then this information is processed into 12 features by counting connections between a source IP and destination IP.
+Some example features include the number of: packets, source ports, and destination ports.
+The following are examples of data records, where each one has 12 feature values.
+
+\begin{itemize}
+\item 21,21,15,0,21,0,0,0,0,0,0,0, attack
+\item 102,2,2,0,0,1,102,0,0,0,0,0, normal
+\end{itemize}
 
 \subsection{Algorithm Overview}
+The algorithm that is used in~\cite{6496342, 6559603} first randomly finds a rule.
+Then the rule is improved in the training phase.
+After that, the rules are used to classify the data into classes in the testing phase.
+The pseudo code below describes the fuzzy genetic algorithm that is used.
+
+\mycomment{Ask Nic how to put in algorithm.}
+
+The fitness function to be maximized is:
+\begin{equation*}
+fitness function = \frac{\alpha}{A} - \frac{\beta}{B}
+\end{equation*}
+In the implementation, a population size of 10 was used for each generation.
+An individual in the population represents a possible detection rule.
+The two best individuals from a present generation are preserved for the next generation.
+The other individuals in the new generation come from single-point crossover.
+
+\subsection{Fuzzy Algorithm}
+In order to measure the probability of an attack, a trapezoidal shape was used in the algorithm.
+The trapezoidal shape has four parameters: a, b, c, and d.
+The following algorithm calculates the probability of being attacked:
+
+\mycomment{Ask Nic how to put in algorithm.}
+
+\mycomment{Put in trapezoidal shape figure.}
+
+\subsection{Features and Rules}
+The KDD99 dataset is composed of 41 features.
+The authors of~\cite{6496342, 6559603} used the following eight features in their system: duration, src\_bytes, num\_failed\_logins, root\_shell, num\_access\_files, srv\_count, serror\_rate, same\_srv\_rate.
+
+The value of each feature is normalized to be a number between 0.0 and 7.0, and then is encoded into blocks of binary strings.
+See Figure~\ref{fig:fuzEncodingForFeature} for an example of a block.
+A rule has 12 blocks of features, and at the end of the string is the type of attack.
+
+\begin{figure}
+\centering
+\caption{Fuzzy encoding for a feature}
+\begin{tabular}{|c|c|c|c|} \hline
+010 & 011 & 100 & 101\\
+a & b & c & d\\
+\hline\end{tabular}
+\label{fig:fuzEncodingForFeature}
+\end{figure}
+
+One record is passed into a rule.
+Each feature in a record is matched to one block of the rule.
+The parameters of each block measure the probability of an attack using the trapezoidal fuzzy rule shape.
+The probabilities of each block are then combined, and the average of the probabilities is compared with a threshold to determine if the record is an attack class or normal class.
+Then the predicted result is compared with the actual result.
 

### 2013-10-12 paper.tex
Wrote sections on types of IDS, types of networking attacks, and detection methodologies.

@@
 \subsection{Types of Intrusion Detection Systems}
+There are two types of intrusion detection systems: \emph{host based} and \emph{network based}.
+Host based intrusion detection systems look at and analyze information that is on a single host or multiple host systems, such as the contents of file systems and system calls to determine if there is a possible threat to the system.
+Network based intrusion detection looks at information that is from the network.
+Network based is useful for analyzing the traffic of multiple systems all at once.
+This type of intrusion detection system analyzes packets that travel through the network.~\cite{DBLP:journals/corr/abs-1204-1336}
 
 \subsection{Types of Networking Attacks}
+There are four different types of networking attacks: \emph{denial of service}, \emph{remote to user attacks}, \emph{user to root attacks}, and \emph{probing}.
+Each attack that happens on a network can be placed into one of these categories.
+
+Denial of service attacks happen when an attacker makes a machine inaccessible to a user by making it too busy to serve legitimate network requests.
+Remote to user attacks happen when an attacker sends packets to a machine over the network in order to get access to things a local user would have on the machine.
+User to root attacks happen when an attacker starts out with access on the machine and then tries to gain root access to the system.
+Probing happens when an attacker examines a machine in order to collect information about weaknesses or vulnerabilities that in the future could be used in such a way that they compromise the system.~\cite{DBLP:journals/corr/abs-1204-1336}
 
 \subsection{Detection Methodologies}
+There are two different ways of detecting attacks: \emph{signature-based detection} and \emph{anomaly-based detection}.
+
+A signature is a pattern that corresponds to a known attack.
+Signature-based detection compares well-known patterns of attacks that are already in the intrusion detection system against captured events in order to identify a possible attack.
+It is a simple and effective way to detect known attacks.
+A disadvantage of it is that it is an ineffective way of detecting unknown attacks.
+Signature-based detection is also called \emph{knowledge-based detection} or \emph{misuse detection}.~\cite{Liao201316}
+
+An anomaly is something that deviates from what is normal.
+Anomaly-based detection looks for patterns of activity that are rare and uncommon.
+It is an effective way to detect new vulnerabilities.
+Some disadvantages of this type of system is that they are expensive and they can detect an intrusive behavior as being normal because of insufficient data.
+Anomaly-based detection is also called \emph{behavior-based detection}.~\cite{DBLP:journals/corr/abs-1204-1336}
 
@@
 \subsection{RLD09 Data Set}
-The KDD99 dataset is 14 years old, and newer attack types are not included in it because of its age.
+The KDD99 data set is 14 years old, and newer attack types are not included in it because of its age.
 Because of this, the authors of~\cite{6496342, 6559603} created their own data set, RLD09.
@@
 
-\subsection{Fuzzy Algorithm}
 In order to measure the probability of an attack, a trapezoidal shape was used in the algorithm.
@@
 \subsection{Features and Rules}
-The KDD99 dataset is composed of 41 features.
+The KDD99 data set is composed of 41 features.
 The authors of~\cite{6496342, 6559603} used the following eight features in their system: duration, src\_bytes, num\_failed\_logins, root\_shell, num\_access\_files, srv\_count, serror\_rate, same\_srv\_rate.

### 2013-10-12 paper.tex
Added stuff about false positives and false negatives, also started writing the experimental design and results section.

@@
 
+\subsection{False Positives and False Negatives}
+\emph{False positives} and {False negatives} are two values that are used to evaluate the accuracy of an intrusion detection system.
+A false positive happens when an intrusion detection system incorrectly identifies normal activity as being an attack.
+A false negative happens when an intrusion detection system fails to identify harmful activity.~\cite{Liao201316}
+
 \subsection{Genetic Algorithm}
@@
 \subsection{Experimental Design and Results}
+The experiments that the authors of~\cite{6496342, 6559603} performed used a total of 16,000 records of normal activity and 10,500 records of attack activity.
+Of the attack activity, 4,000 of them were denial of service attacks and 6,500 were probe attacks.
+
+In the first experiment, the fuzzy genetic algorithm was used to create denial of service and probe detection rules and then the rules were verified with known attack types.
+10,000 records were used for the training set and all 26,500 records were used for the testing set.
+The two steps in the training process were to find a denial of service rule, and find a probe rule from the training set.
+Both of these rules were then used in the testing process to identify attacks from the testing data set.
+The following algorithm was used in the testing process:
+
+\mycomment{Ask Nic how to put in algorithm.}
+
+The detection rate of denial of service attacks in training was 91.64\% and the detection rate of probe attacks in training was 94.79\%.
+The detection rate of the testing data set increased to 97.92\%.
+The false positive rate was 1.13\% and the false negative rate was 4.10\%.
+Detection of an attack happened 2-3 seconds after the packet data arrived at the intrusion detection system.
+Results from this experiment are shown in Table~\ref{tab:fuzGenExp1}.
+
+\mycomment{Fix table}
+
+\begin{table}
+\caption{Results from Experiment 1}
+\begin{tabular}{|ccccccc|} \hline
+ & Attack & Normal & Total Records & FP & FN & DR\\
+DoS Training & 1499 & 8501 & 10000 & 1.46 & 47.50 & 91.64\\
+Probe Training & 2496 & 7504 & 10000 & 1.83 & 15.38 & 94.79\\
+Testing & 10500 & 1600 & 26500 & 1.13 & 4.10 & 97.92\\
+\hline\end{tabular}
+\label{tab:fuzGenExp1}
+\end{table}
 

### 2013-10-12 paper.tex
Added more to the results section.

@@
 % A category with the (minimum) three required fields
-%\category{H.4}{Information Systems Applications}{Miscellaneous}
-%A category including the fourth, optional field follows...
-%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
+\category{}{Security and Privacy}{Intrusion Detection Systems}
 
-%\terms{Delphi theory}
+\terms{}
 
@@
 
+\mycomment{Explain the 8 features}
+
 The value of each feature is normalized to be a number between 0.0 and 7.0, and then is encoded into blocks of binary strings.
@@
 \subsection{Experimental Design and Results}
+\subsubsection{Experiments using Only RLD09}
 The experiments that the authors of~\cite{6496342, 6559603} performed used a total of 16,000 records of normal activity and 10,500 records of attack activity.
@@
 
+In the second experiment, seven tests were run.
+For each test case there were 13 attack types plus normal activity that were in the training data set.
+Three attack types were used for the unknown testing data set.
+For example, the first test case used the training data set that does not have Advance Port Scan, Ack Scan, and Xmas Tree, (which are all probe attacks).
+These three attacks were then used for the testing data set.
+Table~\ref{tab:fuzGenExp2} shows some of the results from the fuzzy genetic algorithm and a decision tree algorithm.
+The fuzzy genetic algorithm has an accuracy of at least 95\% in identifying unknown attacks.
+When compared with the decision tree algorithm, the fuzzy genetic algorithm has a higher detection rate in all cases except 3 and 5.
+It can be seen that in cases 2 and 4 the decision tree has low detection rates, while the fuzzy genetic algorithm has much higher detection rates.
+
+\begin{table}
+\caption{Unknown Attack Experiment}
+\begin{tabular}{|lllcrl|} \hline
+Test & Unknown & Decision & & Fuzzy & \\
+Case & Attacks & Tree DR & & Genetic DR & \\ \hline
+
+1 & Advance Port Scan & 99.7 & Avg = & 100 & Avg =\\
+  & Ack Scan          & 100  & 98.33 & 100 & 100\\
+  & Xmas Tree         & 95.3 &       & 100 &\\ \hline
+
+2 & UDP Flood & \textbf{26.7} & Avg = & \textbf{100} & Avg =\\
+  & Host Scan & 98.15         & 46.65 & 99.20        & 99.80\\
+  & UDP Scan  & \textbf{17.6} &       & \textbf{100} &\\ \hline
+
+3 & Jping    & 100 & Avg =          & 99.20 & Avg =\\
+  & Syn Scan & 99.1& \textbf{99.70} & 96.60 & 98.75\\
+  & Fin Scan & 100 &                & 100 &\\ \hline
+
+4 & UDP Flood & \textbf{26.6} & Avg = & \textbf{100} & Avg =\\
+  & RCP Scan  & 91.56         & 70.35 & 92.60        & 98.15\\
+  & Fin Scan  & 99.8          &       & 100          &\\ \hline
+
+5 & Http Flood & 85.7& Avg =          & 100 & Avg =\\
+  & RCP Scan  & 92.3 & \textbf{99.94} & 91.80 & 97.50\\
+  & Fin Scan  & 99.8 &                & 98.20 &\\
+
+\hline\end{tabular}
+\label{tab:fuzGenExp2}
+\end{table}
+
+\subsubsection{Experiments using Both RLD09 and KDD99}
+The authors of~\cite{6496342, 6559603} also ran experiments that compared the RLD09 data set with the KDD99 data set.
+They used the KDD99 10\% version file for both the training dataset and testing dataset.
+The number of records of normal and attack activity for the two data sets are shown in Table~\ref{tab:RecordsForKDD99}.
+
+\begin{table}
+\caption{Number of Records for KDD99}
+\begin{tabular}{|lll|} \hline
+Attack & Training & Testing\\
+Type   & Data set & Data set\\ \hline
+Normal & 39,387 & 39,337\\
+DoS & 158,597 & 158,503\\
+Probe & 1,550 & 1,674\\
+ & &\\
+ Total & 199,534 & 199,514\\
+
+\hline\end{tabular}
+\label{tab:RecordsForKDD99}
+\end{table}
+
 \section{Genetic Algorithm Implementation}

### 2013-10-12 paper.tex
Wrote about experiments that used both the RLD09 and KDD99 data sets.

@@
 \subsection{False Positives and False Negatives}
-\emph{False positives} and {False negatives} are two values that are used to evaluate the accuracy of an intrusion detection system.
+\emph{False positives} and \emph{False negatives} are two values that are used to evaluate the accuracy of an intrusion detection system.
 A false positive happens when an intrusion detection system incorrectly identifies normal activity as being an attack.
@@
   & Fin Scan  & 99.8 &                & 98.20 &\\
-
 \hline\end{tabular}
@@
 \subsubsection{Experiments using Both RLD09 and KDD99}
-The authors of~\cite{6496342, 6559603} also ran experiments that compared the RLD09 data set with the KDD99 data set.
+The authors of~\cite{6496342, 6559603} also ran experiments that used and compared the RLD09 data set with the KDD99 data set.
 They used the KDD99 10\% version file for both the training dataset and testing dataset.
-The number of records of normal and attack activity for the two data sets are shown in Table~\ref{tab:RecordsForKDD99}.
+The number of records of normal and attack activity for the KDD99 data set is shown in Table~\ref{tab:RecordsForKDD99}, and the number of records of normal and attack activity for the RLD09 data is shown in Table~\ref{tab:RecordsForRLD09}.
 
@@
  Total & 199,534 & 199,514\\
-
 \hline\end{tabular}
 \label{tab:RecordsForKDD99}
+\end{table}
+
+\begin{table}
+\caption{Number of Records for RLD09}
+\begin{tabular}{|lll|} \hline
+Attack & Training & Testing\\
+Type   & Data set & Data set\\ \hline
+Normal & 8,000 & 16,000\\
+DoS & 2,400 & 4,000\\
+Probe & 3,900 & 6,500\\
+ & &\\
+ Total & 14,300 & 26,500\\
+\hline\end{tabular}
+\label{tab:RecordsForRLD09}
+\end{table}
+
+The authors of~\cite{6496342, 6559603} first trained and tested the fuzzy genetic algorithm with the KDD99 data set.
+There were 6 different types of denial of service attacks and 4 different types of probe attacks.
+The detection rate of the KDD99 data set was 98.72\%, and it had a false negative rate of 1.55\%.
+Then 26,500 records of the RLD09 data set were used as the training set.
+The detection rate was 97.97\%.
+The results of this experiment are shown in Table~\ref{tab:bothSetsResults}.
+
+\begin{table}
+\caption{KDD99 and RLD09 Results}
+\begin{tabular}{|cccccc|} \hline
+Data & Testing & Testing & False    & False    & Detection\\
+set  & Attack  & Normal  & Positive & Negative & Rate\\ \hline
+KDD99 & 160,117 & 39,337 & 0.13 & 1.55 & 98.72\\
+RLD09 & 10,500 & 16,000 & 1.14 & 3.39 & 97.97\\
+\hline\end{tabular}
+\label{tab:bothSetsResults}
+\end{table}
+
+The next experiment was the use of the KDD99 training set with the fuzzy genetifc algorithm to separate the data into two classes.
+Then each specific attack type was extracted and combined with normal activity.
+Ten tests were run, and Table~\ref{tab:kddAttacks} shows the accuracy of detecting some of the cases.
+The results showed that the detection rate of most of the cases were greater than 93\%.
+The results also showed that the behavior of the attacks were highly distinctive from normal activity.
+There were only two cases that had low detection rates, one of which is case 1 in Table~\ref{tab:kddAttacks}.
+
+\begin{table}
+\caption{Results for KDD99 with Certain Attacks}
+\begin{tabular}{|cccccc|} \hline
+Test & Attack & Type & FP & FN & DR\\
+Case & Name   &&&&\\ \hline
+1 & Back & DoS & 85.33 & 0.00 & 16.56\\
+2 & Land & DoS & 0.39 & 0.00 & 99.61\\
+3 & Smurf & DoS & 0.76 & 0.10 & 99.73\\
+4 & Portsweep & Probe & 6.40 & 0.00 & 93.66\\
+5 & Satan & Probe & 0.74 & 3.75 & 99.22\\
+\hline\end{tabular}
+\label{tab:kddAttacks}
+\end{table}
+
+The final experiment that was run used only the RLD09 data set with the fuzzy genetic algorithm.
+17 tests were run, and Table~\ref{tab:rldAttacks} shows the accuracy of detecting some of the cases.
+The results showed that the detection rate of a majority of the cases were greater than 97.5\%.
+Again, there were only two cases that had low detection rates, one of which is case 2 in Table~\ref{tab:rldAttacks}.
+
+\begin{table}
+\caption{Results for RLD99 with Certain Attacks}
+\begin{tabular}{|cccccc|} \hline
+Test & Attack & Type & FP & FN & DR\\
+Case & Name   &&&&\\ \hline
+1 & Smurf & DoS & 0.02 & 0 & 99.98\\
+2 & UDP Flood & DoS & 11.06 & 0 & 89.59\\
+3 & Ackscan & Probe & 0.03 & 0 & 99.97\\
+4 & Ipscan & Probe & 13.01 & 16.4 & 86.89\\
+5 & Synscan & Probe & 0.65 & 4.2 & 99.24\\
+\hline\end{tabular}
+\label{tab:rldAttacks}
 \end{table}

### 2013-10-13 paper.tex
Started writing the section on the genetic algorithm implementation.

@@
 
-\keywords{Intrusion detection, genetic algorithm, fuzzy logic}
+\keywords{Intrusion detection, genetic algorithm, fuzzy logic, KDD99, RLD09}
 
@@
 This kind of algorithm is able to learn new attacks, and has a high detection rate.
-The system used is evaluated in terms of detection speed, detection rate, and false alarm rate.
+The system used is evaluated in terms of detection speed, detection rate, and false alarm rate.~\cite{6496342, 6559603}
 
@@
 
-\mycomment{Ask Nic how to put in algorithm.}
+\mycomment{Figure out how to put in algorithm.}
+
+for each record
+
+   for each rule
+
+      for each attribute
+
+         prob = fuzzy();
+
+         totalprob = totalprob + prob;
+
+      If (totalprob > threshold)
+
+         class is attack;
+
+      else
+
+         class is normal;
+
+      compare the predicted result with actual result
+
+\mycomment{put in alpha and beta symbols}
+
+      find A, B, alpha, and beta,
+
+   calculate fitness
+
+   // create next generation
+
+   preserve\_best()
+
+   crossover()
+
+   mutation()
+
+// A is total number of attack records.
+B is total number of normal records.
+alpha is total number of attack records correctly identified as attack.
+beta is total number of normal records incorrectly classified as attack (false positive).
 
@@
 
-\mycomment{Ask Nic how to put in algorithm.}
+\mycomment{Figure out how to put in algorithm.}
+
+if (data value is between b and c)
+
+   prob =1.0
+
+else if (data value is between a and b)
+\begin{equation*}
+   prob = \frac{data - a}{b-a}
+\end{equation*}
+
+else if (data value is between c and d)
+\begin{equation*}
+   prob = \frac{d-data}{d-c}
+\end{equation*}
+
+else
+
+   prob = 0.0
 
@@
 
-\mycomment{Explain the 8 features}
+duration is the length of the connection in seconds.
+src\_bytes is the number of bytes sent from source to destination.
+num\_failed\_logins is the number of failed login attempts.
+root\_shell returns 1 if root shell is obtained and 0 otherwise.
+num\_access\_files is the number of operations on access control files.
+srv\_count is the number of connections to the same service as the current connection in the past two seconds.
+serror\_rate is the percentage of connections that have ``SYN" errors.
+\mycomment{What are SYN errors?}
+ same\_srv\_rate is the percentage of connections to the same service.~\cite{KDD99Features}
 
@@
 
-\mycomment{Ask Nic how to put in algorithm.}
+\mycomment{Figure out how to put in algorithm.}
+
+If (dos\_rule = yes or probe\_rule = yes)
+
+   This record is an attack;
+
+else
+
+   This record is normal;
 
@@
 
-\mycomment{Fix table}
+\mycomment{Fix tables}
 
@@
 \subsection{Algorithm Overview}
+The authors of~\cite{DBLP:journals/corr/abs-1204-1336} used a genetic algorithm to make their intrusion detection system.
+The system is divided into two phases: a precalculation phase and a detection phase.
+In the precalculation phase, a set of chromosomes are created using training data.
+Then this set of chromosomes is used in the detection phase for comparison.
+The following algorithm is used in the precalculation phase:
+
+\mycomment{Figure out how to put in algorithm.}
+
+Algorithm : Initialize chromosomes for comparison
+
+Input : Network audit data (for training)
+
+Output : A set of chromosomes
+
+Range = 0.125
+
+For each training data
+
+   If it has neighboring chromosome within Range
+
+      Merge it with the nearest chromosome
+
+   Else
+
+      Create new chromosome with it
+
+   End if
+
+End for
+
+In the detection phase, a population is created for test data and then the type of the data is predicted.
+The set of chromosomes that was created in the precalculation phase is used in the detection phase to find the fitness of each chromosome in the population.
+The following algorithm is used in the detection phase:
+
+\mycomment{Figure out how to put in algorithm.}
+
+Algorithm : Predict data/intrusion type (using GA)
+
+Input : Network audit data (for testing), Precalculated set of chromosomes
+
+Output : Type of data.
+
+Initialize the population
+
+CrossoverRate = 0.15, MutationRate = 0.35
+
+While number of generation is not reached
+
+   For each chromosome in the population
+
+      For each precalculated chromosome
+
+         Find fitness
+
+      End for
+
+      Assign optimal fitness as the fitness of that chromosome
+
+   End for
+
+   Remove some chromosomes with worse fitness
+
+   Apply crossover to the selected pair of chromosomes of the population
+
+   Apply mutation to each chromosome of the population
+
+End while
+
+The authors of~\cite{DBLP:journals/corr/abs-1204-1336} used the KDD99 data set.
+For the training data set, the KDD99 10\% version file was used, and for the testing data set, the KDD99 corrected version file was used.
 

### 2013-10-13 paper.tex
Wrote some things in the KDD99 data set section.

@@
 
-\subsection{False Positives and False Negatives}
-\emph{False positives} and \emph{False negatives} are two values that are used to evaluate the accuracy of an intrusion detection system.
-A false positive happens when an intrusion detection system incorrectly identifies normal activity as being an attack.
-A false negative happens when an intrusion detection system fails to identify harmful activity.~\cite{Liao201316}
-
 \subsection{Genetic Algorithm}
@@
 \subsection{KDD99 Data Set}
+Knowledge Discovery in Databases (KDD) is the process of extracting useful and interesting information from data in order to understand patterns in the data, with a limited amount of human intervention.
+Data is collected from several different sources and then is analyzed in order to produce patterns or models over the data.~\cite{bc1_ecindm} The KDD99 data set is a benchmark data set that was simulated in a military network environment in 1998, and then it was formed into the KDD99 data set in 1999~\cite{6559603}.
+In an experiment, the researchers divide the data set into two subsets, a \emph{training set} and a \emph{testing set}.
+Algorithms are trained on the training set to look for patterns.
+These patterns are then verified when the testing set is used.~\cite{bc1_ecindm} Four different measures are then used to determine the accuracy of an algorithm, which are described in Secion~\ref{sec:PosNeg}.
+
+\subsection{False Positives, False Negatives, True Positives, True Negatives}
+\label{sec:PosNeg}
+\emph{False positives} and \emph{False negatives} are two values that are used to evaluate the accuracy of an intrusion detection system.
+A false positive happens when an intrusion detection system incorrectly identifies normal activity as being an attack.
+A false negative happens when an intrusion detection system fails to identify harmful activity.~\cite{Liao201316}
+
+\mycomment{Write more here}
 

### 2013-10-15 paper.tex
Put two algorithms into paper, also moved some sections around.

@@
 \usepackage{color}
+\usepackage{algorithmic}
+\usepackage{algorithm}
 
@@
 
+\subsection{Fuzzy Algorithm}
+In order to measure the probability of an attack, a trapezoidal shape was used in the algorithm. The trapezoidal shape has four parameters: a, b, c, and d. 
+\mycomment{Put in trapezoidal shape figure.} Algorithm~\ref{alg:fuzAlg} calculates the probability of being attacked.
+
+\begin{algorithm}
+\caption{Fuzzy Algorithm}
+\label{alg:fuzAlg}
+\begin{algorithmic}
+\IF{data value is between b and c}
+  \STATE{prob = 1.0}
+\ELSIF{data value is between a and b}
+  \STATE{
+    prob = $(\textrm{data}-a)/(b-a)$
+  }
+\ELSIF{data value is between c and d}
+  \STATE{
+    prob = $(d-\textrm{data})/(d-c)$
+  }
+\ELSE \STATE{prob = 0.0}
+\ENDIF
+\end{algorithmic}
+\end{algorithm}
+
 \subsection{Algorithm Overview}
@@
 After that, the rules are used to classify the data into classes in the testing phase.
-The pseudo code below describes the fuzzy genetic algorithm that is used.
+Algorithm~\ref{alg:fuzGenAlg} describes the fuzzy genetic algorithm that is used.
 
-\mycomment{Figure out how to put in algorithm.}
+\begin{algorithm}
+\caption{Fuzzy Genetic Algorithm}
+\label{alg:fuzGenAlg}
+\begin{algorithmic}
+\FOR{each record}
 
-for each record
+  \FOR{each rule}
+    \FOR{each attribute}
+      \STATE{prob = fuzzy();}
+      \STATE{totalprob = totalprob + prob;}
+    \ENDFOR
 
-   for each rule
+    \IF{totalprob > threshold} 
+      \STATE{class is attack;}
+      \ELSE \STATE{class is normal;}
+    \ENDIF
+  \ENDFOR
 
-      for each attribute
-
-         prob = fuzzy();
-
-         totalprob = totalprob + prob;
-
-      If (totalprob > threshold)
-
-         class is attack;
-
-      else
-
-         class is normal;
-
-      compare the predicted result with actual result
-
-\mycomment{put in alpha and beta symbols}
-
-      find A, B, alpha, and beta,
-
-   calculate fitness
-
-   // create next generation
-
-   preserve\_best()
-
-   crossover()
-
-   mutation()
-
-// A is total number of attack records.
+  \STATE{compare the predicted result with actual result}
+  \STATE{find A, B, $\alpha$, and $\beta$}
+  \STATE{// A is total number of attack records.
 B is total number of normal records.
-alpha is total number of attack records correctly identified as attack.
-beta is total number of normal records incorrectly classified as attack (false positive).
+$\alpha$ is total number of attack records correctly identified as attack.
+$\beta$ is total number of normal records incorrectly classified as attack.}
+\ENDFOR
+\STATE{calculate fitness}
+\STATE{//create next generation}
+\STATE{preserve\_best()}
+\STATE{crossover()}
+\STATE{mutation()}
+\end{algorithmic}
+\end{algorithm}
 
@@
 The other individuals in the new generation come from single-point crossover.
-
-In order to measure the probability of an attack, a trapezoidal shape was used in the algorithm.
-The trapezoidal shape has four parameters: a, b, c, and d.
-The following algorithm calculates the probability of being attacked:
-
-\mycomment{Figure out how to put in algorithm.}
-
-if (data value is between b and c)
-
-   prob =1.0
-
-else if (data value is between a and b)
-\begin{equation*}
-   prob = \frac{data - a}{b-a}
-\end{equation*}
-
-else if (data value is between c and d)
-\begin{equation*}
-   prob = \frac{d-data}{d-c}
-\end{equation*}
-
-else
-
-   prob = 0.0
-
-\mycomment{Put in trapezoidal shape figure.}
 

### 2013-10-16 paper.tex
Moved a few sections around, and made sure the algorithms were formated nicely.

@@
 
+\subsection{Rules}
+Rules are represented by IF-THEN statements with the following format:
+
+IF (cond1) AND...AND (condm) THEN class = ci
+
+\mycomment{Write more here and figure out how to format the rule}
+
 \subsection{KDD99 Data Set}
 Knowledge Discovery in Databases (KDD) is the process of extracting useful and interesting information from data in order to understand patterns in the data, with a limited amount of human intervention.
-Data is collected from several different sources and then is analyzed in order to produce patterns or models over the data.~\cite{bc1_ecindm} The KDD99 data set is a benchmark data set that was simulated in a military network environment in 1998, and then it was formed into the KDD99 data set in 1999~\cite{6559603}.
+Data is collected from several different sources and then is analyzed in order to produce patterns or models over the data.~\cite{bc1_ecindm}
+
+The KDD99 data set is a benchmark data set that was simulated in a military network environment in 1999~\cite{6559603}.
+It is composed of 41 \emph{features}, which are things that help to distinguish normal connections from attacks.
+
 In an experiment, the researchers divide the data set into two subsets, a \emph{training set} and a \emph{testing set}.
@@
 
-\mycomment{Write more here}
+\mycomment{Write more here about True Pos, True Neg, Detection Rate}
 
 \section{Fuzzy Genetic Algorithm Implementation}
-The focus of the research is on detecting new or unknown types of attacks in a network.
-The intrusion detection system is able to identify normal network activity as well as classify different attack types using a fuzzy genetic algorithm.
-This kind of algorithm is able to learn new attacks, and has a high detection rate.
-The system used is evaluated in terms of detection speed, detection rate, and false alarm rate.~\cite{6496342, 6559603}
+\label{sec:fuzGenAlgImp}
+\mycomment{Figure out better title for Section~\ref{sec:fuzGenAlgImp}}
 
-\subsection{RLD09 Data Set}
-The KDD99 data set is 14 years old, and newer attack types are not included in it because of its age.
-Because of this, the authors of~\cite{6496342, 6559603} created their own data set, RLD09.
-To create the data set, the authors captured network data from the Computer Engineering Department at King Mongkut's University of Technology Thonburi, in Bangkok, Thailand.
-The data has around ten million preprocessed data packets.
-It has 17 different types of attacks, as well as normal network activity.
-The attacks can be divided into denial of service attacks and probe attacks.
-A packet sniffer was used to get information about TCP, UDP, and ICMP headers from protocol packets.
-Then this information is processed into 12 features by counting connections between a source IP and destination IP.
-Some example features include the number of: packets, source ports, and destination ports.
-The following are examples of data records, where each one has 12 feature values.
+\subsection{Setup}
+The authors of~\cite{6496342, 6559603} used 8 out of the 41 KDD99 features in their system: duration, src\_bytes, num\_failed\_logins, root\_shell, num\_access\_files, srv\_count, serror\_rate, same\_srv\_rate.
 
-\begin{itemize}
-\item 21,21,15,0,21,0,0,0,0,0,0,0, attack
-\item 102,2,2,0,0,1,102,0,0,0,0,0, normal
-\end{itemize}
+Duration is the length of the connection in seconds.
+src\_bytes is the number of bytes sent from source to destination.
+num\_failed\_logins is the number of failed login attempts.
+root\_shell returns 1 if root shell is obtained and 0 otherwise.
+num\_access\_files is the number of operations on access control files.
+srv\_count is the number of connections to the same service as the current connection in the past two seconds.
+serror\_rate is the percentage of connections that have ``SYN" errors.
+\mycomment{What are SYN errors?}
+ same\_srv\_rate is the percentage of connections to the same service.~\cite{KDD99Features}
 
 \subsection{Fuzzy Algorithm}
+The focus of~\cite{6496342, 6559603} is on detecting new or unknown types of attacks in a network.
+The intrusion detection system used is able to identify normal network activity as well as attacks using a fuzzy genetic algorithm.
+This kind of algorithm is able to learn new attacks, and has a high detection rate.
+The system used is evaluated in terms of detection speed, detection rate, and false alarm rate.
+
 In order to measure the probability of an attack, a trapezoidal shape was used in the algorithm. The trapezoidal shape has four parameters: a, b, c, and d. 
@@
 
+The four parameters a, b, c, and d are encoded into blocks of binary strings, where each block is a feature with parameters between 0.0 and 7.0.
+See Figure~\ref{fig:fuzEncodingForFeature} for an example of a block.
+A rule has 12 blocks of features, and at the end of the string is the type of attack.
+
+\begin{figure}
+\centering
+\caption{Fuzzy encoding for a feature}
+\begin{tabular}{|c|c|c|c|} \hline
+010 & 011 & 100 & 101\\
+a & b & c & d\\
+\hline\end{tabular}
+\label{fig:fuzEncodingForFeature}
+\end{figure}
+
+One record (either an attack or normal activity) is passed into a rule.
+Each feature in a record is matched to one block of the rule.
+The parameters of each block measure the probability of an attack using the trapezoidal fuzzy rule shape.
+The probabilities of each block are then combined by taking the average of the probabilities and comparing them with a threshold to determine if the record is an attack class or normal class.
+
 \subsection{Algorithm Overview}
-The algorithm that is used in~\cite{6496342, 6559603} first randomly finds a rule.
+The algorithm that is used in~\cite{6496342, 6559603} first randomly generates a rule.
 Then the rule is improved in the training phase.
@@
 \begin{equation*}
-fitness function = \frac{\alpha}{A} - \frac{\beta}{B}
+\textrm{fitness function} = \frac{\alpha}{A} - \frac{\beta}{B}
 \end{equation*}
@@
 
-\subsection{Features and Rules}
-The KDD99 data set is composed of 41 features.
-The authors of~\cite{6496342, 6559603} used the following eight features in their system: duration, src\_bytes, num\_failed\_logins, root\_shell, num\_access\_files, srv\_count, serror\_rate, same\_srv\_rate.
+\subsection{Experimental Design and Results}
 
-duration is the length of the connection in seconds.
-src\_bytes is the number of bytes sent from source to destination.
-num\_failed\_logins is the number of failed login attempts.
-root\_shell returns 1 if root shell is obtained and 0 otherwise.
-num\_access\_files is the number of operations on access control files.
-srv\_count is the number of connections to the same service as the current connection in the past two seconds.
-serror\_rate is the percentage of connections that have ``SYN" errors.
-\mycomment{What are SYN errors?}
- same\_srv\_rate is the percentage of connections to the same service.~\cite{KDD99Features}
+\subsubsection{RLD09 Data Set}
+The KDD99 data set is 14 years old, and newer attack types are not included in it because of its age.
+Because of this, the authors of~\cite{6496342, 6559603} created their own data set, RLD09, to use in their experiments.
+To create the data set, the authors captured network data from the Computer Engineering Department at King Mongkut's University of Technology Thonburi, in Bangkok, Thailand.
+The data has around ten million preprocessed data packets.
+It has 17 different types of attacks, as well as normal network activity.
+The attacks can be divided into denial of service attacks and probe attacks.
+A packet sniffer was used to get information about TCP, UDP, and ICMP headers from protocol packets.
+Then this information was processed into 12 features.
+The features are the number of TCP, UDP, and ICMP packets, the number of TCP and UDP source ports, the number of TCP and UDP destination ports, and the number of TCP fin flags, syn flags, push flags, ack flags, urgent flags.
 
-The value of each feature is normalized to be a number between 0.0 and 7.0, and then is encoded into blocks of binary strings.
-See Figure~\ref{fig:fuzEncodingForFeature} for an example of a block.
-A rule has 12 blocks of features, and at the end of the string is the type of attack.
-
-\begin{figure}
-\centering
-\caption{Fuzzy encoding for a feature}
-\begin{tabular}{|c|c|c|c|} \hline
-010 & 011 & 100 & 101\\
-a & b & c & d\\
-\hline\end{tabular}
-\label{fig:fuzEncodingForFeature}
-\end{figure}
-
-One record is passed into a rule.
-Each feature in a record is matched to one block of the rule.
-The parameters of each block measure the probability of an attack using the trapezoidal fuzzy rule shape.
-The probabilities of each block are then combined, and the average of the probabilities is compared with a threshold to determine if the record is an attack class or normal class.
-Then the predicted result is compared with the actual result.
-
-\subsection{Experimental Design and Results}
 \subsubsection{Experiments using Only RLD09}
@@
 Both of these rules were then used in the testing process to identify attacks from the testing data set.
-The following algorithm was used in the testing process:
+Algorithm~\ref{alg:testProc} was used in the testing process.
 
-\mycomment{Figure out how to put in algorithm.}
-
-If (dos\_rule = yes or probe\_rule = yes)
-
-   This record is an attack;
-
-else
-
-   This record is normal;
+\begin{algorithm}
+\caption{}
+\label{alg:testProc}
+\begin{algorithmic}
+\IF{dos\_rule = yes or probe\_rule = yes}
+  \STATE{This record is an attack;}
+\ELSE \STATE{This record is normal;}
+\ENDIF
+\end{algorithmic}
+\end{algorithm}
 
@@
 
-\begin{table}
+\begin{table*}
 \caption{Results from Experiment 1}
@@
 \label{tab:fuzGenExp1}
-\end{table}
+\end{table*}
 
@@
 \section{Genetic Algorithm Implementation}
+\label{sec:genAlgImp}
+\mycomment{Figure out better title for Section~\ref{sec:genAlgImp}}
 
@@
 Then this set of chromosomes is used in the detection phase for comparison.
-The following algorithm is used in the precalculation phase:
+Algorithm~\ref{alg:precalc} is used in the precalculation phase.
 
-\mycomment{Figure out how to put in algorithm.}
-
-Algorithm : Initialize chromosomes for comparison
-
-Input : Network audit data (for training)
-
-Output : A set of chromosomes
-
-Range = 0.125
-
-For each training data
-
-   If it has neighboring chromosome within Range
-
-      Merge it with the nearest chromosome
-
-   Else
-
-      Create new chromosome with it
-
-   End if
-
-End for
+\begin{algorithm}
+\caption{Major steps in precalculation}
+\label{alg:precalc}
+\begin{algorithmic}
+\STATE{range = 0.125}
+\FOR{each training data}
+  \IF{it has neighboring chromosome within range}
+    \STATE{Merge it with the nearest chromosome}
+  \ELSE \STATE{Create new chromosome with it}
+  \ENDIF
+\ENDFOR
+\end{algorithmic}
+\end{algorithm}
 
@@
 The set of chromosomes that was created in the precalculation phase is used in the detection phase to find the fitness of each chromosome in the population.
-The following algorithm is used in the detection phase:
+Algorithm~\ref{alg:detection} is used in the detection phase.
 
-\mycomment{Figure out how to put in algorithm.}
-
-Algorithm : Predict data/intrusion type (using GA)
-
-Input : Network audit data (for testing), Precalculated set of chromosomes
-
-Output : Type of data.
-
-Initialize the population
-
-CrossoverRate = 0.15, MutationRate = 0.35
-
-While number of generation is not reached
-
-   For each chromosome in the population
-
-      For each precalculated chromosome
-
-         Find fitness
-
-      End for
-
-      Assign optimal fitness as the fitness of that chromosome
-
-   End for
-
-   Remove some chromosomes with worse fitness
-
-   Apply crossover to the selected pair of chromosomes of the population
-
-   Apply mutation to each chromosome of the population
-
-End while
+\begin{algorithm}
+\caption{Major steps in detection}
+\label{alg:detection}
+\begin{algorithmic}
+\STATE{Initialize the population}
+\STATE{crossoverRate = 0.15, mutationRate = 0.35}
+\WHILE{number of generation is not reached}
+  \FOR{each chromosome in the population}
+    \FOR{each precalculated chromosome}
+      \STATE{Find fitness}
+    \ENDFOR
+    \STATE{Assign optimal fitness as the fitness of that chromosome}
+  \ENDFOR
+  \STATE{Remove some chromosomes with worse fitness}
+  \STATE{Apply crossover to the selected pair of chromosomes of the population}
+  \STATE{Apply mutation to each chromosome of the population}
+\ENDWHILE
+\end{algorithmic}
+\end{algorithm}
 

### 2013-10-16 paper.tex
Made some small changes in the fuzzy genetic algorithm section and also the genetic algorithm section.

@@
 \subsection{Rules}
-Rules are represented by IF-THEN statements with the following format:
+Rules are represented by IF-THEN statements in the following format:
 
-IF (cond1) AND...AND (condm) THEN class = ci
+If <condition> then <action>
 
-\mycomment{Write more here and figure out how to format the rule}
+\mycomment{Write more here}
 
@@
 Duration is the length of the connection in seconds.
+
 src\_bytes is the number of bytes sent from source to destination.
@@
 
-One record (either an attack or normal activity) is passed into a rule.
-Each feature in a record is matched to one block of the rule.
-The parameters of each block measure the probability of an attack using the trapezoidal fuzzy rule shape.
-The probabilities of each block are then combined by taking the average of the probabilities and comparing them with a threshold to determine if the record is an attack class or normal class.
-
 \subsection{Algorithm Overview}
@@
 Algorithm~\ref{alg:fuzGenAlg} describes the fuzzy genetic algorithm that is used.
+One record (either an attack or normal activity) is passed into a rule.
+Each feature in a record is matched to one block of the rule.
+The parameters of each block measure the probability of an attack using the trapezoidal fuzzy rule shape.
+The probabilities of each block are then combined by taking the average of the probabilities and comparing them with a threshold to determine if the record is an attack class or normal class.
 
@@
   \FOR{each rule}
-    \FOR{each attribute}
+    \FOR{each feature}
       \STATE{prob = fuzzy();}
@@
     \ENDFOR
-    \STATE{Assign optimal fitness as the fitness of that chromosome}
+    \STATE{Assign optimal fitness as the fitness of that
+
+    chromosome}
   \ENDFOR
@@
 
+\subsection{Experimental Design and Results}
 The authors of~\cite{DBLP:journals/corr/abs-1204-1336} used the KDD99 data set.
 For the training data set, the KDD99 10\% version file was used, and for the testing data set, the KDD99 corrected version file was used.
-
-\subsection{Experimental Design and Results}
+\mycomment{Write more here}
 

### 2013-10-18 paper.tex
Changed what's in some of the tables.

@@
 It is composed of 41 \emph{features}, which are things that help to distinguish normal connections from attacks.
+The experiments that~\cite{6559603, DBLP:journals/corr/abs-1204-1336} ran, which I will describe in Sections~\ref{sec:fuzGenAlgImp} and~\ref{sec:genAlgImp}, used the KDD99 data set.
 
@@
 The detection rate of the testing data set increased to 97.92\%.
-The false positive rate was 1.13\% and the false negative rate was 4.10\%.
 Detection of an attack happened 2-3 seconds after the packet data arrived at the intrusion detection system.
@@
 Three attack types were used for the unknown testing data set.
-For example, the first test case used the training data set that does not have Advance Port Scan, Ack Scan, and Xmas Tree, (which are all probe attacks).
+For example, test case 1 used the training data set that does not have UDP Flood, Host Scan, and UDP Scan, (the first is DoS, and the last two are probe).
 These three attacks were then used for the testing data set.
@@
 The fuzzy genetic algorithm has an accuracy of at least 95\% in identifying unknown attacks.
-When compared with the decision tree algorithm, the fuzzy genetic algorithm has a higher detection rate in all cases except 3 and 5.
-It can be seen that in cases 2 and 4 the decision tree has low detection rates, while the fuzzy genetic algorithm has much higher detection rates.
+When compared with the decision tree algorithm, the fuzzy genetic algorithm has a higher detection rate in all cases except 2 and 4.
+It can be seen that in cases 1 and 3 the decision tree has low detection rates, while the fuzzy genetic algorithm has much higher detection rates.
 
@@
 \caption{Unknown Attack Experiment}
-\begin{tabular}{|lllcrl|} \hline
-Test & Unknown & Decision & & Fuzzy & \\
-Case & Attacks & Tree DR & & Genetic DR & \\ \hline
+\begin{tabular}{|llcc|} \hline
+Test & Unknown & Decision & Fuzzy\\
+Case & Attacks & Tree DR  & Genetic DR\\ \hline
 
-1 & Advance Port Scan & 99.7 & Avg = & 100 & Avg =\\
-  & Ack Scan          & 100  & 98.33 & 100 & 100\\
-  & Xmas Tree         & 95.3 &       & 100 &\\ \hline
+1 & UDP Flood & Avg = & Avg =\\
+  & Host Scan & 46.65 & 99.80\\
+  & UDP Scan  &       &\\ \hline
 
-2 & UDP Flood & \textbf{26.7} & Avg = & \textbf{100} & Avg =\\
-  & Host Scan & 98.15         & 46.65 & 99.20        & 99.80\\
-  & UDP Scan  & \textbf{17.6} &       & \textbf{100} &\\ \hline
+2 & Jping    & Avg =          & Avg =\\
+  & Syn Scan & 99.70 & 98.75\\
+  & Fin Scan &                &\\ \hline
 
-3 & Jping    & 100 & Avg =          & 99.20 & Avg =\\
-  & Syn Scan & 99.1& \textbf{99.70} & 96.60 & 98.75\\
-  & Fin Scan & 100 &                & 100 &\\ \hline
+3 & UDP Flood & Avg = & Avg =\\
+  & RCP Scan  & 70.35 & 98.15\\
+  & Fin Scan  &       &\\ \hline
 
-4 & UDP Flood & \textbf{26.6} & Avg = & \textbf{100} & Avg =\\
-  & RCP Scan  & 91.56         & 70.35 & 92.60        & 98.15\\
-  & Fin Scan  & 99.8          &       & 100          &\\ \hline
-
-5 & Http Flood & 85.7& Avg =          & 100 & Avg =\\
-  & RCP Scan  & 92.3 & \textbf{99.94} & 91.80 & 97.50\\
-  & Fin Scan  & 99.8 &                & 98.20 &\\
+4 & Http Flood & Avg =          & Avg =\\
+  & RCP Scan  & 99.94 & 97.50\\
+  & Fin Scan  &                &\\
 \hline\end{tabular}
@@
 \begin{tabular}{|lll|} \hline
-Attack & Training & Testing\\
-Type   & Data set & Data set\\ \hline
+Attack & Training & Testing\\ \hline
 Normal & 39,387 & 39,337\\
@@
 \begin{tabular}{|lll|} \hline
-Attack & Training & Testing\\
-Type   & Data set & Data set\\ \hline
+Attack & Training & Testing\\ \hline
 Normal & 8,000 & 16,000\\
@@
 There were 6 different types of denial of service attacks and 4 different types of probe attacks.
-The detection rate of the KDD99 data set was 98.72\%, and it had a false negative rate of 1.55\%.
+The detection rate of the KDD99 data set was 98.72\%.
 Then 26,500 records of the RLD09 data set were used as the training set.
@@
 \begin{tabular}{|cccccc|} \hline
-Test & Attack & Type & FP & FN & DR\\
-Case & Name   &&&&\\ \hline
+Test & Attack & Type & FP & FN & DR\\ \hline
 1 & Back & DoS & 85.33 & 0.00 & 16.56\\
-2 & Land & DoS & 0.39 & 0.00 & 99.61\\
-3 & Smurf & DoS & 0.76 & 0.10 & 99.73\\
-4 & Portsweep & Probe & 6.40 & 0.00 & 93.66\\
-5 & Satan & Probe & 0.74 & 3.75 & 99.22\\
+2 & Smurf & DoS & 0.76 & 0.10 & 99.73\\
+3 & Portsweep & Probe & 6.40 & 0.00 & 93.66\\
+4 & Satan & Probe & 0.74 & 3.75 & 99.22\\
 \hline\end{tabular}
@@
 The results showed that the detection rate of a majority of the cases were greater than 97.5\%.
-Again, there were only two cases that had low detection rates, one of which is case 2 in Table~\ref{tab:rldAttacks}.
+Again, there were only two test cases that had low detection rates, one of which is case 2 in Table~\ref{tab:rldAttacks}.
 
@@
 \begin{tabular}{|cccccc|} \hline
-Test & Attack & Type & FP & FN & DR\\
-Case & Name   &&&&\\ \hline
+Test & Attack & Type & FP & FN & DR\\ \hline
 1 & Smurf & DoS & 0.02 & 0 & 99.98\\
@@
 3 & Ackscan & Probe & 0.03 & 0 & 99.97\\
-4 & Ipscan & Probe & 13.01 & 16.4 & 86.89\\
-5 & Synscan & Probe & 0.65 & 4.2 & 99.24\\
+4 & Synscan & Probe & 0.65 & 4.2 & 99.24\\
 \hline\end{tabular}

### 2013-10-27 paper.tex
Made some changes based on feedback from Nic.

@@
 
-\keywords{Intrusion detection, genetic algorithm, fuzzy logic, KDD99, RLD09}
+\keywords{Intrusion detection, genetic algorithm, fuzzy logic, KDD99, RLD09, computer security}
 
@@
 There are two types of intrusion detection systems: \emph{host based} and \emph{network based}.
-Host based intrusion detection systems look at and analyze information that is on a single host or multiple host systems, such as the contents of file systems and system calls to determine if there is a possible threat to the system.
-Network based intrusion detection looks at information that is from the network.
-Network based is useful for analyzing the traffic of multiple systems all at once.
-This type of intrusion detection system analyzes packets that travel through the network.~\cite{DBLP:journals/corr/abs-1204-1336}
+Host based intrusion detection systems look at and analyze information on a single host to determine if there is a possible threat to the system.
+The information that is analyzed include file systems and system calls.
+Network based intrusion detection systems look at information from multiple systems all at once.
+The information that is analyzed is the packets that pass through the network.~\cite{DBLP:journals/corr/abs-1204-6416}
 
 \subsection{Types of Networking Attacks}
-There are four different types of networking attacks: \emph{denial of service}, \emph{remote to user attacks}, \emph{user to root attacks}, and \emph{probing}.
-Each attack that happens on a network can be placed into one of these categories.
+There are four main types of networking attacks that this paper will address: \emph{denial of service}, \emph{remote to user attacks}, \emph{user to root attacks}, and \emph{probing}.
+Each attack that happens on a network can be placed into one of these categories.~\cite{DBLP:journals/corr/abs-1204-1336}
 
-Denial of service attacks happen when an attacker makes a machine inaccessible to a user by making it too busy to serve legitimate network requests.
-Remote to user attacks happen when an attacker sends packets to a machine over the network in order to get access to things a local user would have on the machine.
+Denial of service (DoS) attacks happen when an attacker makes a machine inaccessible to a user by making it too busy to serve legitimate requests.
+For example, many systems lock out a user from an account after a certain number of failed login attempts.
+An attacker would be able to use this to prevent legitimate users from logging in.~\cite{DoSAttacks} Remote to user attacks happen when an attacker sends packets to a machine over the network in order to gain access to things a local user would have on the machine.
+An example is when an attacker tries to gain access to a machine by guessing possible usernames and passwords.
 User to root attacks happen when an attacker starts out with access on the machine and then tries to gain root access to the system.
-Probing happens when an attacker examines a machine in order to collect information about weaknesses or vulnerabilities that in the future could be used in such a way that they compromise the system.~\cite{DBLP:journals/corr/abs-1204-1336}
+For example, if a program expects a user to input their name, the programmer has to decide how many characters that name buffer will require.
+Assume the program allocates 20 characters for the name buffer.
+Suppose the user's name has 35 characters.
+The last 15 characters will overflow the buffer, which will then overwrite the instructions that are to be executed next.
+An attacker can cause commands to be executed by manipulating the data that overflows.
+Probing happens when an attacker examines a machine in order to collect information about weaknesses or vulnerabilities that in the future could be used to compromise the system.~\cite{DBLP:journals/corr/abs-1204-1336, typesOfAttacks}
 
@@
 It is a simple and effective way to detect known attacks.
-A disadvantage of it is that it is an ineffective way of detecting unknown attacks.
 Signature-based detection is also called \emph{knowledge-based detection} or \emph{misuse detection}.~\cite{Liao201316}
@@
 Anomaly-based detection looks for patterns of activity that are rare and uncommon.
-It is an effective way to detect new vulnerabilities.
-Some disadvantages of this type of system is that they are expensive and they can detect an intrusive behavior as being normal because of insufficient data.
+It is an effective way to detect new attacks.
 Anomaly-based detection is also called \emph{behavior-based detection}.~\cite{DBLP:journals/corr/abs-1204-1336}
@@
 \subsection{KDD99 Data Set}
-Knowledge Discovery in Databases (KDD) is the process of extracting useful and interesting information from data in order to understand patterns in the data, with a limited amount of human intervention.
-Data is collected from several different sources and then is analyzed in order to produce patterns or models over the data.~\cite{bc1_ecindm}
-
-The KDD99 data set is a benchmark data set that was simulated in a military network environment in 1999~\cite{6559603}.
-It is composed of 41 \emph{features}, which are things that help to distinguish normal connections from attacks.
+The KDD99 data set is a benchmark data set that was generated by simulating a military network environment in 1999~\cite{6559603}.
+It has long been a standard data set for intrusion detection.
+The data in the set is classified as normal or attack activity.
+KDD99 uses 41 \emph{features}, which are properties of a \emph{record} (either an attack or normal activity), that are used to describe the activity.
 The experiments that~\cite{6559603, DBLP:journals/corr/abs-1204-1336} ran, which I will describe in Sections~\ref{sec:fuzGenAlgImp} and~\ref{sec:genAlgImp}, used the KDD99 data set.
 
-In an experiment, the researchers divide the data set into two subsets, a \emph{training set} and a \emph{testing set}.
-Algorithms are trained on the training set to look for patterns.
-These patterns are then verified when the testing set is used.~\cite{bc1_ecindm} Four different measures are then used to determine the accuracy of an algorithm, which are described in Secion~\ref{sec:PosNeg}.
+\mycomment{Say something about the features.}
 
@@
 \label{sec:PosNeg}
-\emph{False positives} and \emph{False negatives} are two values that are used to evaluate the accuracy of an intrusion detection system.
+In a machine learning experiment, a common technique is to divide the data set into two subsets, a \emph{training set} and a \emph{testing set}.
+The given algorithm is then trained on the training set to look for patterns.
+These patterns are then verified using the test set is used.~\cite{bc1_ecindm} Four different measures are then used to determine the accuracy of the algorithm in question.
+
+\emph{False positives} and \emph{False negatives} are two concepts that are used to evaluate the accuracy of an intrusion detection system.
 A false positive happens when an intrusion detection system incorrectly identifies normal activity as being an attack.
@@
 
-\subsection{Setup}
-The authors of~\cite{6496342, 6559603} used 8 out of the 41 KDD99 features in their system: duration, src\_bytes, num\_failed\_logins, root\_shell, num\_access\_files, srv\_count, serror\_rate, same\_srv\_rate.
-
-Duration is the length of the connection in seconds.
-
-src\_bytes is the number of bytes sent from source to destination.
-num\_failed\_logins is the number of failed login attempts.
-root\_shell returns 1 if root shell is obtained and 0 otherwise.
-num\_access\_files is the number of operations on access control files.
-srv\_count is the number of connections to the same service as the current connection in the past two seconds.
-serror\_rate is the percentage of connections that have ``SYN" errors.
-\mycomment{What are SYN errors?}
- same\_srv\_rate is the percentage of connections to the same service.~\cite{KDD99Features}
-
-\subsection{Fuzzy Algorithm}
 The focus of~\cite{6496342, 6559603} is on detecting new or unknown types of attacks in a network.
@@
 This kind of algorithm is able to learn new attacks, and has a high detection rate.
-The system used is evaluated in terms of detection speed, detection rate, and false alarm rate.
+The authors of~\cite{6496342, 6559603} used 8 out of the 41 KDD99 features in their system:
+\begin{enumerate}
+  \item duration: the length of the normal or attack activity in seconds.
+  \item src\_bytes: the number of bytes sent from source to destination. Source is the user who may or may not be an attacker, and destination is the server being potentially attacked.
+  \item num\_failed\_logins: the number of failed login attempts.
+  \item root\_shell: returns 1 if root shell is obtained, which means that the user is able to log in as root.
+This gives them the ability to do things like add accounts and change user passwords.
+If root shell is not obtained, 0 is returned.
+  \item num\_access\_files: the number of operations on access control files.
+Access control files specify which users are granted access to objects and what operations are allowed on the objects.
+An example would be (Alice, delete), which would give Alice permission to delete the file.~\cite{accessControl}
+  \item srv\_count: the number of connections to the same service as the current connection in the past two seconds.
+  \item serror\_rate: the percentage of connections that have ``SYN" errors.
+\mycomment{What are SYN errors?} When a client attempts to connect to a server, it first sends a SYN (synchronize) message to the server.
+The server then acknowledges the request by sending a SYN-ACK to the client.
+The connection is established when the client sends an ACK back to the server.
+A SYN error is a failure that happens early in this process.~\cite{TCP}
+  \item same\_srv\_rate: the percentage of connections to the same service.
+\end{enumerate}
 
-In order to measure the probability of an attack, a trapezoidal shape was used in the algorithm. The trapezoidal shape has four parameters: a, b, c, and d. 
-\mycomment{Put in trapezoidal shape figure.} Algorithm~\ref{alg:fuzAlg} calculates the probability of being attacked.
+~\cite{KDD99}
 
+\subsection{Fuzzy Algorithm}
+In order to measure the probability that a record is an attack, a trapezoidal shape was used in the algorithm. The trapezoidal shape has four parameters: $a, b, c$, and $d$. 
+\mycomment{Put in trapezoidal shape figure.} Algorithm~\ref{alg:fuzAlg} calculates the probability of a record being an attack.
+
+\mycomment{Figure out how to float algorithms.}
 \begin{algorithm}
@@
 \begin{algorithmic}
-\IF{data value is between b and c}
+\IF{data value is between $b$ and $c$}
   \STATE{prob = 1.0}
-\ELSIF{data value is between a and b}
+\ELSIF{data value is between $a$ and $b$}
   \STATE{
@@
   }
-\ELSIF{data value is between c and d}
+\ELSIF{data value is between $c$ and $d$}
   \STATE{
@@
 
-The four parameters a, b, c, and d are encoded into blocks of binary strings, where each block is a feature with parameters between 0.0 and 7.0.
+The four parameters a, b, c, and d are encoded into blocks of binary strings, where each block is a feature with values between 0.0 and 7.0.
 See Figure~\ref{fig:fuzEncodingForFeature} for an example of a block.
 A rule has 12 blocks of features, and at the end of the string is the type of attack.
+\mycomment{Put in an example/figure of a rule.}
 
@@
 \subsection{Algorithm Overview}
-The algorithm that is used in~\cite{6496342, 6559603} first randomly generates a rule.
-Then the rule is improved in the training phase.
-After that, the rules are used to classify the data into classes in the testing phase.
+The algorithm that is used in~\cite{6496342, 6559603} first randomly generates rules.
+Then the rules are improved in the training phase, which can be seen in Algorithm~\ref{alg:fuzGenAlg}.
+After that, the rules are used to classify the data in the testing phase.
 Algorithm~\ref{alg:fuzGenAlg} describes the fuzzy genetic algorithm that is used.
@@
 The parameters of each block measure the probability of an attack using the trapezoidal fuzzy rule shape.
-The probabilities of each block are then combined by taking the average of the probabilities and comparing them with a threshold to determine if the record is an attack class or normal class.
+The probabilities of each block are then compared with a threshold to determine if the record represents an attack or normal behavior.
 
@@
     \FOR{each feature}
-      \STATE{prob = fuzzy();}
+      \STATE{prob = fuzzy(); // Algorithm~\ref{alg:fuzAlg}}
       \STATE{totalprob = totalprob + prob;}
@@
   \STATE{compare the predicted result with actual result}
-  \STATE{find A, B, $\alpha$, and $\beta$}
-  \STATE{// A is total number of attack records.
-B is total number of normal records.
+  \STATE{find $A$, $B$, $\alpha$, and $\beta$}
+  \STATE{// $A$ is total number of attack records.
+$B$ is total number of normal records.
 $\alpha$ is total number of attack records correctly identified as attack.
@@
 \end{equation*}
-In the implementation, a population size of 10 was used for each generation.
+In this implementation, a population size of 10 was used for each generation.
 An individual in the population represents a possible detection rule.
 The two best individuals from a present generation are preserved for the next generation.
-The other individuals in the new generation come from single-point crossover.
+The other individuals in the new generation come from mutation and single-point crossover.
 
@@
 \subsubsection{RLD09 Data Set}
+\mycomment{Possibly move this section to KDD99 section.}
 The KDD99 data set is 14 years old, and newer attack types are not included in it because of its age.
@@
 The data has around ten million preprocessed data packets.
-It has 17 different types of attacks, as well as normal network activity.
-The attacks can be divided into denial of service attacks and probe attacks.
+It has 17 different types of attacks that can be divided into denial of service attacks and probe attacks.
+It also has normal network activity.
 A packet sniffer was used to get information about TCP, UDP, and ICMP headers from protocol packets.
-Then this information was processed into 12 features.
-The features are the number of TCP, UDP, and ICMP packets, the number of TCP and UDP source ports, the number of TCP and UDP destination ports, and the number of TCP fin flags, syn flags, push flags, ack flags, urgent flags.
+Then this information was processed into 12 features:
+\begin{itemize}
+  \item Features of TCP
+
+    \# of packets
+
+    \# of source ports
+
+    \# of destination ports
+
+    \# of fin flags, syn flags, push flags, ack flags, urgent flags
+  \item Features of UDP
+
+    \# of packets
+
+    \# of source ports
+
+    \# of destination ports
+  \item Features of ICMP
+
+    \# of ICMP packets
+\end{itemize}
 
@@
 The experiments that the authors of~\cite{6496342, 6559603} performed used a total of 16,000 records of normal activity and 10,500 records of attack activity.
-Of the attack activity, 4,000 of them were denial of service attacks and 6,500 were probe attacks.
+Of the attack records, 4,000 of them were denial of service attacks and 6,500 were probe attacks.
 
@@
 10,000 records were used for the training set and all 26,500 records were used for the testing set.
-The two steps in the training process were to find a denial of service rule, and find a probe rule from the training set.
-Both of these rules were then used in the testing process to identify attacks from the testing data set.
-Algorithm~\ref{alg:testProc} was used in the testing process.
-
-\begin{algorithm}
-\caption{}
-\label{alg:testProc}
-\begin{algorithmic}
-\IF{dos\_rule = yes or probe\_rule = yes}
-  \STATE{This record is an attack;}
-\ELSE \STATE{This record is normal;}
-\ENDIF
-\end{algorithmic}
-\end{algorithm}
+The two steps in the training process were to find a denial of service rule, and find a probe rule.
+Both of these rules were then used together in the testing process to identify attacks from the testing data set.
 
@@
 The detection rate of the testing data set increased to 97.92\%.
-Detection of an attack happened 2-3 seconds after the packet data arrived at the intrusion detection system.
 Results from this experiment are shown in Table~\ref{tab:fuzGenExp1}.
@@
 \begin{table*}
+\center
 \caption{Results from Experiment 1}
-\begin{tabular}{|ccccccc|} \hline
- & Attack & Normal & Total Records & FP & FN & DR\\
+\begin{tabular}{ccccccc}
+ & Attack & Normal & Total Records & FP (\%) & FN (\%) & DR (\%)\\
 DoS Training & 1499 & 8501 & 10000 & 1.46 & 47.50 & 91.64\\
 Probe Training & 2496 & 7504 & 10000 & 1.83 & 15.38 & 94.79\\
-Testing & 10500 & 1600 & 26500 & 1.13 & 4.10 & 97.92\\
-\hline\end{tabular}
+Testing & 10500 & 16000 & 26500 & 1.13 & 4.10 & 97.92\\
+\end{tabular}
 \label{tab:fuzGenExp1}
+\center
 \end{table*}
@@
 Three attack types were used for the unknown testing data set.
-For example, test case 1 used the training data set that does not have UDP Flood, Host Scan, and UDP Scan, (the first is DoS, and the last two are probe).
+For example, test case 1 used the training data set that does not have Advance Port Scan, Ack Scan, and Xmas Tree, which are all probe attacks.
 These three attacks were then used for the testing data set.
-Table~\ref{tab:fuzGenExp2} shows some of the results from the fuzzy genetic algorithm and a decision tree algorithm.
-The fuzzy genetic algorithm has an accuracy of at least 95\% in identifying unknown attacks.
-When compared with the decision tree algorithm, the fuzzy genetic algorithm has a higher detection rate in all cases except 2 and 4.
-It can be seen that in cases 1 and 3 the decision tree has low detection rates, while the fuzzy genetic algorithm has much higher detection rates.
+Table~\ref{tab:fuzGenExp2} shows some of the results from the fuzzy genetic algorithm and a decision tree algorithm, which is another common way of addressing these problems.
+For further information, see~\cite{decisionTree}.
+When compared with the decision tree algorithm, the fuzzy genetic algorithm has a higher detection rate in all cases except 3 and 5.
+It can be seen that in cases 2 and 4 the decision tree has low detection rates, while the fuzzy genetic algorithm has much higher detection rates.
 
@@
 Test & Unknown & Decision & Fuzzy\\
-Case & Attacks & Tree DR  & Genetic DR\\ \hline
+Case & Attacks & Tree DR (\%)  & Genetic DR (\%)\\ \hline
 
-1 & UDP Flood & Avg = & Avg =\\
+1 & Advance Port Scan & Avg = & Avg =\\
+  & Ack Scan		  & 98.33 & 100\\
+  & Xmas Tree		  &		  &\\ \hline
+
+2 & UDP Flood & Avg = & Avg =\\
   & Host Scan & 46.65 & 99.80\\
@@
 
-2 & Jping    & Avg =          & Avg =\\
+3 & Jping    & Avg =          & Avg =\\
   & Syn Scan & 99.70 & 98.75\\
@@
 
-3 & UDP Flood & Avg = & Avg =\\
+4 & UDP Flood & Avg = & Avg =\\
   & RCP Scan  & 70.35 & 98.15\\
@@
 
-4 & Http Flood & Avg =          & Avg =\\
+5 & Http Flood & Avg =          & Avg =\\
   & RCP Scan  & 99.94 & 97.50\\
@@
 The authors of~\cite{6496342, 6559603} also ran experiments that used and compared the RLD09 data set with the KDD99 data set.
-They used the KDD99 10\% version file for both the training dataset and testing dataset.
-The number of records of normal and attack activity for the KDD99 data set is shown in Table~\ref{tab:RecordsForKDD99}, and the number of records of normal and attack activity for the RLD09 data is shown in Table~\ref{tab:RecordsForRLD09}.
-
-\begin{table}
-\caption{Number of Records for KDD99}
-\begin{tabular}{|lll|} \hline
-Attack & Training & Testing\\ \hline
-Normal & 39,387 & 39,337\\
-DoS & 158,597 & 158,503\\
-Probe & 1,550 & 1,674\\
- & &\\
- Total & 199,534 & 199,514\\
-\hline\end{tabular}
-\label{tab:RecordsForKDD99}
-\end{table}
-
-\begin{table}
-\caption{Number of Records for RLD09}
-\begin{tabular}{|lll|} \hline
-Attack & Training & Testing\\ \hline
-Normal & 8,000 & 16,000\\
-DoS & 2,400 & 4,000\\
-Probe & 3,900 & 6,500\\
- & &\\
- Total & 14,300 & 26,500\\
-\hline\end{tabular}
-\label{tab:RecordsForRLD09}
-\end{table}
+They used the KDD99 10\% version file~\cite{KDD99} for both the training dataset and testing dataset.
 
@@
 \begin{tabular}{|cccccc|} \hline
-Data & Testing & Testing & False    & False    & Detection\\
-set  & Attack  & Normal  & Positive & Negative & Rate\\ \hline
+Data set & Attack & Normal & FP (\%) & FN (\%) & DR (\%)\\ \hline
 KDD99 & 160,117 & 39,337 & 0.13 & 1.55 & 98.72\\
@@
 \begin{tabular}{|cccccc|} \hline
-Test & Attack & Type & FP & FN & DR\\ \hline
+Test & Attack & Type & FP (\%) & FN (\%) & DR (\%)\\ \hline
 1 & Back & DoS & 85.33 & 0.00 & 16.56\\
@@
 \begin{tabular}{|cccccc|} \hline
-Test & Attack & Type & FP & FN & DR\\ \hline
+Test & Attack & Type & FP (\%) & FN (\%) & DR (\%)\\ \hline
 1 & Smurf & DoS & 0.02 & 0 & 99.98\\
@@
 \mycomment{Figure out better title for Section~\ref{sec:genAlgImp}}
+
+\mycomment{Switch Sections~\ref{sec:fuzGenAlgImp} and~\ref{sec:genAlgImp}}
 
@@
 \STATE{Initialize the population}
-\STATE{crossoverRate = 0.15, mutationRate = 0.35}
 \WHILE{number of generation is not reached}

### 2013-10-27 paper.tex
Added the trapezoidal shape to my paper.

@@
 \subsection{Fuzzy Algorithm}
-In order to measure the probability that a record is an attack, a trapezoidal shape was used in the algorithm. The trapezoidal shape has four parameters: $a, b, c$, and $d$. 
-\mycomment{Put in trapezoidal shape figure.} Algorithm~\ref{alg:fuzAlg} calculates the probability of a record being an attack.
+In order to measure the probability that a record is an attack, a trapezoidal shape was used in the algorithm.
+This is shown in Figure~\ref{fig:trapFigure}.
+The trapezoidal shape has four parameters: $a, b, c$, and $d$.
+Algorithm~\ref{alg:fuzAlg} calculates the probability of a record being an attack.
+
+\begin{figure}
+\centering
+\psfig{file=trapFigure.jpg}
+\caption{Trapezoidal shape with 4 parameters where $a\leq b\leq c\leq d$}
+\label{fig:trapFigure}
+\end{figure}
 
@@
 A rule has 12 blocks of features, and at the end of the string is the type of attack.
-\mycomment{Put in an example/figure of a rule.}
+An example of this is Figure~\ref{fig:rule}.
 
@@
 \caption{Fuzzy encoding for a feature}
-\begin{tabular}{|c|c|c|c|} \hline
+\begin{tabular}{|cccc|} \hline
 010 & 011 & 100 & 101\\
-a & b & c & d\\
+a=2 & b=3 & c=4 & d=5\\
 \hline\end{tabular}
@@
 \end{figure}
+
+\begin{figure*}
+\centering
+\caption{A rule with 12 blocks of features}
+\begin{tabular}{|cccc|c|cccc|c|} \hline
+010 & 011 & 100 & 101   & ...... & 010 & 011 & 101 & 111   & DoS\\
+a=2 & b=3 & c=4 & d=5   & ...... & a=2 & b=3 & c=5 & d=7   &\\ 
+    &     & Block 1&    &        &     & Block 12& &       & Type\\
+\hline\end{tabular}
+\label{fig:rule}
+\end{figure*}
 

### 2013-10-27 paper.tex
Wrote some stuff in the rules section and fuzzy logic section.

@@
 
-\subsection{Genetic Algorithm}
+\subsection{Rules}
+Rules are a way in which elements of one set are separated into different sets, or classes, in order to differentiate between normal connections and attacks.
+Rules are represented by if-then statements in the following format: If <\emph{condition}> then <\emph{action}>.~\cite{bapuji2012soft} Rules can specify the details of a packet such as the IP address, port number and protocal.
+If a packet matches any of the rules in the intrusion detection system then the system will take appropriate action, which may include stopping the connection or logging off the system.~\cite{DBLP:journals/corr/abs-1204-6416}
 
 \subsection{Fuzzy Logic}
+Attacks on systems do not always have a fixed patter, so fuzzy logic is used to detect patterns that have a behavior that is between normal and unusual.
+A fuzzy logic rule has the following format: If <\emph{condition}> then <\emph{consequence}>, where \emph{condition} is a fuzzy variable, and \emph{consequence} is a fuzzy set.
+An example of this is: if the number of packets with the same destination address is high, then the pattern is unusual.
+To determine what is considered high, the values of the packets are divided into fuzzy sets.
+In Figure~\ref{fig:triangleFigure} there are three sets: low (region A), medium (region B), and high (region C).
+\begin{figure}
+\centering
+\psfig{file=triangleFigure.JPG}
+\caption{Fuzzy shape with three sets: low (A), medium (B), and high (C)}
+\label{fig:triangleFigure}
+\end{figure}
+The x-axis are the values in the fuzzy set (number of packets), and the y-axis is the membership function, which determines the region.
+For example, if the number of packets with the same destination address is 15 then this will be considered as low for a degree of 0.4, but will be considered medium or high for other degrees.
+If the number of packets is determined to be high, then the connection will be aborted.~\cite{DBLP:journals/corr/abs-1204-6416}
 
-\subsection{Rules}
-Rules are represented by IF-THEN statements in the following format:
-
-If <condition> then <action>
-
-\mycomment{Write more here}
+\subsection{Genetic Algorithm}
 

### 2013-10-27 paper.tex
Wrote some things in the genetic algorithm section, also wrote about true positives and true negatives.

@@
 \psfig{file=triangleFigure.JPG}
-\caption{Fuzzy shape with three sets: low (A), medium (B), and high (C)}
+\caption{Fuzzy shape with three sets: low (A), medium (B), and high (C)~\cite{DBLP:journals/corr/abs-1204-6416}}
 \label{fig:triangleFigure}
@@
 \subsection{Genetic Algorithm}
+Genetic algorithms are a search technique used to find solutions to problems \emph{Mutation}, \emph{selection}, and \emph{crossover} are used to evolve and improve rules.
+
+Mutation is where random bits in a solution are changed.
+Selection is where solutions that have a better \emph{fitness} are chosen over the other solutions.
+The fitness function determines the quality of a particular solution.
+Crossover is where two solutions swap one of their characteristics with the other to form two new solutions.
+
+The solution to a problem is represented as a chromosome.
+First, a randomly generated population of chromosomes is created.
+According to the characteristics of the problem, the positions of each chromosome are encoded as bits, characters, or numbers.
+Then mutation, selection, and crossover are applied to each generation and eventually the best solution is found.~\cite{DBLP:journals/corr/abs-1204-6416}
 
@@
 The given algorithm is then trained on the training set to look for patterns.
-These patterns are then verified using the test set is used.~\cite{bc1_ecindm} Four different measures are then used to determine the accuracy of the algorithm in question.
+These patterns are then verified using the test set.~\cite{bc1_ecindm} Four different measures are then used to determine the accuracy of the algorithm in question.
 
-\emph{False positives} and \emph{False negatives} are two concepts that are used to evaluate the accuracy of an intrusion detection system.
-A false positive happens when an intrusion detection system incorrectly identifies normal activity as being an attack.
-A false negative happens when an intrusion detection system fails to identify harmful activity.~\cite{Liao201316}
+A \emph{false positive} (FP) happens when an intrusion detection system incorrectly identifies normal activity as being an attack.
+A \emph{false negative} (FN) happens when an intrusion detection system fails to identify harmful activity.
+A \emph{true positive} (TP) happens when an intrusion detection system correctly identifies activities to be attacks.
+A \emph{true negative} (TN) happens when an intrusion detection system correctly identifies activities to be normal.
 
-\mycomment{Write more here about True Pos, True Neg, Detection Rate}
+The \emph{detection rate} (DR) of an intrusion detection system is the number of true positives divided by the total number of intrusions that happen.~\cite{ids}
 
@@
 
-\mycomment{Fix tables}
-
 \begin{table*}

### 2013-10-28 paper.tex
Added more to the GA section

@@
 \usepackage{algorithm}
+\usepackage{multirow}
 
@@
 For example, many systems lock out a user from an account after a certain number of failed login attempts.
-An attacker would be able to use this to prevent legitimate users from logging in.~\cite{DoSAttacks} Remote to user attacks happen when an attacker sends packets to a machine over the network in order to gain access to things a local user would have on the machine.
+An attacker would be able to use this to prevent legitimate users from logging in.~\cite{DoSAttacks} Remote to user (R2L) attacks happen when an attacker sends packets to a machine over the network in order to gain access to things a local user would have on the machine.
 An example is when an attacker tries to gain access to a machine by guessing possible usernames and passwords.
-User to root attacks happen when an attacker starts out with access on the machine and then tries to gain root access to the system.
+User to root (U2R) attacks happen when an attacker starts out with access on the machine and then tries to gain root access to the system.
 For example, if a program expects a user to input their name, the programmer has to decide how many characters that name buffer will require.
@@
     \FOR{each precalculated chromosome}
-      \STATE{Find fitness}
+      \STATE{Find fitness // Fitness function is the standard deviation equation with distance}
     \ENDFOR
@@
 The authors of~\cite{DBLP:journals/corr/abs-1204-1336} used the KDD99 data set.
-For the training data set, the KDD99 10\% version file was used, and for the testing data set, the KDD99 corrected version file was used.
-\mycomment{Write more here}
+They used only the numerical features of the KDD99 data set (34 out of 41 total features).
+For the training data set, the KDD99 10\% version file was used, and for the test data set, the KDD99 corrected version file was used.
+Table~\ref{tab:numberOfRecords} shows the number of records of normal and attack activity in the training and test sets.
+
+\begin{table}
+\center
+\caption{Number of records}
+\begin{tabular}{lll}
+  & Training & Testing \\ 
+Normal & 97,280 & 60,593\\
+Probe  & 4,107  & 4,166\\
+DoS	   & 391,458 & 229,853\\
+U2R    & 52      & 228\\
+R2L    & 1,124   & 16,189\\
+Total  & 494,021 & 311,029\\
+\end{tabular}
+\center
+\label{tab:numberOfRecords}
+\end{table}
+
+In the detection phase, an initial population is created and then is compared with each chromosome that was created in the training phase.
+Crossover and mutation happen in the population in order to create a new population.
+The results from running the genetic algorithm are shown in Tables~\ref{tab:genAlgResults} and~\ref{tab:genAlgResults2}.
+The detection rate was 94.9\%.
+The false positive rate was 30.5\%.
+The false negative rate was 5\%.
+The true negative rate was 69.5\%.
+
+\begin{table*}
+\center
+\caption{Results for Genetic Algorithm Experiment}
+\begin{tabular}{cccccccc}
+& & \multicolumn{5}{ c }{Predicted label} \\ 
+& & Normal & Probe & DoS & U2R & R2L & \% Correct \\ \cline{3-8}
+\multicolumn{1}{ c }{\multirow{2}{*}{Actual class} } &
+\multicolumn{1}{ c| }{Normal} & 42138 & 1421 & 15835 & 486 & 713 & 69.5\\
+\multicolumn{1}{ c  }{}                        &
+\multicolumn{1}{ c| }{Probe} & 398 & 2963 & 654 & 2 & 149 & 71.1   \\
+\multicolumn{1}{ c  }{\multirow{2}{*}{} } &
+\multicolumn{1}{ c| }{Dos} & 921 & 432 & 228489 & 1 & 10 & 99.4\\
+\multicolumn{1}{ c  }{}                        &
+\multicolumn{1}{ c| }{U2R} & 146 & 21 & 8 & 43 & 10 & 18.9\\
+\multicolumn{1}{ c  }{}                        &
+\multicolumn{1}{ c| }{R2L} & 11191 & 578 & 3398 & 141 & 881 & 5.4\\
+
+\multicolumn{1}{ c  }{}                        &
+\multicolumn{1}{ c| }{} &  &  &  &  &  & \\
+
+\multicolumn{1}{ c  }{\% Correct}                        &
+\multicolumn{1}{ c| }{} & 76.9 & 54.7 & 92.0 & 6.4 & 50.0 \\
+\end{tabular}
+\center
+\label{tab:genAlgResults}
+\end{table*}
+
+\begin{table*}
+\center
+\caption{Results for Genetic Algorithm Experiment}
+\begin{tabular}{cccc}
+%\cline{3-8}
+& & \multicolumn{2}{ c }{Predicted label} \\ 
+& & Normal & Intrusion\\ \cline{3-4}
+\multicolumn{1}{ c }{\multirow{2}{*}{Actual class} } &
+\multicolumn{1}{ c| }{Normal} & True Negative (42,138) & False Positive (18,455)\\
+\multicolumn{1}{ c  }{}                        &
+\multicolumn{1}{ c| }{Probe} & False Negative (12,528) & True Positive (237,908)\\
+\end{tabular}
+\center
+\label{tab:genAlgResults2}
+\end{table*}
+
+The authors of~\cite{DBLP:journals/corr/abs-1204-1336} compared their results with the winning entry of the KDD99 Classifier Learning Contest~\cite{KDD99Contest}, and found that they had a better detection rate for denial of service and user to root attacks than the winning entry.
 

### 2013-10-28 paper.tex
Changed the KDD99 section so that it talks about both data sets, and also swapped sections 3 and 4.

@@
 
-\subsection{KDD99 Data Set}
+\subsection{Data Sets and Features}
+Two different data sets are used in this paper: KDD99 and RLD09.
+
 The KDD99 data set is a benchmark data set that was generated by simulating a military network environment in 1999~\cite{6559603}.
@@
 KDD99 uses 41 \emph{features}, which are properties of a \emph{record} (either an attack or normal activity), that are used to describe the activity.
-The experiments that~\cite{6559603, DBLP:journals/corr/abs-1204-1336} ran, which I will describe in Sections~\ref{sec:fuzGenAlgImp} and~\ref{sec:genAlgImp}, used the KDD99 data set.
 
-\mycomment{Say something about the features.}
+The authors of~\cite{6496342, 6559603} used 8 out of the 41 KDD99 features~\cite{KDD99} in their system: 
+\begin{enumerate}
+  \item duration: the length of the normal or attack activity in seconds.
+  \item src\_bytes: the number of bytes sent from source to destination. Source is the user who may or may not be an attacker, and destination is the server being potentially attacked.
+  \item num\_failed\_logins: the number of failed login attempts.
+  \item root\_shell: returns 1 if root shell is obtained, which means that the user is able to log in as root.
+This gives them the ability to do things like add accounts and change user passwords.
+If root shell is not obtained, 0 is returned.
+  \item num\_access\_files: the number of operations on access control files.
+Access control files specify which users are granted access to objects and what operations are allowed on the objects.
+An example would be (Alice, delete), which would give Alice permission to delete the file.~\cite{accessControl}
+  \item srv\_count: the number of connections to the same service as the current connection in the past two seconds.
+  \item serror\_rate: the percentage of connections that have ``SYN" errors.
+When a client attempts to connect to a server, it first sends a SYN (synchronize) message to the server.
+The server then acknowledges the request by sending a SYN-ACK to the client.
+The connection is established when the client sends an ACK back to the server.
+A SYN error is a failure that happens early in this process.~\cite{TCP}
+  \item same\_srv\_rate: the percentage of connections to the same service.
+\end{enumerate}
+
+The KDD99 data set is 14 years old, and newer attack types are not included in it because of its age.
+Because of this, the authors of~\cite{6496342, 6559603} created their own data set, RLD09, to use in their experiments.
+To create the data set, the authors captured network data from the Computer Engineering Department at King Mongkut's University of Technology Thonburi, in Bangkok, Thailand.
+The data has around ten million preprocessed data packets.
+It has 17 different types of attacks that can be divided into denial of service attacks and probe attacks.
+It also has normal network activity.
+A packet sniffer was used to get information about TCP, UDP, and ICMP headers from protocol packets.
+Then this information was processed into 12 features:
+\begin{itemize}
+  \item Features of TCP
+
+    \# of packets
+
+    \# of source ports
+
+    \# of destination ports
+
+    \# of fin flags, syn flags, push flags, ack flags, urgent flags
+  \item Features of UDP
+
+    \# of packets
+
+    \# of source ports
+
+    \# of destination ports
+  \item Features of ICMP
+
+    \# of ICMP packets
+\end{itemize}
 
@@
 
+\section{Genetic Algorithm Implementation}
+\label{sec:genAlgImp}
+\mycomment{Figure out better title for Section~\ref{sec:genAlgImp}}
+
+\subsection{Algorithm Overview}
+The authors of~\cite{DBLP:journals/corr/abs-1204-1336} used a genetic algorithm to make their intrusion detection system.
+The system is divided into two phases: a precalculation phase and a detection phase.
+In the precalculation phase, a set of chromosomes are created using training data.
+Then this set of chromosomes is used in the detection phase for comparison.
+Algorithm~\ref{alg:precalc} is used in the precalculation phase.
+
+\begin{algorithm}
+\caption{Major steps in precalculation}
+\label{alg:precalc}
+\begin{algorithmic}
+\STATE{range = 0.125}
+\FOR{each training data}
+  \IF{it has neighboring chromosome within range}
+    \STATE{Merge it with the nearest chromosome}
+  \ELSE \STATE{Create new chromosome with it}
+  \ENDIF
+\ENDFOR
+\end{algorithmic}
+\end{algorithm}
+
+In the detection phase, a population is created for test data and then the type of the data is predicted.
+The set of chromosomes that was created in the precalculation phase is used in the detection phase to find the fitness of each chromosome in the population.
+Algorithm~\ref{alg:detection} is used in the detection phase.
+
+\begin{algorithm}
+\caption{Major steps in detection}
+\label{alg:detection}
+\begin{algorithmic}
+\STATE{Initialize the population}
+\WHILE{number of generation is not reached}
+  \FOR{each chromosome in the population}
+    \FOR{each precalculated chromosome}
+      \STATE{Find fitness // Fitness function is the standard deviation equation with distance}
+    \ENDFOR
+    \STATE{Assign optimal fitness as the fitness of that
+
+    chromosome}
+  \ENDFOR
+  \STATE{Remove some chromosomes with worse fitness}
+  \STATE{Apply crossover to the selected pair of chromosomes of the population}
+  \STATE{Apply mutation to each chromosome of the population}
+\ENDWHILE
+\end{algorithmic}
+\end{algorithm}
+
+\subsection{Experimental Design and Results}
+The authors of~\cite{DBLP:journals/corr/abs-1204-1336} used the KDD99 data set.
+They used only the numerical features of the KDD99 data set (34 out of 41 total features).
+For the training data set, the KDD99 10\% version file was used, and for the test data set, the KDD99 corrected version file was used.
+Table~\ref{tab:numberOfRecords} shows the number of records of normal and attack activity in the training and test sets.
+
+\begin{table}
+\center
+\caption{Number of records}
+\begin{tabular}{lll}
+  & Training & Testing \\ 
+Normal & 97,280 & 60,593\\
+Probe  & 4,107  & 4,166\\
+DoS	   & 391,458 & 229,853\\
+U2R    & 52      & 228\\
+R2L    & 1,124   & 16,189\\
+Total  & 494,021 & 311,029\\
+\end{tabular}
+\center
+\label{tab:numberOfRecords}
+\end{table}
+
+In the detection phase, an initial population is created and then is compared with each chromosome that was created in the training phase.
+Crossover and mutation happen in the population in order to create a new population.
+The results from running the genetic algorithm are shown in Tables~\ref{tab:genAlgResults} and~\ref{tab:genAlgResults2}.
+The detection rate was 94.9\%.
+The false positive rate was 30.5\%.
+The false negative rate was 5\%.
+The true negative rate was 69.5\%.
+
+\begin{table*}
+\center
+\caption{Results for Genetic Algorithm Experiment}
+\begin{tabular}{cccccccc}
+& & \multicolumn{5}{ c }{Predicted label} \\ 
+& & Normal & Probe & DoS & U2R & R2L & \% Correct \\ \cline{3-8}
+\multicolumn{1}{ c }{\multirow{2}{*}{Actual class} } &
+\multicolumn{1}{ c| }{Normal} & 42138 & 1421 & 15835 & 486 & 713 & 69.5\\
+\multicolumn{1}{ c  }{}                        &
+\multicolumn{1}{ c| }{Probe} & 398 & 2963 & 654 & 2 & 149 & 71.1   \\
+\multicolumn{1}{ c  }{\multirow{2}{*}{} } &
+\multicolumn{1}{ c| }{Dos} & 921 & 432 & 228489 & 1 & 10 & 99.4\\
+\multicolumn{1}{ c  }{}                        &
+\multicolumn{1}{ c| }{U2R} & 146 & 21 & 8 & 43 & 10 & 18.9\\
+\multicolumn{1}{ c  }{}                        &
+\multicolumn{1}{ c| }{R2L} & 11191 & 578 & 3398 & 141 & 881 & 5.4\\
+
+\multicolumn{1}{ c  }{}                        &
+\multicolumn{1}{ c| }{} &  &  &  &  &  & \\
+
+\multicolumn{1}{ c  }{\% Correct}                        &
+\multicolumn{1}{ c| }{} & 76.9 & 54.7 & 92.0 & 6.4 & 50.0 \\
+\end{tabular}
+\center
+\label{tab:genAlgResults}
+\end{table*}
+
+\begin{table*}
+\center
+\caption{Results for Genetic Algorithm Experiment}
+\begin{tabular}{cccc}
+%\cline{3-8}
+& & \multicolumn{2}{ c }{Predicted label} \\ 
+& & Normal & Intrusion\\ \cline{3-4}
+\multicolumn{1}{ c }{\multirow{2}{*}{Actual class} } &
+\multicolumn{1}{ c| }{Normal} & True Negative (42,138) & False Positive (18,455)\\
+\multicolumn{1}{ c  }{}                        &
+\multicolumn{1}{ c| }{Probe} & False Negative (12,528) & True Positive (237,908)\\
+\end{tabular}
+\center
+\label{tab:genAlgResults2}
+\end{table*}
+
+The authors of~\cite{DBLP:journals/corr/abs-1204-1336} compared their results with the winning entry of the KDD99 Classifier Learning Contest~\cite{KDD99Contest}, and found that they had a better detection rate for denial of service and user to root attacks than the winning entry.
+
 \section{Fuzzy Genetic Algorithm Implementation}
@@
 This kind of algorithm is able to learn new attacks, and has a high detection rate.
-The authors of~\cite{6496342, 6559603} used 8 out of the 41 KDD99 features in their system:
-\begin{enumerate}
-  \item duration: the length of the normal or attack activity in seconds.
-  \item src\_bytes: the number of bytes sent from source to destination. Source is the user who may or may not be an attacker, and destination is the server being potentially attacked.
-  \item num\_failed\_logins: the number of failed login attempts.
-  \item root\_shell: returns 1 if root shell is obtained, which means that the user is able to log in as root.
-This gives them the ability to do things like add accounts and change user passwords.
-If root shell is not obtained, 0 is returned.
-  \item num\_access\_files: the number of operations on access control files.
-Access control files specify which users are granted access to objects and what operations are allowed on the objects.
-An example would be (Alice, delete), which would give Alice permission to delete the file.~\cite{accessControl}
-  \item srv\_count: the number of connections to the same service as the current connection in the past two seconds.
-  \item serror\_rate: the percentage of connections that have ``SYN" errors.
-\mycomment{What are SYN errors?} When a client attempts to connect to a server, it first sends a SYN (synchronize) message to the server.
-The server then acknowledges the request by sending a SYN-ACK to the client.
-The connection is established when the client sends an ACK back to the server.
-A SYN error is a failure that happens early in this process.~\cite{TCP}
-  \item same\_srv\_rate: the percentage of connections to the same service.
-\end{enumerate}
-
-~\cite{KDD99}
 
@@
 \subsection{Experimental Design and Results}
-
-\subsubsection{RLD09 Data Set}
-\mycomment{Possibly move this section to KDD99 section.}
-The KDD99 data set is 14 years old, and newer attack types are not included in it because of its age.
-Because of this, the authors of~\cite{6496342, 6559603} created their own data set, RLD09, to use in their experiments.
-To create the data set, the authors captured network data from the Computer Engineering Department at King Mongkut's University of Technology Thonburi, in Bangkok, Thailand.
-The data has around ten million preprocessed data packets.
-It has 17 different types of attacks that can be divided into denial of service attacks and probe attacks.
-It also has normal network activity.
-A packet sniffer was used to get information about TCP, UDP, and ICMP headers from protocol packets.
-Then this information was processed into 12 features:
-\begin{itemize}
-  \item Features of TCP
-
-    \# of packets
-
-    \# of source ports
-
-    \# of destination ports
-
-    \# of fin flags, syn flags, push flags, ack flags, urgent flags
-  \item Features of UDP
-
-    \# of packets
-
-    \# of source ports
-
-    \# of destination ports
-  \item Features of ICMP
-
-    \# of ICMP packets
-\end{itemize}
 
@@
 
-\section{Genetic Algorithm Implementation}
-\label{sec:genAlgImp}
-\mycomment{Figure out better title for Section~\ref{sec:genAlgImp}}
-
-\mycomment{Switch Sections~\ref{sec:fuzGenAlgImp} and~\ref{sec:genAlgImp}}
-
-\subsection{Algorithm Overview}
-The authors of~\cite{DBLP:journals/corr/abs-1204-1336} used a genetic algorithm to make their intrusion detection system.
-The system is divided into two phases: a precalculation phase and a detection phase.
-In the precalculation phase, a set of chromosomes are created using training data.
-Then this set of chromosomes is used in the detection phase for comparison.
-Algorithm~\ref{alg:precalc} is used in the precalculation phase.
-
-\begin{algorithm}
-\caption{Major steps in precalculation}
-\label{alg:precalc}
-\begin{algorithmic}
-\STATE{range = 0.125}
-\FOR{each training data}
-  \IF{it has neighboring chromosome within range}
-    \STATE{Merge it with the nearest chromosome}
-  \ELSE \STATE{Create new chromosome with it}
-  \ENDIF
-\ENDFOR
-\end{algorithmic}
-\end{algorithm}
-
-In the detection phase, a population is created for test data and then the type of the data is predicted.
-The set of chromosomes that was created in the precalculation phase is used in the detection phase to find the fitness of each chromosome in the population.
-Algorithm~\ref{alg:detection} is used in the detection phase.
-
-\begin{algorithm}
-\caption{Major steps in detection}
-\label{alg:detection}
-\begin{algorithmic}
-\STATE{Initialize the population}
-\WHILE{number of generation is not reached}
-  \FOR{each chromosome in the population}
-    \FOR{each precalculated chromosome}
-      \STATE{Find fitness // Fitness function is the standard deviation equation with distance}
-    \ENDFOR
-    \STATE{Assign optimal fitness as the fitness of that
-
-    chromosome}
-  \ENDFOR
-  \STATE{Remove some chromosomes with worse fitness}
-  \STATE{Apply crossover to the selected pair of chromosomes of the population}
-  \STATE{Apply mutation to each chromosome of the population}
-\ENDWHILE
-\end{algorithmic}
-\end{algorithm}
-
-\subsection{Experimental Design and Results}
-The authors of~\cite{DBLP:journals/corr/abs-1204-1336} used the KDD99 data set.
-They used only the numerical features of the KDD99 data set (34 out of 41 total features).
-For the training data set, the KDD99 10\% version file was used, and for the test data set, the KDD99 corrected version file was used.
-Table~\ref{tab:numberOfRecords} shows the number of records of normal and attack activity in the training and test sets.
-
-\begin{table}
-\center
-\caption{Number of records}
-\begin{tabular}{lll}
-  & Training & Testing \\ 
-Normal & 97,280 & 60,593\\
-Probe  & 4,107  & 4,166\\
-DoS	   & 391,458 & 229,853\\
-U2R    & 52      & 228\\
-R2L    & 1,124   & 16,189\\
-Total  & 494,021 & 311,029\\
-\end{tabular}
-\center
-\label{tab:numberOfRecords}
-\end{table}
-
-In the detection phase, an initial population is created and then is compared with each chromosome that was created in the training phase.
-Crossover and mutation happen in the population in order to create a new population.
-The results from running the genetic algorithm are shown in Tables~\ref{tab:genAlgResults} and~\ref{tab:genAlgResults2}.
-The detection rate was 94.9\%.
-The false positive rate was 30.5\%.
-The false negative rate was 5\%.
-The true negative rate was 69.5\%.
-
-\begin{table*}
-\center
-\caption{Results for Genetic Algorithm Experiment}
-\begin{tabular}{cccccccc}
-& & \multicolumn{5}{ c }{Predicted label} \\ 
-& & Normal & Probe & DoS & U2R & R2L & \% Correct \\ \cline{3-8}
-\multicolumn{1}{ c }{\multirow{2}{*}{Actual class} } &
-\multicolumn{1}{ c| }{Normal} & 42138 & 1421 & 15835 & 486 & 713 & 69.5\\
-\multicolumn{1}{ c  }{}                        &
-\multicolumn{1}{ c| }{Probe} & 398 & 2963 & 654 & 2 & 149 & 71.1   \\
-\multicolumn{1}{ c  }{\multirow{2}{*}{} } &
-\multicolumn{1}{ c| }{Dos} & 921 & 432 & 228489 & 1 & 10 & 99.4\\
-\multicolumn{1}{ c  }{}                        &
-\multicolumn{1}{ c| }{U2R} & 146 & 21 & 8 & 43 & 10 & 18.9\\
-\multicolumn{1}{ c  }{}                        &
-\multicolumn{1}{ c| }{R2L} & 11191 & 578 & 3398 & 141 & 881 & 5.4\\
-
-\multicolumn{1}{ c  }{}                        &
-\multicolumn{1}{ c| }{} &  &  &  &  &  & \\
-
-\multicolumn{1}{ c  }{\% Correct}                        &
-\multicolumn{1}{ c| }{} & 76.9 & 54.7 & 92.0 & 6.4 & 50.0 \\
-\end{tabular}
-\center
-\label{tab:genAlgResults}
-\end{table*}
-
-\begin{table*}
-\center
-\caption{Results for Genetic Algorithm Experiment}
-\begin{tabular}{cccc}
-%\cline{3-8}
-& & \multicolumn{2}{ c }{Predicted label} \\ 
-& & Normal & Intrusion\\ \cline{3-4}
-\multicolumn{1}{ c }{\multirow{2}{*}{Actual class} } &
-\multicolumn{1}{ c| }{Normal} & True Negative (42,138) & False Positive (18,455)\\
-\multicolumn{1}{ c  }{}                        &
-\multicolumn{1}{ c| }{Probe} & False Negative (12,528) & True Positive (237,908)\\
-\end{tabular}
-\center
-\label{tab:genAlgResults2}
-\end{table*}
-
-The authors of~\cite{DBLP:journals/corr/abs-1204-1336} compared their results with the winning entry of the KDD99 Classifier Learning Contest~\cite{KDD99Contest}, and found that they had a better detection rate for denial of service and user to root attacks than the winning entry.
-
 \section{Conclusions}

### 2013-10-29 paper.tex
Wrote abstract, introduction, and conclusion.

@@
 \begin{abstract}
-
+This paper describes two different ways of training an intrusion detection system about possible attacks to a system: genetic algorithms and fuzzy logic.
+I will explain how genetic algorithms are used in intrusion detection systems and compare the results to the winning entry of the KDD99 Classifier Learning Contest.
+I will then explain how fuzzy genetic algorithms are used and compare those results with a decision tree algorithm.
 \end{abstract}
@@
 
-\terms{}
+\terms{Security}
 
@@
 \section{Introduction}
+An attack on important and confidential data is a concern for many people, so it is important to have a way to detect and analyze these attacks.
+A way of detecting attacks is by using an intrusion detection system.
+An intrusion detection system is a device that monitors network activities for malicious or abnormal behaviors and then produces reports~\cite{DBLP:journals/corr/abs-1204-1336}.
+There are various ways of training the intrusion detection system about the possible threats.
+Two different ways, which I will talk about in this paper, are the use of genetic algorithms and fuzzy logic.
+
+In Section~\ref{sec:background} I will give some background on intrusion detection systems.
+I will explain the types of intrusion detection systems, the main types of networking attacks, and ways of detecting attacks.
+I will explain what \emph{rules} are in Section~\ref{sec:rules}, and then give information on fuzzy logic and genetic algorithms in Sections~\ref{sec:fuzzyLogic} and~\ref{sec:GA}.
+I will then explain two data sets that are used in intrusion detection.
+After that, I will explain the measures that are used in determining the accuracy of an intrusion detection system.
 
 \section{Background}
-
-\subsection{Types of Intrusion Detection Systems}
-There are two types of intrusion detection systems: \emph{host based} and \emph{network based}.
-Host based intrusion detection systems look at and analyze information on a single host to determine if there is a possible threat to the system.
-The information that is analyzed include file systems and system calls.
-Network based intrusion detection systems look at information from multiple systems all at once.
-The information that is analyzed is the packets that pass through the network.~\cite{DBLP:journals/corr/abs-1204-6416}
+\label{sec:background}
 
 \subsection{Types of Networking Attacks}
+\label{sec:typesAttacks}
 There are four main types of networking attacks that this paper will address: \emph{denial of service}, \emph{remote to user attacks}, \emph{user to root attacks}, and \emph{probing}.
@@
 \subsection{Detection Methodologies}
+\label{sec:detectionMeth}
 There are two different ways of detecting attacks: \emph{signature-based detection} and \emph{anomaly-based detection}.
@@
 
-\subsection{Rules}
-Rules are a way in which elements of one set are separated into different sets, or classes, in order to differentiate between normal connections and attacks.
-Rules are represented by if-then statements in the following format: If <\emph{condition}> then <\emph{action}>.~\cite{bapuji2012soft} Rules can specify the details of a packet such as the IP address, port number and protocal.
-If a packet matches any of the rules in the intrusion detection system then the system will take appropriate action, which may include stopping the connection or logging off the system.~\cite{DBLP:journals/corr/abs-1204-6416}
-
-\subsection{Fuzzy Logic}
-Attacks on systems do not always have a fixed patter, so fuzzy logic is used to detect patterns that have a behavior that is between normal and unusual.
-A fuzzy logic rule has the following format: If <\emph{condition}> then <\emph{consequence}>, where \emph{condition} is a fuzzy variable, and \emph{consequence} is a fuzzy set.
-An example of this is: if the number of packets with the same destination address is high, then the pattern is unusual.
-To determine what is considered high, the values of the packets are divided into fuzzy sets.
-In Figure~\ref{fig:triangleFigure} there are three sets: low (region A), medium (region B), and high (region C).
-\begin{figure}
-\centering
-\psfig{file=triangleFigure.JPG}
-\caption{Fuzzy shape with three sets: low (A), medium (B), and high (C)~\cite{DBLP:journals/corr/abs-1204-6416}}
-\label{fig:triangleFigure}
-\end{figure}
-The x-axis are the values in the fuzzy set (number of packets), and the y-axis is the membership function, which determines the region.
-For example, if the number of packets with the same destination address is 15 then this will be considered as low for a degree of 0.4, but will be considered medium or high for other degrees.
-If the number of packets is determined to be high, then the connection will be aborted.~\cite{DBLP:journals/corr/abs-1204-6416}
-
-\subsection{Genetic Algorithm}
-Genetic algorithms are a search technique used to find solutions to problems \emph{Mutation}, \emph{selection}, and \emph{crossover} are used to evolve and improve rules.
-
-Mutation is where random bits in a solution are changed.
-Selection is where solutions that have a better \emph{fitness} are chosen over the other solutions.
-The fitness function determines the quality of a particular solution.
-Crossover is where two solutions swap one of their characteristics with the other to form two new solutions.
-
-The solution to a problem is represented as a chromosome.
-First, a randomly generated population of chromosomes is created.
-According to the characteristics of the problem, the positions of each chromosome are encoded as bits, characters, or numbers.
-Then mutation, selection, and crossover are applied to each generation and eventually the best solution is found.~\cite{DBLP:journals/corr/abs-1204-6416}
-
-\subsection{Data Sets and Features}
+\subsection{Data Sets}
+\label{sec:dataSets}
 Two different data sets are used in this paper: KDD99 and RLD09.
-
 The KDD99 data set is a benchmark data set that was generated by simulating a military network environment in 1999~\cite{6559603}.
@@
 To create the data set, the authors captured network data from the Computer Engineering Department at King Mongkut's University of Technology Thonburi, in Bangkok, Thailand.
-The data has around ten million preprocessed data packets.
-It has 17 different types of attacks that can be divided into denial of service attacks and probe attacks.
+The data has around ten million preprocessed data \emph{packets}.
+A packet is a unit of data that is carried throughout a network~\cite{networkPacket}.
+RLD09 has 17 different types of attacks that can be divided into denial of service attacks and probe attacks.
 It also has normal network activity.
-A packet sniffer was used to get information about TCP, UDP, and ICMP headers from protocol packets.
-Then this information was processed into 12 features:
-\begin{itemize}
-  \item Features of TCP
+RLD09 uses 12 features, which include the number of packets, source ports, and destination ports.
 
-    \# of packets
+\subsection{Rules}
+\label{sec:rules}
+Rules are a way in which elements of one set are separated into different sets, or classes, in order to differentiate between normal connections and attacks.
+Rules are represented by if-then statements in the following format: If <\emph{condition}> then <\emph{action}>.~\cite{bapuji2012soft} Rules can specify the details of a packet such as the IP address, port number and protocal.
+If a packet matches any of the rules in the intrusion detection system then the system will take appropriate action, which may include stopping the connection or logging off the system.~\cite{DBLP:journals/corr/abs-1204-6416}
 
-    \# of source ports
+\subsection{Fuzzy Logic}
+\label{sec:fuzzyLogic}
+Attacks on systems do not always have a fixed patter, so fuzzy logic is used to detect patterns that have a behavior that is between normal and unusual.
+A fuzzy logic rule has the following format: If <\emph{condition}> then <\emph{consequence}>, where \emph{condition} is a fuzzy variable, and \emph{consequence} is a fuzzy set.
+An example of this is: if the number of packets with the same destination address is high, then the pattern is unusual.
+To determine what is considered high, the values of the packets are divided into fuzzy sets.
+In Figure~\ref{fig:triangleFigure} there are three sets: low (region A), medium (region B), and high (region C).
+\begin{figure}
+\centering
+\psfig{file=triangleFigure.JPG}
+\caption{Fuzzy shape with three sets: low (A), medium (B), and high (C)~\cite{DBLP:journals/corr/abs-1204-6416}}
+\label{fig:triangleFigure}
+\end{figure}
+The x-axis are the values in the fuzzy set (number of packets), and the y-axis is the membership function, which determines the region.
+For example, if the number of packets with the same destination address is 15 then this will be considered as low for a degree of 0.4, but will be considered medium or high for other degrees.
+If the number of packets is determined to be high, then the connection will be aborted.~\cite{DBLP:journals/corr/abs-1204-6416}
 
-    \# of destination ports
+\subsection{Genetic Algorithm}
+\label{sec:GA}
+Genetic algorithms are a search technique used to find solutions to problems \emph{Mutation}, \emph{selection}, and \emph{crossover} are used to evolve and improve rules.
 
-    \# of fin flags, syn flags, push flags, ack flags, urgent flags
-  \item Features of UDP
+Mutation is where random bits in a solution are changed.
+Selection is where solutions that have a better \emph{fitness} are chosen over the other solutions.
+The fitness function determines the quality of a particular solution.
+Crossover is where two solutions swap one of their characteristics with the other to form two new solutions.
 
-    \# of packets
-
-    \# of source ports
-
-    \# of destination ports
-  \item Features of ICMP
-
-    \# of ICMP packets
-\end{itemize}
+The solution to a problem is represented as a chromosome.
+First, a randomly generated population of chromosomes is created.
+According to the characteristics of the problem, the positions of each chromosome are encoded as bits, characters, or numbers.
+Then mutation, selection, and crossover are applied to each generation and eventually the best solution is found.~\cite{DBLP:journals/corr/abs-1204-6416}
 
@@
 \label{sec:genAlgImp}
-\mycomment{Figure out better title for Section~\ref{sec:genAlgImp}}
 
@@
 
-In the detection phase, a population is created for test data and then the type of the data is predicted.
+In the detection phase, a population is created for test data.
+Selection, crossover, and mutation occur, and then the type of the data (whether it is an attack or normal behavior) is predicted.
 The set of chromosomes that was created in the precalculation phase is used in the detection phase to find the fitness of each chromosome in the population.
@@
 They used only the numerical features of the KDD99 data set (34 out of 41 total features).
-For the training data set, the KDD99 10\% version file was used, and for the test data set, the KDD99 corrected version file was used.
+For the training data set, the KDD99 10\% version file was used, and for the test data set, the KDD99 corrected version file~\cite{KDD99} was used.
+The training set has a total of 494,021 records, and 396,741 of them are attacks.
+The test set has a total of 311,029 records, and 250,436 of them are attacks.
 Table~\ref{tab:numberOfRecords} shows the number of records of normal and attack activity in the training and test sets.
@@
 \end{table}
-
-In the detection phase, an initial population is created and then is compared with each chromosome that was created in the training phase.
-Crossover and mutation happen in the population in order to create a new population.
-The results from running the genetic algorithm are shown in Tables~\ref{tab:genAlgResults} and~\ref{tab:genAlgResults2}.
-The detection rate was 94.9\%.
-The false positive rate was 30.5\%.
-The false negative rate was 5\%.
-The true negative rate was 69.5\%.
 
@@
 
+In the detection phase, an initial population is created and then is compared with each chromosome that was created in the precalculation phase.
+Crossover and mutation happen in the population in order to create a new population.
+The results from running the genetic algorithm are shown in Tables~\ref{tab:genAlgResults} and~\ref{tab:genAlgResults2}.
+The detection rate was 94.9\%.
+The false positive rate was 30.5\%.
+The false negative rate was 5\%.
+The true negative rate was 69.5\%.
+
 The authors of~\cite{DBLP:journals/corr/abs-1204-1336} compared their results with the winning entry of the KDD99 Classifier Learning Contest~\cite{KDD99Contest}, and found that they had a better detection rate for denial of service and user to root attacks than the winning entry.
@@
 \label{sec:fuzGenAlgImp}
-\mycomment{Figure out better title for Section~\ref{sec:fuzGenAlgImp}}
-
 The focus of~\cite{6496342, 6559603} is on detecting new or unknown types of attacks in a network.
@@
 
-\mycomment{Figure out how to float algorithms.}
 \begin{algorithm}
@@
 \caption{Unknown Attack Experiment}
-\begin{tabular}{|llcc|} \hline
+\begin{tabular}{llll}
 Test & Unknown & Decision & Fuzzy\\
@@
 
-1 & Advance Port Scan & Avg = & Avg =\\
-  & Ack Scan		  & 98.33 & 100\\
-  & Xmas Tree		  &		  &\\ \hline
+1 & Adv Port Scan (Probe) & Avg = & Avg =\\
+  & Ack Scan (Probe)		  & 98.33 & 100\\
+  & Xmas Tree (Probe)		  &		  &\\ \hline
 
-2 & UDP Flood & Avg = & Avg =\\
-  & Host Scan & 46.65 & 99.80\\
-  & UDP Scan  &       &\\ \hline
+2 & UDP Flood (DoS) & Avg = & Avg =\\
+  & Host Scan (Probe) & 46.65 & 99.80\\
+  & UDP Scan (Probe)  &       &\\ \hline
 
-3 & Jping    & Avg =          & Avg =\\
-  & Syn Scan & 99.70 & 98.75\\
-  & Fin Scan &                &\\ \hline
+3 & Jping (DoS)    & Avg =          & Avg =\\
+  & Syn Scan (Probe) & 99.70 & 98.75\\
+  & Fin Scan (Probe) &                &\\ \hline
 
-4 & UDP Flood & Avg = & Avg =\\
-  & RCP Scan  & 70.35 & 98.15\\
-  & Fin Scan  &       &\\ \hline
+4 & UDP Flood (DoS) & Avg = & Avg =\\
+  & RCP Scan (Probe)  & 70.35 & 98.15\\
+  & Fin Scan (Probe)  &       &\\ \hline
 
-5 & Http Flood & Avg =          & Avg =\\
-  & RCP Scan  & 99.94 & 97.50\\
-  & Fin Scan  &                &\\
+5 & Http Flood (DoS) & Avg =          & Avg =\\
+  & RCP Scan (Probe)  & 99.94 & 97.50\\
+  & Fin Scan (Probe) &                &\\
 \hline\end{tabular}
@@
 \caption{KDD99 and RLD09 Results}
-\begin{tabular}{|cccccc|} \hline
+\begin{tabular}{cccccc}
 Data set & Attack & Normal & FP (\%) & FN (\%) & DR (\%)\\ \hline
@@
 RLD09 & 10,500 & 16,000 & 1.14 & 3.39 & 97.97\\
-\hline\end{tabular}
+\end{tabular}
 \label{tab:bothSetsResults}
@@
 \caption{Results for KDD99 with Certain Attacks}
-\begin{tabular}{|cccccc|} \hline
+\begin{tabular}{cccccc}
 Test & Attack & Type & FP (\%) & FN (\%) & DR (\%)\\ \hline
@@
 4 & Satan & Probe & 0.74 & 3.75 & 99.22\\
-\hline\end{tabular}
+\end{tabular}
 \label{tab:kddAttacks}
@@
 \caption{Results for RLD99 with Certain Attacks}
-\begin{tabular}{|cccccc|} \hline
+\begin{tabular}{cccccc}
 Test & Attack & Type & FP (\%) & FN (\%) & DR (\%)\\ \hline
@@
 4 & Synscan & Probe & 0.65 & 4.2 & 99.24\\
-\hline\end{tabular}
+\end{tabular}
 \label{tab:rldAttacks}
@@
 \section{Conclusions}
+\label{sec:conclusion}
+This paper showed that the use of genetic algorithms and fuzzy logic in intrusion detection are effective ways of detecting attacks.
+The genetic algorithm that was used in~\cite{DBLP:journals/corr/abs-1204-1336} had a high detection rate for denial of service attacks.
+When compared with the winning entry of the KDD99 Classifier Learning Contest, it was shown to have a better detection rate for both denial of service and user to root attacks.
+The fuzzy genetic algorithm that was used in~\cite{6496342, 6559603} had a higher detection rate than a decision tree algorithm in most cases.
+It was also shown that fuzzy genetic algorithms are good at detecting unknown attacks.
 
-\section{Acknowledgments}
+%\section{Acknowledgments}
 

### 2013-10-30 paper.tex
Wrote more in the introduction.

@@
 
-In Section~\ref{sec:background} I will give some background on intrusion detection systems.
-I will explain the types of intrusion detection systems, the main types of networking attacks, and ways of detecting attacks.
+In Section~\ref{sec:background} I will give some background on intrusion detection.
+I will explain the main types of networking attacks in Section~\ref{sec:typesAttacks}, and ways of detecting attacks in Section~\ref{sec:detectionMeth}.
+In Section~\ref{sec:dataSets} I will explain two data sets that are used in intrusion detection.
 I will explain what \emph{rules} are in Section~\ref{sec:rules}, and then give information on fuzzy logic and genetic algorithms in Sections~\ref{sec:fuzzyLogic} and~\ref{sec:GA}.
-I will then explain two data sets that are used in intrusion detection.
-After that, I will explain the measures that are used in determining the accuracy of an intrusion detection system.
+After that, I will explain the measures that are used in determining the accuracy of an intrusion detection system in Section~\ref{sec:PosNeg}.
+In Section~\ref{sec:genAlgImp} I will explain the use of a genetic algorithm in an intrusion detection system, and in Section~\ref{sec:fuzGenAlgImp} I will explain the use of a fuzzy genetic algorithm.
+Section~\ref{sec:conclusion} has some conclusions about the use of genetic algorithms and fuzzy genetic algorithms in intrusion detection systems.
 
@@
 \label{sec:fuzzyLogic}
-Attacks on systems do not always have a fixed patter, so fuzzy logic is used to detect patterns that have a behavior that is between normal and unusual.
+Attacks on systems do not always have a fixed pattern, so fuzzy logic is used to detect patterns that have a behavior that is between normal and unusual.
 A fuzzy logic rule has the following format: If <\emph{condition}> then <\emph{consequence}>, where \emph{condition} is a fuzzy variable, and \emph{consequence} is a fuzzy set.
@@
 
+\vspace{0.30cm}
+
 \subsection{Genetic Algorithm}
 \label{sec:GA}
-Genetic algorithms are a search technique used to find solutions to problems \emph{Mutation}, \emph{selection}, and \emph{crossover} are used to evolve and improve rules.
+Genetic algorithms are a search technique used to find solutions to problems. \emph{Mutation}, \emph{selection}, and \emph{crossover} are used to evolve and improve rules.
 
-Mutation is where random bits in a solution are changed.
-Selection is where solutions that have a better \emph{fitness} are chosen over the other solutions.
-The fitness function determines the quality of a particular solution.
-Crossover is where two solutions swap one of their characteristics with the other to form two new solutions.
+Mutation is where random bits in an \emph{individual}, or possible solution, are changed.
+Selection is where individuals that have a better \emph{fitness} are chosen over the other individuals.
+The fitness function determines the quality of a particular individual.
+Crossover is where two individuals swap one of their characteristics with the other to form two new individuals.
 
@@
 \caption{Number of records}
+\vspace{0.20cm}
 \begin{tabular}{lll}
@@
 \caption{Results for Genetic Algorithm Experiment}
+\vspace{0.20cm}
 \begin{tabular}{cccccccc}
@@
 \caption{Results for Genetic Algorithm Experiment}
+\vspace{0.20cm}
 \begin{tabular}{cccc}
@@
 \caption{Fuzzy encoding for a feature}
+\vspace{0.20cm}
 \begin{tabular}{|cccc|} \hline
@@
 \caption{A rule with 12 blocks of features}
+\vspace{0.20cm}
 \begin{tabular}{|cccc|c|cccc|c|} \hline
@@
 \caption{Results from Experiment 1}
+\vspace{0.20cm}
 \begin{tabular}{ccccccc}
@@
 Table~\ref{tab:fuzGenExp2} shows some of the results from the fuzzy genetic algorithm and a decision tree algorithm, which is another common way of addressing these problems.
-For further information, see~\cite{decisionTree}.
+For further information on decision trees, see~\cite{decisionTree}.
 When compared with the decision tree algorithm, the fuzzy genetic algorithm has a higher detection rate in all cases except 3 and 5.
@@
 \caption{Unknown Attack Experiment}
+\vspace{0.20cm}
 \begin{tabular}{llll}
@@
 \caption{KDD99 and RLD09 Results}
+\vspace{0.20cm}
 \begin{tabular}{cccccc}
@@
 \caption{Results for KDD99 with Certain Attacks}
+\vspace{0.20cm}
 \begin{tabular}{cccccc}
@@
 \caption{Results for RLD99 with Certain Attacks}
+\vspace{0.20cm}
 \begin{tabular}{cccccc}

### 2013-11-05 Talk/Presentation.tex
Adding the presentation slides.

\documentclass{beamer}

\mode<presentation>
{
  \usetheme{CambridgeUS}
  \setbeamercovered{transparent}
}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage[T1]{fontenc} 
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.
\usepackage{amsmath}

\newcommand{\linespace}{\vskip 0.25cm}

\definecolor{MyForestGreen}{rgb}{0,0.7,0} 
\newcommand{\tableemph}[1]{{#1}}
\newcommand{\tablewin}[1]{\tableemph{#1}}
\newcommand{\tablemid}[1]{\tableemph{#1}}
\newcommand{\tablelose}[1]{\tableemph{#1}}

\definecolor{MyLightGray}{rgb}{0.6,0.6,0.6}
\newcommand{\tabletie}[1]{\color{MyLightGray} {#1}}

% The text in square brackets is the short version of your title and will be used in the
% header/footer depending on your theme.
\title[Intrusion Detection]{Intrusion Detection with \\ Genetic Algorithms and Fuzzy Logic}

% Sub-titles are optional - uncomment and edit the next line if you want one.
% \subtitle{Why does sub-tree crossover work?}

% The text in square brackets is the short version of your name(s) and will be used in the
% header/footer depending on your theme.
\author[Ireland]{Emma Ireland}

% The text in square brackets is the short version of your institution and will be used in the
% header/footer depending on your theme.
\institute[U of Minn, Morris]
{
  Division of Science and Mathematics \\
  University of Minnesota, Morris \\
  Morris, Minnesota, USA
}

% The text in square brackets is the short version of the date if you need that.
\date[December '13] % (optional)
{December 2013 \\ UMM CSci Senior Seminar Conference}

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
\AtBeginSection[]
{
  \begin{frame}<beamer>
    \frametitle{Outline}
    \tableofcontents[currentsection, hideothersubsections]
  \end{frame}
}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

% For a 20-25 minute senior seminar talk you probably want something like:
% - Two or three major sections (other than the summary).
% - At *most* three subsections per section.
% - Talk about 30s to 2min per frame. So there should probably be between
%   15 and 30 frames, all told.

\section*{Overview}

\subsection*{The big picture}

\begin{frame}
  \frametitle{The big picture}

  \begin{columns}
  \begin{column}{0.6\textwidth}
  \begin{itemize}
  	\item 
	\item 
	\item 
	\item 
	\item 
  \end{itemize}
  \end{column}
  \begin{column}{0.4\textwidth}

  \end{column}
  \end{columns}
\end{frame}

\subsection*{Outline}

\begin{frame}
  \frametitle{Outline}
  \tableofcontents[hideallsubsections]
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Background]{Background}
\subsection{Types of Networking Attacks}
\begin{frame}
  \frametitle{title here}

\end{frame}

\subsection{Detection Methodologies}
\begin{frame}
  \frametitle{title}

\end{frame}

\subsection{Data Sets}
\begin{frame}
  \frametitle{title}

\end{frame}

\subsection{Rules}
\begin{frame}
  \frametitle{title}

\end{frame}

\subsection{Fuzzy Logic}
\begin{frame}
  \frametitle{title}

\end{frame}

\subsection{Genetic Algorithm}
\begin{frame}
  \frametitle{title}

\end{frame}

\subsection{Determining the accuracy of an algorithm}
\begin{frame}
  \frametitle{title}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Genetic Algorithm Implementation]{Genetic Algorithm Implementation}
\subsection{Algorithm Overview}
\begin{frame}
  \frametitle{title here}

\end{frame}

\subsection{Experimental Design and Results}
\begin{frame}
  \frametitle{title}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Fuzzy Genetic Algorithm Implementation]{Fuzzy Genetic Algorithm Implementation}

\subsection{Fuzzy Algorithm}

\begin{frame}
  \frametitle{Measuring the probability of a record being an attack}

\end{frame}

\begin{frame}
	\frametitle{Encoding of features and rules}

\end{frame}

\subsection{Algorithm Overview}

\begin{frame}
	\frametitle{title}

\end{frame}

\subsection{Experimental Design and Results}

\begin{frame}
	\frametitle{title}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Conclusions]{Conclusions}

\begin{frame}
\frametitle{Conclusions}

\begin{itemize}
  \item

  \linespace

  \item

  \linespace

  \item
\end{itemize}

\end{frame}

\begin{frame}
	\frametitle{Thanks!}

	Thank you for your time and attention!

	\linespace
	\linespace

	\begin{center}
	{\huge Questions?}
	\end{center}
\end{frame}

\section*{References}

\begin{frame} 
	\frametitle{References}

	\begin{thebibliography}{lskdjf}

  	\end{thebibliography}

\end{frame}

\end{document}

### 2013-11-05 Talk/Presentation.tex
Added trapezoidal shape, fuzzy algorithm, and things about features and rules to my presentation.
Added a few slides on KDD99.

@@
 \usepackage{amsmath}
+\usepackage{algorithmic}
 
@@
 
-\subsection{Data Sets}
+\subsection{Data Sets - KDD99 and RLD09}
 \begin{frame}
-  \frametitle{title}
+  \frametitle{KDD99}
+	\begin{itemize}
+		\item Generated by simulating a military network environment in 1999.
+		\item Has long been a standard data set for intrusion detection.
+		\item Data in the set is classified as normal or attack activity.
+
+		\linespace
+		\linespace
+
+		\item KDD99 uses 41 features.
+		\begin{itemize}
+		  	\item \emph{Features} are properties of a \emph{record}, (either an attack or normal activity), that are used to describe the activity.
+		\end{itemize}
+
+	\end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Some features of KDD99}
+	\begin{enumerate}
+		\item duration: length of the normal or attack activity
+in seconds.
+		\item src\_bytes: number of bytes sent from source to destination.
+		\item num\_failed\_logins: number of failed login attempts.
+		\item root\_shell: returns 1 if root shell is obtained, else returns 0.
+		\item num\_access\_files: number of operations on access control files.
+		\item srv\_count: number of connections to the same service as the current connection in the past two seconds.
+		\item serror\_rate: percentage of connections that have ``SYN" errors.
+		\item same\_srv\_rate: percentage of connections to the same service.
+	\end{enumerate}
+\end{frame}
+
+\begin{frame}
+  \frametitle{RLD09}
 
@@
 
+  \begin{columns}
+  \begin{column}{0.4\textwidth}
+  \begin{itemize}
+  	\item Trapezoidal shape
+  \end{itemize}
+
+  \includegraphics[width=0.95\textwidth]{../trapFigure.jpg}
+
+  \begin{itemize}
+	\item The parameters are the values of a feature.
+  \end{itemize}
+
+  \end{column}
+
+  \begin{column}{0.6\textwidth}
+  \begin{itemize}
+  	\item Fuzzy algorithm
+  \end{itemize}
+
+\begin{algorithmic}
+\IF{data value is between $b$ and $c$}
+  \STATE{prob = 1.0}
+\ELSIF{data value is between $a$ and $b$}
+  \STATE{
+    prob = $(\textrm{data}-a)/(b-a)$
+  }
+\ELSIF{data value is between $c$ and $d$}
+  \STATE{
+    prob = $(d-\textrm{data})/(d-c)$
+  }
+\ELSE \STATE{prob = 0.0}
+\ENDIF
+\end{algorithmic}
+
+  \end{column}
+  \end{columns}
 \end{frame}
@@
 	\frametitle{Encoding of features and rules}
+	\begin{itemize}
+	\item The four parameters are encoded into blocks.
+	\item Each block is a feature with values between 0.0 and 7.0.
+	\end{itemize}
+
+\begin{figure}
+\begin{tabular}{|cccc|} \hline
+010 & 011 & 100 & 101\\
+a=2 & b=3 & c=4 & d=5\\
+\hline\end{tabular}
+\end{figure}
+
+	\begin{itemize}
+	\item A rule has 12 blocks of features, at the end is the type of attack.
+	\end{itemize}
+
+\begin{figure}
+\begin{tabular}{|cccc|c|cccc|c|} \hline
+010 & 011 & 100 & 101   & ...... & 010 & 011 & 101 & 111   & DoS\\
+a=2 & b=3 & c=4 & d=5   & ...... & a=2 & b=3 & c=5 & d=7   &\\ 
+    &     & Block 1&    &        &     & Block 12& &       & Type\\
+\hline\end{tabular}
+\end{figure}
 

### 2013-11-06 Talk/Presentation.tex
Added titles to each slide, also wrote things on the RLD09 slide and the rules slide, and started the fuzzy logic slide.

@@
 
-\subsection*{The big picture}
+\subsection*{The Big Picture}
 
 \begin{frame}
-  \frametitle{The big picture}
+  \frametitle{The Big Picture}
 
@@
 \begin{frame}
-  \frametitle{title here}
-
+  \frametitle{Types of Networking Attacks}
+Explain DoS, remote to user, user to root, probe
 \end{frame}
@@
 \begin{frame}
-  \frametitle{title}
-
+  \frametitle{Detection Methodologies}
+Explain signature-based and anomaly-based detection
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Some features of KDD99}
+  \frametitle{Some Features of KDD99}
 	\begin{enumerate}
@@
   \frametitle{RLD09}
-
+	\begin{itemize}
+		\item RLD09 was created because KDD99 is 14 years old.
+		\item Data was captured from a university in Bangkok, Thailand.
+		\item The data has 10 million data packets.
+		\item 17 different types of attacks - divided into denial of service attacks, probe attacks. It also has normal activity.
+		\item 12 features, which include the number of packets, source ports, and destination ports.
+	\end{itemize}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{title}
+  \frametitle{Rules}
+	\begin{itemize}
+		\item Elements of one set are separated into different sets in order to differentiate between normal connections and attacks.
 
+		\linespace
+		\linespace
+
+		\item If <\emph{condition}> then <\emph{action}>.
+
+		\linespace
+		\linespace
+
+		\item Specify the details of a packet such as the IP address or port number.
+		\item If a packet matches any of the rules in the intrusion detection system, the system will take appropriate action, which may include stopping the connection or logging off the system.
+
+	\end{itemize}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{title}
+  \frametitle{Fuzzy Logic}
+  \begin{columns}
+  \begin{column}{0.6\textwidth}
+  \begin{itemize}
+  	\item Used to detect patterns that have a behavior that is between normal and unusual.
+  	\item Fuzzy logic rule: If <\emph{condition}> then <\emph{consequence}>.
+  	\begin{itemize}
+  		\item \emph{condition} is a fuzzy variable
+  		\item \emph{consequence} is a fuzzy set
+  	\end{itemize}
+  \end{itemize}  
+  \end{column}
+
+  \begin{column}{0.4\textwidth}
+  \includegraphics[width=0.95\textwidth]{../triangleFigure.JPG}
+  \end{column}
+  \end{columns}
+\end{frame}
+
+\subsection{Genetic Algorithms}
+\begin{frame}
+  \frametitle{Genetic Algorithms}
 
@@
 
-\subsection{Genetic Algorithm}
+\subsection{Determining the Accuracy of an Algorithm}
 \begin{frame}
-  \frametitle{title}
-
-\end{frame}
-
-\subsection{Determining the accuracy of an algorithm}
-\begin{frame}
-  \frametitle{title}
-
+  \frametitle{Determining the Accuracy of an Algorithm}
+Explain training and test set, false positive, false negative, true positive, true negative, detection rate.
 \end{frame}
@@
 \begin{frame}
-  \frametitle{title here}
+  \frametitle{Algorithm Overview}
 
@@
 \begin{frame}
-  \frametitle{title}
+  \frametitle{Experimental Design}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Results}
 
@@
 
+\subsection{Main Points of Research}
+\begin{frame}
+  \frametitle{Main Points of Research}
+	\begin{itemize}
+		\item Detecting new or unknown types of attacks in a network.
+		\item The intrusion detection system used is able to identify normal network activity as well as attacks using a fuzzy genetic algorithm.
+
+		\linespace
+		\linespace
+
+		\item Ran experiments using only RLD09, and experiments using KDD99 and RLD09 together.
+	\end{itemize}
+\end{frame}
+
 \subsection{Fuzzy Algorithm}
@@
 \begin{frame}
-  \frametitle{Measuring the probability of a record being an attack}
+  \frametitle{Measuring the Probability of a Record Being an Attack}
 
@@
 \begin{frame}
-	\frametitle{Encoding of features and rules}
+	\frametitle{Encoding of Features and Rules}
 	\begin{itemize}
@@
 \begin{frame}
-	\frametitle{title}
+	\frametitle{Algorithm Overview}
 
@@
 \begin{frame}
-	\frametitle{title}
+	\frametitle{Experiments Using Only RLD09}
+Experiment 1
+
+\end{frame}
+
+\begin{frame}
+	\frametitle{Experiments Using Only RLD09}
+Experiment 2
+
+\end{frame}
+
+\begin{frame}
+	\frametitle{Experiments Using Both RLD09 and KDD99}
+Experiment 1
+
+\end{frame}
+
+\begin{frame}
+	\frametitle{Experiments Using Both RLD09 and KDD99}
+Experiment 2
+
+\end{frame}
+
+\begin{frame}
+	\frametitle{Experiments Using Both RLD09 and KDD99}
+Experiment 3
 

### 2013-11-07 Talk/Presentation.tex
Added more to the fuzzy logic slide.

@@
   \frametitle{Fuzzy Logic}
-  \begin{columns}
-  \begin{column}{0.6\textwidth}
+
   \begin{itemize}
   	\item Used to detect patterns that have a behavior that is between normal and unusual.
-  	\item Fuzzy logic rule: If <\emph{condition}> then <\emph{consequence}>.
+  	\item If <\emph{condition}> then <\emph{consequence}>.
   	\begin{itemize}
-  		\item \emph{condition} is a fuzzy variable
-  		\item \emph{consequence} is a fuzzy set
+  		\item \emph{condition} is a fuzzy variable and \emph{consequence} is a fuzzy set
   	\end{itemize}
+  	\item If the number of packets with the same destination address is 20, and $a$=10, $b$=25, $c$=45, then the degree=.6 and the region=B so the number of packets=medium.
   \end{itemize}
+
+  \begin{columns}
+  \begin{column}{0.4\textwidth}
+      \includegraphics[width=0.95\textwidth]{../triangleFigure.JPG} 
   \end{column}
 
-  \begin{column}{0.4\textwidth}
-  \includegraphics[width=0.95\textwidth]{../triangleFigure.JPG}
+  \begin{column}{0.6\textwidth}
+  \begin{algorithmic}
+	\IF{x is between $a$ and $b$}
+		\STATE{degree = $(\textrm{x}-a)/(b-a)$}
+	\ELSIF{x is between $b$ and $c$}
+		\STATE{degree = $(c-\textrm{x})/(c-b)$}
+	\ELSE \STATE{degree = 0.0}
+	\ENDIF
+  \end{algorithmic}
   \end{column}
@@
   \begin{column}{0.6\textwidth}
-  \begin{itemize}
-  	\item Fuzzy algorithm
-  \end{itemize}
+%  \begin{itemize}
+%  	\item Fuzzy algorithm
+%  \end{itemize}
 

### 2013-11-09 Talk/Presentation.tex
Worked on the fuzzy algorithm slide
Added a different trapezoid figure, also added stuff to the fuzzy genetic algorithm slide.

--- 2013-11-06 Talk/Presentation.tex
@@
 		\item Data in the set is classified as normal or attack activity.
-
-		\linespace
-		\linespace
-
 		\item KDD99 uses 41 features.
@@
   \frametitle{Some Features of KDD99}
-	\begin{enumerate}
+	\begin{itemize}
 		\item duration: length of the normal or attack activity
 in seconds.
-		\item src\_bytes: number of bytes sent from source to destination.
+		%\item src\_bytes: number of bytes sent from source to destination.
 		\item num\_failed\_logins: number of failed login attempts.
 		\item root\_shell: returns 1 if root shell is obtained, else returns 0.
-		\item num\_access\_files: number of operations on access control files.
-		\item srv\_count: number of connections to the same service as the current connection in the past two seconds.
+		%\item num\_access\_files: number of operations on access control files.
+		%\item srv\_count: number of connections to the same service as the current connection in the past two seconds.
 		\item serror\_rate: percentage of connections that have ``SYN" errors.
-		\item same\_srv\_rate: percentage of connections to the same service.
-	\end{enumerate}
+		%\item same\_srv\_rate: percentage of connections to the same service.
+	\end{itemize}
 \end{frame}
@@
 		\item Data was captured from a university in Bangkok, Thailand.
-		\item The data has 10 million data packets.
-		\item 17 different types of attacks - divided into denial of service attacks, probe attacks. It also has normal activity.
-		\item 12 features, which include the number of packets, source ports, and destination ports.
+		\item The data has 10 million data packets, 17 different types of attacks (divided into denial of service and probe attacks), and 12 features.
 	\end{itemize}
@@
 		\item Elements of one set are separated into different sets in order to differentiate between normal connections and attacks.
-
-		\linespace
-		\linespace
-
-		\item If <\emph{condition}> then <\emph{action}>.
-
-		\linespace
-		\linespace
-
-		\item Specify the details of a packet such as the IP address or port number.
-		\item If a packet matches any of the rules in the intrusion detection system, the system will take appropriate action, which may include stopping the connection or logging off the system.
-
-	\end{itemize}
-\end{frame}
-
-\subsection{Fuzzy Logic}
-\begin{frame}
-  \frametitle{Fuzzy Logic}
-  \begin{columns}
-  \begin{column}{0.6\textwidth}
+		\item If-Then format
 		\begin{itemize}
-  	\item Used to detect patterns that have a behavior that is between normal and unusual.
-  	\item Fuzzy logic rule: If <\emph{condition}> then <\emph{consequence}>.
-  	\begin{itemize}
-  		\item \emph{condition} is a fuzzy variable
-  		\item \emph{consequence} is a fuzzy set
+			\item If the length of the activity is 4 seconds, then the probability of it being an attack is 100\%.
 		\end{itemize}				
 	\end{itemize}
-  \end{column}
-
-  \begin{column}{0.4\textwidth}
-  \includegraphics[width=0.95\textwidth]{../triangleFigure.JPG}
-  \end{column}
-  \end{columns}
 \end{frame}
@@
   \frametitle{Determining the Accuracy of an Algorithm}
-Explain training and test set, false positive, false negative, true positive, true negative, detection rate.
+Explain false positive, false negative, true positive, true negative, detection rate.
 \end{frame}
@@
 
-\subsection{Main Points of Research}
-\begin{frame}
-  \frametitle{Main Points of Research}
-	\begin{itemize}
-		\item Detecting new or unknown types of attacks in a network.
-		\item The intrusion detection system used is able to identify normal network activity as well as attacks using a fuzzy genetic algorithm.
-
-		\linespace
-		\linespace
-
-		\item Ran experiments using only RLD09, and experiments using KDD99 and RLD09 together.
-	\end{itemize}
-\end{frame}
-
 \subsection{Fuzzy Algorithm}
@@
   \begin{columns}
-  \begin{column}{0.4\textwidth}
-  \begin{itemize}
-  	\item Trapezoidal shape
-  \end{itemize}
-
-  \includegraphics[width=0.95\textwidth]{../trapFigure.jpg}
-
-  \begin{itemize}
-	\item The parameters are the values of a feature.
-  \end{itemize}
-
+  \begin{column}{0.6\textwidth}
+  \includegraphics[width=0.95\textwidth]{../TrapezoidFigure.pdf}
   \end{column}
 
-  \begin{column}{0.6\textwidth}
+  \begin{column}{0.4\textwidth}
+  Example:
   \begin{itemize}
-  	\item Fuzzy algorithm
+  	\item Feature: duration (length of the activity in seconds).
+  	\item a=1, b=3, c=5, d=7
+  	\item The length of the activity is 6 seconds (between c and d).
+  	\linespace
+  	\item prob = $\frac{d-\textrm{data}}{d-c}$ = $\frac{7-6}{7-5}$ = 0.5
   \end{itemize}
-
-\begin{algorithmic}
-\IF{data value is between $b$ and $c$}
-  \STATE{prob = 1.0}
-\ELSIF{data value is between $a$ and $b$}
-  \STATE{
-    prob = $(\textrm{data}-a)/(b-a)$
-  }
-\ELSIF{data value is between $c$ and $d$}
-  \STATE{
-    prob = $(d-\textrm{data})/(d-c)$
-  }
-\ELSE \STATE{prob = 0.0}
-\ENDIF
-\end{algorithmic}
 
@@
 	\frametitle{Algorithm Overview}
+\begin{columns}
+\begin{column}{0.6\textwidth}
+\begin{algorithmic}
+\FOR{each record}
+  \FOR{each rule}
+    \FOR{each feature}
+      \STATE{prob = fuzzy(); // Trapezoidal fuzzy rule shape}
+      \STATE{totalprob = totalprob + prob;}
+    \ENDFOR    
+    \IF{totalprob > threshold} 
+      \STATE{class is attack;}
+      %\ELSE \STATE{class is normal;}
+    \ENDIF
+  \ENDFOR
+  \STATE{find $A$, $B$, $\alpha$, and $\beta$  %// $A$: \# of attack records.
+$B$: \# of normal records.
+$\alpha$: \# of attack records correctly identified as attack.
+$\beta$: \# of normal records incorrectly classified as attack.}
 
+  } 
+\ENDFOR
+\STATE{calculate fitness}
+\STATE{crossover(), mutation()}
+
+\end{algorithmic}
+\end{column}
+
+\begin{column}{0.4\textwidth}
+	Fitness function:
+	\begin{equation*}
+	\frac{\alpha}{A} - \frac{\beta}{B}
+	\end{equation*}
+
+	$A$: \# of attack records.
+
+	$B$: \# of normal records.
+
+	$\alpha$: \# of attack records correctly identified as attack.
+
+	$\beta$: \# of normal records incorrectly classified as attack.
+
+\end{column}
+\end{columns}
 \end{frame}
@@
 \subsection{Experimental Design and Results}
+
+\begin{frame}
+	\frametitle{Experiments}
+A variety of experiments were run. Two experiments used just RLD09, and three experiments used KDD99 and RLD09 together.
+
+\end{frame}
 

### 2013-11-10 Talk/Presentation.tex
Added stuff about false positives/negatives and true positives/negatives, also added stuff to the experiments slide.

@@
   \frametitle{Determining the Accuracy of an Algorithm}
-Explain false positive, false negative, true positive, true negative, detection rate.
+	\begin{itemize}
+		\item False positive (FP): intrusion detection system incorrectly identifies normal activity as being an attack. 
+		\item False negative (FN): intrusion detection system fails to identify harmful activity. 
+		\item True positive (TP): intrusion detection system correctly identifies activities to be attacks. 
+		\item True negative (TN): intrusion detection system correctly
+identifies activities to be normal.
+		\item Detection rate (DR): the number of true positives divided by the total number of intrusions that happen.
+	\end{itemize}
 \end{frame}
@@
 	\frametitle{Experiments}
-A variety of experiments were run. Two experiments used just RLD09, and three experiments used KDD99 and RLD09 together.
-
+	\begin{itemize}
+		\item A variety of experiments were run. Two experiments used just RLD09, and three experiments used KDD99 and RLD09 together.
+		\item The experiments used a total of 16,000 records of normal activity and 10,500 records of attack activity.
+Of the attack records, 4,000 were denial of service attacks and 6,500 were probe attacks.
+	\end{itemize}
 \end{frame}
@@
 Experiment 1
-
 \end{frame}

### 2013-11-12 paper.tex
Made some changes based on feedback.

@@
 A way of detecting attacks is by using an intrusion detection system.
-An intrusion detection system is a device that monitors network activities for malicious or abnormal behaviors and then produces reports~\cite{DBLP:journals/corr/abs-1204-1336}.
-There are various ways of training the intrusion detection system about the possible threats.
-Two different ways, which I will talk about in this paper, are the use of genetic algorithms and fuzzy logic.
+An intrusion detection system is a device that monitors network activities for malicious or abnormal behaviors and then produces reports and alerts~\cite{DBLP:journals/corr/abs-1204-1336}.
+There are various ways of training the intrusion detection system about possible threats.
+Two approaches that I will talk about in this paper are the use of genetic algorithms and fuzzy logic.
 
@@
 I will explain the main types of networking attacks in Section~\ref{sec:typesAttacks}, and ways of detecting attacks in Section~\ref{sec:detectionMeth}.
-In Section~\ref{sec:dataSets} I will explain two data sets that are used in intrusion detection.
+In Section~\ref{sec:dataSets} I will describe two data sets that are used in intrusion detection.
 I will explain what \emph{rules} are in Section~\ref{sec:rules}, and then give information on fuzzy logic and genetic algorithms in Sections~\ref{sec:fuzzyLogic} and~\ref{sec:GA}.
@@
 For example, many systems lock out a user from an account after a certain number of failed login attempts.
-An attacker would be able to use this to prevent legitimate users from logging in.~\cite{DoSAttacks} Remote to user (R2L) attacks happen when an attacker sends packets to a machine over the network in order to gain access to things a local user would have on the machine.
+An attacker would be able to use this to prevent legitimate users from logging in~\cite{DoSAttacks}.
+Remote to user (R2L) attacks happen when an attacker sends packets to a machine over the network in order to gain access to things a local user would have on the machine.
 An example is when an attacker tries to gain access to a machine by guessing possible usernames and passwords.
@@
 Suppose the user's name has 35 characters.
-The last 15 characters will overflow the buffer, which will then overwrite the instructions that are to be executed next.
+The last 15 characters will overflow the buffer, which could then overwrite the instructions that are to be executed next.
 An attacker can cause commands to be executed by manipulating the data that overflows.
@@
 
-An anomaly is something that deviates from what is normal.
 Anomaly-based detection looks for patterns of activity that are rare and uncommon.
@@
 The data in the set is classified as normal or attack activity.
-KDD99 uses 41 \emph{features}, which are properties of a \emph{record} (either an attack or normal activity), that are used to describe the activity.
+KDD99 uses 41 \emph{features}, which are properties of a \emph{record} (either an attack or normal activity), that are used to describe the activity and help to distinguish normal connections from attacks.
 
-The authors of~\cite{6496342, 6559603} used 8 out of the 41 KDD99 features~\cite{KDD99} in their system: 
+In research conducted by P. Jongsuebsuk, N. Wattanapongsakorn, and
+C. Charnsripinyo~\cite{6496342, 6559603}, 8 out of the 41 KDD99 features~\cite{KDD99} were used in their system:
 \begin{enumerate}
@@
 
-The KDD99 data set is 14 years old, and newer attack types are not included in it because of its age.
-Because of this, the authors of~\cite{6496342, 6559603} created their own data set, RLD09, to use in their experiments.
+The KDD99 data set is 14 years old, and newer attack types are not included in it because of its age. Because of this, P. Jongsuebsuk, N. Wattanapongsakorn, and
+C.
+Charnsripinyo~\cite{6496342, 6559603} created their own data set, RLD09, to use in their experiments.
 To create the data set, the authors captured network data from the Computer Engineering Department at King Mongkut's University of Technology Thonburi, in Bangkok, Thailand.
@@
 A fuzzy logic rule has the following format: If <\emph{condition}> then <\emph{consequence}>, where \emph{condition} is a fuzzy variable, and \emph{consequence} is a fuzzy set.
-An example of this is: if the number of packets with the same destination address is high, then the pattern is unusual.
+For example: if the number of packets with the same destination address is high, then the pattern is unusual.
 To determine what is considered high, the values of the packets are divided into fuzzy sets.
@@
 If the number of packets is determined to be high, then the connection will be aborted.~\cite{DBLP:journals/corr/abs-1204-6416}
-
-\vspace{0.30cm}
 
@@
 \subsection{Algorithm Overview}
-The authors of~\cite{DBLP:journals/corr/abs-1204-1336} used a genetic algorithm to make their intrusion detection system.
+In research conducted by M.
+S.
+Hoque, M.
+A.
+Mukit, and M.
+A.
+N.
+Bikas~\cite{DBLP:journals/corr/abs-1204-1336}, a genetic algorithm was used to make their intrusion detection system.
 The system is divided into two phases: a precalculation phase and a detection phase.
@@
   \STATE{Apply crossover to the selected pair of chromosomes of the population}
-  \STATE{Apply mutation to each chromosome of the population}
+  \STATE{Apply mutation to each chromosome of the
+
+  population~//~mutation rate~=~0.35}
 \ENDWHILE
@@
 \STATE{crossover()}
-\STATE{mutation()}
+\STATE{mutation() // mutation rate~=~0.30}
 \end{algorithmic}
@@
 \end{equation*}
+$A$ is total number of attack records.
+$B$ is total number of normal records.
+$\alpha$ is total number of attack records correctly identified as attack.
+$\beta$ is total number of normal records incorrectly classified as attack.
+
 In this implementation, a population size of 10 was used for each generation.
@@
 
-The authors of~\cite{6496342, 6559603} first trained and tested the fuzzy genetic algorithm with the KDD99 data set.
+The authors of~\cite{6496342, 6559603} first trained and tested the fuzzy genetic algorithm (Algorithm~\ref{alg:fuzGenAlg}) with the KDD99 data set.
 There were 6 different types of denial of service attacks and 4 different types of probe attacks.

### 2013-11-14 paper.tex
Made some changes based on feedback from Alex.

@@
 
-\title{Intrusion Detection with Genetic Algorithms and Fuzzy Logic}
+\title{Intrusion Detection with \\ Genetic Algorithms and Fuzzy Logic}
 
@@
 This paper describes two different ways of training an intrusion detection system about possible attacks to a system: genetic algorithms and fuzzy logic.
-I will explain how genetic algorithms are used in intrusion detection systems and compare the results to the winning entry of the KDD99 Classifier Learning Contest.
-I will then explain how fuzzy genetic algorithms are used and compare those results with a decision tree algorithm.
+I will describe how genetic algorithms are used in intrusion detection systems and compare the results to the winning entry of the KDD99 Classifier Learning Contest.
+I will then describe how fuzzy genetic algorithms are used and compare those results with a decision tree algorithm.
 \end{abstract}
@@
 In Section~\ref{sec:background} I will give some background on intrusion detection.
-I will explain the main types of networking attacks in Section~\ref{sec:typesAttacks}, and ways of detecting attacks in Section~\ref{sec:detectionMeth}.
-In Section~\ref{sec:dataSets} I will describe two data sets that are used in intrusion detection.
-I will explain what \emph{rules} are in Section~\ref{sec:rules}, and then give information on fuzzy logic and genetic algorithms in Sections~\ref{sec:fuzzyLogic} and~\ref{sec:GA}.
-After that, I will explain the measures that are used in determining the accuracy of an intrusion detection system in Section~\ref{sec:PosNeg}.
-In Section~\ref{sec:genAlgImp} I will explain the use of a genetic algorithm in an intrusion detection system, and in Section~\ref{sec:fuzGenAlgImp} I will explain the use of a fuzzy genetic algorithm.
+I will describe the main types of networking attacks in Section~\ref{sec:typesAttacks}, and ways of detecting attacks in Section~\ref{sec:detectionMeth}.
+In Section~\ref{sec:dataSets} I will describe two data sets that are commonly used in intrusion detection.
+I will explain how \emph{rules} are used to differentiate between normal connections and attacks in Section~\ref{sec:rules}, and then give information on genetic algorithms in Section~\ref{sec:GA}.
+After that, I will describe the measures that are used in determining the accuracy of an intrusion detection system in Section~\ref{sec:PosNeg}.
+In Section~\ref{sec:genAlgImp} I will describe the use of a genetic algorithm in an intrusion detection system, and in Section~\ref{sec:fuzGenAlgImp} I will describe the use of a fuzzy genetic algorithm.
 Section~\ref{sec:conclusion} has some conclusions about the use of genetic algorithms and fuzzy genetic algorithms in intrusion detection systems.
@@
 \label{sec:dataSets}
-Two different data sets are used in this paper: KDD99 and RLD09.
-The KDD99 data set is a benchmark data set that was generated by simulating a military network environment in 1999~\cite{6559603}.
+Two different data sets are used in this paper to evaluate the performance of intrusion detection systems: KDD99 and RLD09.
+The KDD99 data set is a benchmark data set that was generated by simulating a military network environment in 1999.
 It has long been a standard data set for intrusion detection.
-The data in the set is classified as normal or attack activity.
-KDD99 uses 41 \emph{features}, which are properties of a \emph{record} (either an attack or normal activity), that are used to describe the activity and help to distinguish normal connections from attacks.
+The data was processed into five million records.
+A record is a sequence of TCP packets, between which data flows to and from a source IP address to a target IP address.
+The data in the set is classified as either normal or attack activity.
+KDD99 uses 41 \emph{features}, which are properties of a record that are used to describe the activity and help to distinguish normal connections from attacks.~\cite{KDD99}
 
-In research conducted by P. Jongsuebsuk, N. Wattanapongsakorn, and
-C. Charnsripinyo~\cite{6496342, 6559603}, 8 out of the 41 KDD99 features~\cite{KDD99} were used in their system:
+In research conducted by P. Jongsuebsuk,
+
+N. Wattanapongsakorn, and
+C.
+Charnsripinyo~\cite{6496342, 6559603}, which will be discussed in Section~\ref{sec:fuzGenAlgImp}, eight out of the 41 KDD99 features~\cite{KDD99} were used.
+They decided on these 8 features based on research done in~\cite{Fries:2008:FAN:1388969.1389037}.
+The features are:
 \begin{enumerate}
@@
 Charnsripinyo~\cite{6496342, 6559603} created their own data set, RLD09, to use in their experiments.
-To create the data set, the authors captured network data from the Computer Engineering Department at King Mongkut's University of Technology Thonburi, in Bangkok, Thailand.
-The data has around ten million preprocessed data \emph{packets}.
-A packet is a unit of data that is carried throughout a network~\cite{networkPacket}.
+To create the data set, the authors captured network data from a university in Bangkok, Thailand.
 RLD09 has 17 different types of attacks that can be divided into denial of service attacks and probe attacks.
@@
 RLD09 uses 12 features, which include the number of packets, source ports, and destination ports.
+For further information, see~\cite{6496342, 6559603}.
+RLD09 only has 12 features because the number of features affect the speed and memory consumption of the intrusion detection system.
 
@@
 \label{sec:rules}
-Rules are a way in which elements of one set are separated into different sets, or classes, in order to differentiate between normal connections and attacks.
-Rules are represented by if-then statements in the following format: If <\emph{condition}> then <\emph{action}>.~\cite{bapuji2012soft} Rules can specify the details of a packet such as the IP address, port number and protocal.
-If a packet matches any of the rules in the intrusion detection system then the system will take appropriate action, which may include stopping the connection or logging off the system.~\cite{DBLP:journals/corr/abs-1204-6416}
-
-\subsection{Fuzzy Logic}
-\label{sec:fuzzyLogic}
-Attacks on systems do not always have a fixed pattern, so fuzzy logic is used to detect patterns that have a behavior that is between normal and unusual.
-A fuzzy logic rule has the following format: If <\emph{condition}> then <\emph{consequence}>, where \emph{condition} is a fuzzy variable, and \emph{consequence} is a fuzzy set.
-For example: if the number of packets with the same destination address is high, then the pattern is unusual.
-To determine what is considered high, the values of the packets are divided into fuzzy sets.
-In Figure~\ref{fig:triangleFigure} there are three sets: low (region A), medium (region B), and high (region C).
-\begin{figure}
-\centering
-\psfig{file=triangleFigure.JPG}
-\caption{Fuzzy shape with three sets: low (A), medium (B), and high (C)~\cite{DBLP:journals/corr/abs-1204-6416}}
-\label{fig:triangleFigure}
-\end{figure}
-The x-axis are the values in the fuzzy set (number of packets), and the y-axis is the membership function, which determines the region.
-For example, if the number of packets with the same destination address is 15 then this will be considered as low for a degree of 0.4, but will be considered medium or high for other degrees.
-If the number of packets is determined to be high, then the connection will be aborted.~\cite{DBLP:journals/corr/abs-1204-6416}
+A commonly used approach for detecting intrusions and to differentiate between normal connections and attacks is to use rules.
+Rules are represented by if-then statements in the following format: If (\emph{condition}) then (\emph{consequence}).
+The condition part of the rule is composed of one or more features, and the consequence of the rule says if it is an intrusion or not.
+For example: if (\emph{duration} = 1, and \emph{src\_bytes} = 0, and \emph{serror\_rate} = 50) then \emph{intrusion}.~\cite{Fries:2008:FAN:1388969.1389037}
 
@@
 \label{sec:GA}
-Genetic algorithms are a search technique used to find solutions to problems. \emph{Mutation}, \emph{selection}, and \emph{crossover} are used to evolve and improve rules.
+Genetic algorithms are a search technique used to find solutions to problems.
+Operations analogous to biological \emph{mutation}, \emph{selection}, and \emph{crossover} are used to evolve and improve rules.
 
@@
 
-The \emph{detection rate} (DR) of an intrusion detection system is the number of true positives divided by the total number of intrusions that happen.~\cite{ids}
+The \emph{detection rate} (DR) of an intrusion detection system is the number of intrusions detected by the system divided by the total number of intrusions.~\cite{ids}
 
-\section{Genetic Algorithm Implementation}
+\section{Using Genetic Algorithms}
 \label{sec:genAlgImp}
@@
 Algorithm~\ref{alg:precalc} is used in the precalculation phase.
+It initializes chromosomes for comparison.
+It takes the network data as an input and outputs a set of chromosomes.
 
@@
 
-In the detection phase, a population is created for test data.
+In the detection phase, the precalculated set of chromosomes is used to find the fitness of each chromosome in the population.
 Selection, crossover, and mutation occur, and then the type of the data (whether it is an attack or normal behavior) is predicted.
-The set of chromosomes that was created in the precalculation phase is used in the detection phase to find the fitness of each chromosome in the population.
 Algorithm~\ref{alg:detection} is used in the detection phase.
+It takes the precalculated set of chromosomes as an input and outputs the type of data.
 
@@
   \STATE{Remove some chromosomes with worse fitness}
-  \STATE{Apply crossover to the selected pair of chromosomes of the population}
+  \STATE{Apply crossover to the selected pair of chromosomes of the population // crossover rate~=~0.15}
   \STATE{Apply mutation to each chromosome of the
@@
 They used only the numerical features of the KDD99 data set (34 out of 41 total features).
-For the training data set, the KDD99 10\% version file was used, and for the test data set, the KDD99 corrected version file~\cite{KDD99} was used.
-The training set has a total of 494,021 records, and 396,741 of them are attacks.
+For the training data set, the KDD99 10\% version file was used.
+This is a 10\% subset of the full KDD99 data set.
+For the test data set, the KDD99 corrected version file was used.
+This file corrects various typos that were in other files.~\cite{KDD99} The training set has a total of 494,021 records, and 396,741 of them are attacks.
 The test set has a total of 311,029 records, and 250,436 of them are attacks.
@@
 In the detection phase, an initial population is created and then is compared with each chromosome that was created in the precalculation phase.
-Crossover and mutation happen in the population in order to create a new population.
+Crossover and mutation happen in the detection population in order to create a new population.
 The results from running the genetic algorithm are shown in Tables~\ref{tab:genAlgResults} and~\ref{tab:genAlgResults2}.
@@
 
-The authors of~\cite{DBLP:journals/corr/abs-1204-1336} compared their results with the winning entry of the KDD99 Classifier Learning Contest~\cite{KDD99Contest}, and found that they had a better detection rate for denial of service and user to root attacks than the winning entry.
+The authors of~\cite{DBLP:journals/corr/abs-1204-1336} compared their results with the winning entry of the KDD99 Classifier Learning Contest~\cite{KDD99Contest}.
+The winning entry used a decision tree algorithm, which is another common algorithm that is used.
+For further information on decision trees, see~\cite{decisionTree}.
+The authors of~\cite{DBLP:journals/corr/abs-1204-1336} found that they had a better detection rate for denial of service and user to root attacks than the winning entry.
+For the winning entry, the detection rate of denial of service was 97.1\%, and for user to root it was 13.2\%.
+In~\cite{DBLP:journals/corr/abs-1204-1336} the detection rate of denial of service was 99.4\% and for user to root it was 18.9\%.
 
-\section{Fuzzy Genetic Algorithm Implementation}
+\section{Using Fuzzy Genetic Algorithms}
 \label{sec:fuzGenAlgImp}
@@
 
-\subsection{Fuzzy Algorithm}
-In order to measure the probability that a record is an attack, a trapezoidal shape was used in the algorithm.
+\subsection{Fuzzy Logic Rules and Algorithm}
+Attacks on systems do not always have a fixed pattern, so fuzzy logic is used to detect patterns that have a behavior that is between normal and unusual.
+Fuzzy logic rules are similar to the rules described in Section~\ref{sec:rules}, except that \emph{consequence} is a certainty factor.
+For example, if (\emph{duration} = 6) then (\emph{probability of it being an attack is 50\%}).
+
+In order to measure the probability that a record is an attack, a trapezoidal shape was used in~\cite{6496342, 6559603}.
 This is shown in Figure~\ref{fig:trapFigure}.
@@
 \centering
-\psfig{file=trapFigure.jpg}
-\caption{Trapezoidal shape with 4 parameters where $a\leq b\leq c\leq d$}
+\psfig{file=TrapezoidFigure.pdf, width=3.3in}
+\caption{Used to measure the probability that a record is an attack}
 \label{fig:trapFigure}
@@
 \end{algorithm}
+
+An example of using the fuzzy shape and algorithm is: suppose that the feature is duration, and suppose it is 6 seconds.
+Suppose that $a$ = 1, $b$ = 3, $c$ = 5, $d$ = 7.
+Then the probability of an attack is equal to
+
+\begin{equation*}
+\frac{d-\textrm{data}}{d-c} = \frac{7-6}{7-5} = 0.5.
+\end{equation*}
 
@@
 \STATE{preserve\_best()}
-\STATE{crossover()}
+\STATE{crossover() // single-point crossover}
 \STATE{mutation() // mutation rate~=~0.30}
@@
 
-In this implementation, a population size of 10 was used for each generation.
+A population size of 10 was used for each generation.
 An individual in the population represents a possible detection rule.
@@
 \subsection{Experimental Design and Results}
+A variety of experiments were run in~\cite{6496342, 6559603}. Two experiments used just RLD09, and three experiments used KDD99 and RLD09 together.
 
 \subsubsection{Experiments using Only RLD09}
-The experiments that the authors of~\cite{6496342, 6559603} performed used a total of 16,000 records of normal activity and 10,500 records of attack activity.
+The experiments using only RLD09 that the authors of~\cite{6496342, 6559603} performed used a total of 16,000 records of normal activity and 10,500 records of attack activity.
 Of the attack records, 4,000 of them were denial of service attacks and 6,500 were probe attacks.
@@
 \center
-\caption{Results from Experiment 1}
+\caption{Results from Experiment 1, using only RLD09}
 \vspace{0.20cm}
@@
 
-In the second experiment, seven tests were run.
+In the second experiment,~\cite{6496342, 6559603} pulled some types of attacks out of the training set and kept them for unknown data testing.
+This was to test that the fuzzy genetic algorithm could detect unknown attacks.
+In this experiment, seven tests were run.
 For each test case there were 13 attack types plus normal activity that were in the training data set.
@@
 Table~\ref{tab:fuzGenExp2} shows some of the results from the fuzzy genetic algorithm and a decision tree algorithm, which is another common way of addressing these problems.
-For further information on decision trees, see~\cite{decisionTree}.
+For further information on decision trees, see~\cite{decisionTree}, and for further information on the types of denial of service and probe attacks, see~\cite{typesOfAttacks}.
 When compared with the decision tree algorithm, the fuzzy genetic algorithm has a higher detection rate in all cases except 3 and 5.
-It can be seen that in cases 2 and 4 the decision tree has low detection rates, while the fuzzy genetic algorithm has much higher detection rates.
 
 \begin{table}
-\caption{Unknown Attack Experiment}
+\caption{Unknown Attack Experiment, using only RLD09}
 \vspace{0.20cm}
@@
 \subsubsection{Experiments using Both RLD09 and KDD99}
-The authors of~\cite{6496342, 6559603} also ran experiments that used and compared the RLD09 data set with the KDD99 data set.
+The authors of~\cite{6496342, 6559603} also ran experiments that used both the RLD09 data set and the KDD99 data set in order to compare how the fuzzy genetic algorithm would perform on both.
 They used the KDD99 10\% version file~\cite{KDD99} for both the training dataset and testing dataset.
 
+The first experiment used the fuzzy genetic algorithm to classify normal activity and attacks from both data sets.
 The authors of~\cite{6496342, 6559603} first trained and tested the fuzzy genetic algorithm (Algorithm~\ref{alg:fuzGenAlg}) with the KDD99 data set.
@@
 
-The next experiment was the use of the KDD99 training set with the fuzzy genetifc algorithm to separate the data into two classes.
-Then each specific attack type was extracted and combined with normal activity.
+The next experiment used the fuzzy genetic algorithm to classify types of attacks in the KDD99 data set.
+They used the KDD99 training set, with 158,597 records of denial of service attacks and 1,500 records of probe attacks.
 Ten tests were run, and Table~\ref{tab:kddAttacks} shows the accuracy of detecting some of the cases.
 The results showed that the detection rate of most of the cases were greater than 93\%.
-The results also showed that the behavior of the attacks were highly distinctive from normal activity.
 There were only two cases that had low detection rates, one of which is case 1 in Table~\ref{tab:kddAttacks}.
@@
 
-The final experiment that was run used only the RLD09 data set with the fuzzy genetic algorithm.
+The final experiment that was run used only the RLD09 data set with the fuzzy genetic algorithm to classify types of attacks.
 17 tests were run, and Table~\ref{tab:rldAttacks} shows the accuracy of detecting some of the cases.

### 2013-11-17 Talk/Presentation.tex
Added more to the fuzzy genetic algorithm slides and also the genetic algorithm slides.

--- 2013-11-09 Talk/Presentation.tex
@@
 \usepackage{algorithmic}
+\usepackage{multirow}
 
@@
   \frametitle{Types of Networking Attacks}
-Explain DoS, remote to user, user to root, probe
+  \begin{itemize}
+  	\item Denial of Service (DoS): attacker makes a machine inaccessible to a user by making it too busy to serve legitimate requests.
+
+  	\linespace
+  	\linespace
+
+  	\item Remote to User (R2L): attacker tries to gain access to things a local user would have on the machine.
+
+  	\linespace
+  	\linespace
+
+  	\item User to Root (U2R): attacker starts out with access on the machine and then tries to gain root access to the system.
+
+  	\linespace
+  	\linespace
+
+  	\item Probe: attacker examines a machine in order to collect information about weaknesses or vulnerabilities that in the future could be used to compromise the system.
+  \end{itemize}
 \end{frame}
@@
   \frametitle{Detection Methodologies}
-Explain signature-based and anomaly-based detection
+  \begin{itemize}
+  	\item Signature-based detection: compares well-known patterns of attacks that are already in the intrusion detection system against captured events in order to identify a possible attack.
+  	\item Anomaly-based detection: looks for patterns of activity that are rare and uncommon.
+  \end{itemize}
 \end{frame}
@@
 		\item Has long been a standard data set for intrusion detection.
+		\item Data was processed into 5 million records.
+			\begin{itemize}
+				\item A record is a sequence of TCP packets, between which data flows to and from a source IP address to a target IP address.
+			\end{itemize}
 		\item Data in the set is classified as normal or attack activity.
-		\item KDD99 uses 41 features.
-		\begin{itemize}
-		  	\item \emph{Features} are properties of a \emph{record}, (either an attack or normal activity), that are used to describe the activity.
-		\end{itemize}
+		\item Uses 41 features, which are properties of a record that are used to describe the activity and help to distinguish normal connections from attacks.
 
@@
 		%\item srv\_count: number of connections to the same service as the current connection in the past two seconds.
-		\item serror\_rate: percentage of connections that have ``SYN" errors.
+		%\item serror\_rate: percentage of connections that have ``SYN" errors.
 		%\item same\_srv\_rate: percentage of connections to the same service.
@@
 		\item Data was captured from a university in Bangkok, Thailand.
-		\item The data has 10 million data packets, 17 different types of attacks (divided into denial of service and probe attacks), and 12 features.
+		\item 17 different types of attacks (divided into denial of service and probe attacks), and 12 features.
 	\end{itemize}
@@
 	\begin{itemize}
-		\item Elements of one set are separated into different sets in order to differentiate between normal connections and attacks.
-		\item If-Then format
+		\item A commonly used approach for detecting intrusions and to differentiate between normal connections and attacks is to use rules.
+		\item If-Then format: If (\emph{condition}) then (\emph{consequence}).
 		\begin{itemize}
-			\item If the length of the activity is 4 seconds, then the probability of it being an attack is 100\%.
+			\item The condition is composed of one or more features, and the consequence says if it is an intrusion or not.
+			\item If \emph{duration = 4} then \emph{intrusion}.
 		\end{itemize}				
@@
   \frametitle{Genetic Algorithms}
-
+	\begin{itemize}
+		\item Search technique used to find solutions to problems.
+		\item Mutation: random bits in an individual, or possible solution, are changed.
+		\item Selection: individuals that have a better fitness are chosen over the other individuals.
+		\item Fitness function: determines the quality of a particular individual.
+		\item Crossover: Two individuals swap one of their characteristics with the other to form two new individuals.
+	\end{itemize}
 \end{frame}
@@
   \frametitle{Determining the Accuracy of an Algorithm}
-Explain false positive, false negative, true positive, true negative, detection rate.
+%	\begin{itemize}
+%		\item False positive (FP): intrusion detection system incorrectly identifies normal activity as being an attack. 
+%		\item False negative (FN): intrusion detection system fails to identify harmful activity. 
+%		\item True positive (TP): intrusion detection system correctly identifies activities to be attacks. 
+%		\item True negative (TN): intrusion detection system correctly
+%identifies activities to be normal.
+%		\item Detection rate (DR): number of intrusions detected by the system divided
+%by the total number of intrusions that happen.
+%	\end{itemize}
+\begin{table}
+\begin{tabular}{cccc}
+& & \multicolumn{2}{ c }{Predicted} \\
+& & Not Attack & Attack \\ \cline{3-4}
+\multicolumn{1}{ c }{\multirow{2}{*}{Actual} } &
+\multicolumn{1}{ c| }{Not Attack} & True Negative (TN) & False Positive (FP)\\
+\multicolumn{1}{ c }{} &
+\multicolumn{1}{ c| }{Attack} & False Negative (FN) & True Positive (TP)\\
+\end{tabular}
+\end{table}
+\linespace
+	\begin{itemize}
+		\item Detection rate (DR): number of intrusions detected by the system divided by the total number of intrusions that happen.
+	\end{itemize}
 \end{frame}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section[Genetic Algorithm Implementation]{Genetic Algorithm Implementation}
-\subsection{Algorithm Overview}
-\begin{frame}
-  \frametitle{Algorithm Overview}
+\section[Using Fuzzy Genetic Algorithms]{Using Fuzzy Genetic Algorithms}
 
-\end{frame}
-
-\subsection{Experimental Design and Results}
-\begin{frame}
-  \frametitle{Experimental Design}
-
-\end{frame}
+\subsection{Fuzzy Algorithm}
 
 \begin{frame}
-  \frametitle{Results}
-
+	\frametitle{Fuzzy Logic}
+	\begin{itemize}
+		\item Attacks on systems do not always have a fixed pattern, so fuzzy logic is used to detect patterns that have a behavior that is between normal and unusual.
+		\item Fuzzy logic rules are similar to the rules described before, except that
+consequence is a certainty factor. 
+		\begin{itemize}
+			\item If (\emph{duration} = 6) then (\emph{probability of it being an attack is 50\%}).
+		\end{itemize}
+	\end{itemize}
 \end{frame}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section[Fuzzy Genetic Algorithm Implementation]{Fuzzy Genetic Algorithm Implementation}
-
-\subsection{Fuzzy Algorithm}
 
@@
 \begin{figure}
+\begin{small}
+
 \begin{tabular}{|cccc|c|cccc|c|} \hline
@@
 \hline\end{tabular}
+\end{small}
 \end{figure}
@@
 	\frametitle{Algorithm Overview}
+	\begin{itemize}
+		\item The algorithm generates rules, improves rules, then rules are used to classify the data.
+		\item One record is passed into a rule.
+		\item Each feature in a record is matched to one block of the rule.
+		\item The parameters of each block measure the probability of an attack using the trapezoidal fuzzy rule shape.
+		\item The probabilities of each block are then compared with a threshold to determine if the record represents an attack or normal behavior.
+	\end{itemize}
+\end{frame}
+
+\begin{frame}
+	\frametitle{Algorithm}
 \begin{columns}
 \begin{column}{0.6\textwidth}
+\begin{small}
 \begin{algorithmic}
@@
     \FOR{each feature}
-      \STATE{prob = fuzzy(); // Trapezoidal fuzzy rule shape}
+      \STATE{prob = fuzzy();}
       \STATE{totalprob = totalprob + prob;}
@@
       \STATE{class is attack;}
-      %\ELSE \STATE{class is normal;}
+      \ELSE \STATE{class is normal;}
     \ENDIF
@@
 \end{algorithmic}
+\end{small}
 \end{column}
@@
 	\frametitle{Experiments}
-A variety of experiments were run. Two experiments used just RLD09, and three experiments used KDD99 and RLD09 together.
-
+	\begin{itemize}
+		\item A variety of experiments were run. Two experiments used just RLD09. Three experiments used both the RLD09 and KDD99 in order to compare how the fuzzy GA would perform on both.
+	\end{itemize}
 \end{frame}
@@
 	Experiment 1
+	\begin{itemize}
+		\item Fuzzy GA was used to create DoS and probe detection rules, then the rules were verified with known attack types.
+		\item Two steps in the training process: find a DoS rule, find a probe rule. Both of these rules were then used together in the testing process to identify attacks from the testing data set.
+		\item 10,000 records were used for the training set, 26,500 records were used for the test set.
+	\end{itemize}
+\end{frame}
 
+\begin{frame}
+	\frametitle{Experiments Using Only RLD09}
+	Experiment 1 Results
+\begin{table}
+\begin{small}
+\begin{tabular}{lllllll}
+ & Attack & Normal & Total & FP(\%) & FN(\%) & DR(\%)\\
+DoS Training & 1499 & 8501 & 10000 & 1.46 & 47.50 & 91.64\\
+Probe Training & 2496 & 7504 & 10000 & 1.83 & 15.38 & 94.79\\
+Testing & 10500 & 16000 & 26500 & 1.13 & 4.10 & 97.92\\
+\end{tabular}
+\end{small}
+\end{table}
 \end{frame}
@@
 Experiment 2
+	\begin{itemize}
+		\item Attacks were pulled out of the training set and kept for unknown data testing. This was to test that the fuzzy GA could detect unknown attacks.
+		\item Used fuzzy GA and a decision tree algorithm.
+		\item For each test case there were 13 attack types plus normal activity that were in the training data set. Three attack types were used for the unknown testing data set.
+	\end{itemize}
+\end{frame}
+
+\begin{frame}
+	\frametitle{Experiments Using Only RLD09}
+	Experiment 2 Results
+
+\begin{table}
+\begin{footnotesize}
+\begin{tabular}{llll}
+Test & Unknown & Decision & Fuzzy\\
+Case & Attacks & Tree DR (\%) & Genetic DR (\%)\\ \hline
+
+1 & Adv Port Scan (Probe) & Avg = & Avg =\\
+  & Ack Scan (Probe)                 & 98.33 & 100\\
+  & Xmas Tree (Probe)                 &                 &\\ \hline
+
+2 & UDP Flood (DoS) & Avg = & Avg =\\
+  & Host Scan (Probe) & 46.65 & 99.80\\
+  & UDP Scan (Probe) & &\\ \hline
+
+3 & Jping (DoS) & Avg = & Avg =\\
+  & Syn Scan (Probe) & 99.70 & 98.75\\
+  & Fin Scan (Probe) & &\\ \hline
+
+4 & UDP Flood (DoS) & Avg = & Avg =\\
+  & RCP Scan (Probe) & 70.35 & 98.15\\
+  & Fin Scan (Probe) & &\\ \hline
+
+5 & Http Flood (DoS) & Avg = & Avg =\\
+  & RCP Scan (Probe) & 99.94 & 97.50\\
+  & Fin Scan (Probe) & &\\
+\hline\end{tabular}
+\end{footnotesize}
+\end{table}
 
@@
 Experiment 1
+	\begin{itemize}
+		\item Used fuzzy GA to classify normal activity and attacks from KDD99 and RLD09.
+	\end{itemize}
 
+\begin{table}
+\begin{tabular}{cccccc}
+Data set & Attack & Normal & FP (\%) & FN (\%) & DR (\%)\\ \hline
+KDD99 & 160,117 & 39,337 & 0.13 & 1.55 & 98.72\\
+RLD09 & 10,500 & 16,000 & 1.14 & 3.39 & 97.97\\
+\end{tabular}
+\end{table}
 \end{frame}
@@
 Experiment 2
+	\begin{itemize}
+		\item Used the fuzzy GA to classify types of attacks in KDD99.
+		\item 158,597 records of DoS attacks. 1,500 records of probe attacks.
+	\end{itemize}
+\begin{table}
+\begin{tabular}{cccccc}
+Test & Attack & Type & FP (\%) & FN (\%) & DR (\%)\\ \hline
+1 & Back & DoS & 85.33 & 0.00 & 16.56\\
+2 & Smurf & DoS & 0.76 & 0.10 & 99.73\\
+3 & Portsweep & Probe & 6.40 & 0.00 & 93.66\\
+4 & Satan & Probe & 0.74 & 3.75 & 99.22\\
+\end{tabular}
+\end{table}
 
@@
 Experiment 3
+	\begin{itemize}
+		\item Used the fuzzy GA to classify types of attacks in RLD09.
+		\item 6,400 records of DoS attacks. 10,400 records of probe attacks.
+	\end{itemize}
+\begin{table}
+\begin{tabular}{cccccc}
+Test & Attack & Type & FP (\%) & FN (\%) & DR (\%)\\ \hline
+1 & Smurf & DoS & 0.02 & 0 & 99.98\\
+2 & UDP Flood & DoS & 11.06 & 0 & 89.59\\
+3 & Ackscan & Probe & 0.03 & 0 & 99.97\\
+4 & Synscan & Probe & 0.65 & 4.2 & 99.24\\
+\end{tabular}
+\end{table}
 
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section[Using Genetic Algorithms]{Using Genetic Algorithms}
+\subsection{Algorithm Overview}
+\begin{frame}
+  \frametitle{Algorithm Overview}
+	\begin{itemize}
+		\item The system is divided into 2 phases: a precalculation phase and a detection phase.
+		\item In the precalculation phase, a set of chromosomes are created using training data.
+It takes network data as an input and outputs a set of chromosomes.
+Then this set of chromosomes is used in the detection phase.
+		\item In the detection phase, selection, crossover, and mutation occur, and then the type of the data (whether it is an attack or normal behavior) is predicted.
+	\end{itemize}
+\end{frame}
+
+\subsection{Experimental Design and Results}
+\begin{frame}
+  \frametitle{Experimental Design}
+  \begin{columns}
+
+  \begin{column}{0.5\textwidth}
+	\begin{itemize}
+		\item KDD99 data set.
+		\item Used only the numerical features of KDD99 (34 out of 41 total features).
+		\item Training set: 494,021 records (396,741 of them are attacks). Test set: 311,029 records (250,436 of them are attacks).
+	\end{itemize}
+  \end{column}
+
+  \begin{column}{0.5\textwidth}
+  \begin{table}
+\caption{Number of records}
+\begin{tabular}{lll}
+  & Training & Testing \\
+Normal & 97,280 & 60,593\\
+Probe & 4,107 & 4,166\\
+DoS         & 391,458 & 229,853\\
+U2R & 52 & 228\\
+R2L & 1,124 & 16,189\\
+Total & 494,021 & 311,029\\
+\end{tabular}
+\end{table}
+  \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Results}
+\begin{table}
+\begin{small}
+\begin{tabular}{cccccccc}
+& & \multicolumn{5}{ c }{Predicted} \\
+& & Normal & Probe & DoS & U2R & R2L & \% Correct \\ \cline{3-8}
+\multicolumn{1}{ c }{\multirow{2}{*}{Actual} } &
+\multicolumn{1}{ c| }{Normal} & 42138 & 1421 & 15835 & 486 & 713 & 69.5\\
+\multicolumn{1}{ c }{} &
+\multicolumn{1}{ c| }{Probe} & 398 & 2963 & 654 & 2 & 149 & 71.1 \\
+\multicolumn{1}{ c }{\multirow{2}{*}{} } &
+\multicolumn{1}{ c| }{Dos} & 921 & 432 & 228489 & 1 & 10 & 99.4\\
+\multicolumn{1}{ c }{} &
+\multicolumn{1}{ c| }{U2R} & 146 & 21 & 8 & 43 & 10 & 18.9\\
+\multicolumn{1}{ c }{} &
+\multicolumn{1}{ c| }{R2L} & 11191 & 578 & 3398 & 141 & 881 & 5.4\\
+
+\multicolumn{1}{ c }{} &
+\multicolumn{1}{ c| }{} & & & & & & \\
+
+\multicolumn{1}{ c }{\% Correct} &
+\multicolumn{1}{ c| }{} & 76.9 & 54.7 & 92.0 & 6.4 & 50.0 \\
+\end{tabular}
+\end{small}
+\end{table}
 \end{frame}
@@
 \begin{itemize}
-  \item
-
-  \linespace
-
-  \item
-
-  \linespace
-
-  \item
+  \item The use of genetic algorithms and fuzzy logic in intrusion detection are effective ways of detecting attacks.
+  \item The fuzzy genetic algorithm had a higher detection rate than a decision tree algorithm in most cases.
+  \item Fuzzy genetic algorithms are good at detecting unknown attacks.
 \end{itemize}

### 2013-11-17 Talk/Presentation.tex
Added a better trapezoid figure that has an example, added results for the genetic algorithm, worked on References slide.

@@
 % The text in square brackets is the short version of the date if you need that.
-\date[December '13] % (optional)
+\date[December 2013] % (optional)
 {December 2013 \\ UMM CSci Senior Seminar Conference}
@@
   \frametitle{Measuring the Probability of a Record Being an Attack}
+	\begin{itemize}	
+	\item Feature: duration, 6 seconds.
+	\end{itemize}
+%  \begin{columns}
+%  \begin{column}{0.6\textwidth}
+  \includegraphics[width=0.95\textwidth]{../trapFigExample.pdf}
+%  \end{column}
 
-  \begin{columns}
-  \begin{column}{0.6\textwidth}
-  \includegraphics[width=0.95\textwidth]{../TrapezoidFigure.pdf}
-  \end{column}
+%  \begin{column}{0.4\textwidth}
+%  Example:
+%  \begin{itemize}
+%  	\item Feature: duration (length of the activity in seconds).
+%  	\item a=1, b=3, c=5, d=7
+%  	\item The length of the activity is 6 seconds (between c and d).
+%  	\linespace
+%  	\item prob = $\frac{d-\textrm{data}}{d-c}$ = $\frac{7-6}{7-5}$ = 0.5
+%  \end{itemize}
 
-  \begin{column}{0.4\textwidth}
-  Example:
-  \begin{itemize}
-  	\item Feature: duration (length of the activity in seconds).
-  	\item a=1, b=3, c=5, d=7
-  	\item The length of the activity is 6 seconds (between c and d).
-  	\linespace
-  	\item prob = $\frac{d-\textrm{data}}{d-c}$ = $\frac{7-6}{7-5}$ = 0.5
-  \end{itemize}
-
-  \end{column}
-  \end{columns}
+%  \end{column}
+%  \end{columns}
 \end{frame}
@@
 \begin{small}
-
 \begin{tabular}{|cccc|c|cccc|c|} \hline
@@
 \hline\end{tabular}
+\caption{Based on [Jongsuebsuk-Wattanapongsakorn-Charnsripinyo, 2013]}
 \end{small}
@@
   \frametitle{Results}
-\begin{table}
-\begin{small}
-\begin{tabular}{cccccccc}
-& & \multicolumn{5}{ c }{Predicted} \\
-& & Normal & Probe & DoS & U2R & R2L & \% Correct \\ \cline{3-8}
-\multicolumn{1}{ c }{\multirow{2}{*}{Actual} } &
-\multicolumn{1}{ c| }{Normal} & 42138 & 1421 & 15835 & 486 & 713 & 69.5\\
-\multicolumn{1}{ c }{} &
-\multicolumn{1}{ c| }{Probe} & 398 & 2963 & 654 & 2 & 149 & 71.1 \\
-\multicolumn{1}{ c }{\multirow{2}{*}{} } &
-\multicolumn{1}{ c| }{Dos} & 921 & 432 & 228489 & 1 & 10 & 99.4\\
-\multicolumn{1}{ c }{} &
-\multicolumn{1}{ c| }{U2R} & 146 & 21 & 8 & 43 & 10 & 18.9\\
-\multicolumn{1}{ c }{} &
-\multicolumn{1}{ c| }{R2L} & 11191 & 578 & 3398 & 141 & 881 & 5.4\\
+%\begin{table}
+%\begin{small}
+%\begin{tabular}{cccccccc}
+%& & \multicolumn{5}{ c }{Predicted} \\
+%& & Normal & Probe & DoS & U2R & R2L & \% Correct \\ \cline{3-8}
+%\multicolumn{1}{ c }{\multirow{2}{*}{Actual} } &
+%\multicolumn{1}{ c| }{Normal} & 42138 & 1421 & 15835 & 486 & 713 & 69.5\\
+%\multicolumn{1}{ c }{} &
+%\multicolumn{1}{ c| }{Probe} & 398 & 2963 & 654 & 2 & 149 & 71.1 \\
+%\multicolumn{1}{ c }{\multirow{2}{*}{} } &
+%\multicolumn{1}{ c| }{Dos} & 921 & 432 & 228489 & 1 & 10 & 99.4\\
+%\multicolumn{1}{ c }{} &
+%\multicolumn{1}{ c| }{U2R} & 146 & 21 & 8 & 43 & 10 & 18.9\\
+%\multicolumn{1}{ c }{} &
+%\multicolumn{1}{ c| }{R2L} & 11191 & 578 & 3398 & 141 & 881 & 5.4\\
 
-\multicolumn{1}{ c }{} &
-\multicolumn{1}{ c| }{} & & & & & & \\
+%\multicolumn{1}{ c }{} &
+%\multicolumn{1}{ c| }{} & & & & & & \\
 
-\multicolumn{1}{ c }{\% Correct} &
-\multicolumn{1}{ c| }{} & 76.9 & 54.7 & 92.0 & 6.4 & 50.0 \\
-\end{tabular}
-\end{small}
-\end{table}
+%\multicolumn{1}{ c }{\% Correct} &
+%\multicolumn{1}{ c| }{} & 76.9 & 54.7 & 92.0 & 6.4 & 50.0 \\
+%\end{tabular}
+%\end{small}
+%\end{table}
+	\begin{figure}
+  \includegraphics[width=0.95\textwidth]{../resultsForGA.JPG}
+  \caption{From [Hoque-Mukit-Bikas, 2012]}
+  	\end{figure}
+
 \end{frame}
@@
   \item The use of genetic algorithms and fuzzy logic in intrusion detection are effective ways of detecting attacks.
+  \item The genetic algorithm had a high detection rate for denial of service attacks.
   \item The fuzzy genetic algorithm had a higher detection rate than a decision tree algorithm in most cases.
@@
 
+	\begin{small}
+	\bibitem{6496342}
+Jongsuebsuk, P. and Wattanapongsakorn, N. and Charnsripinyo, C.
+\newblock Network intrusion detection with Fuzzy Genetic Algorithm for unknown attacks.
+\newblock In \emph{2013 International Conference on Information Networking (ICOIN)}, pages 1-5, 2013.
+
+	\bibitem{6559603}
+Jongsuebsuk, P. and Wattanapongsakorn, N. and Charnsripinyo, C.
+\newblock Real-time intrusion detection with fuzzy genetic algorithm.
+\newblock In \emph{2013 10th International Conference on Electrical Engineering/Electronics, Computer, Telecommunications and Information Technology (ECTI-CON)}, pages 1-6, 2013.
+
+	\bibitem{DBLP:journals/corr/abs-1204-1336}
+Mohammad Sazzadul Hoque, Md. Abdul Mukit, Md. Abu Naser Bikas.
+\newblock An Implementation of Intrusion Detection System Using Genetic Algorithm.
+\newblock \emph{CoRR}, abs/1204.1336, 2012.
+	\end{small}
+
   	\end{thebibliography}
+
+  	\linespace
+
+  	\begin{center}
+  	\begin{small}
+  	See my Senior Seminar paper for additional references.
+  	\end{small}
+  	\end{center}
 

### 2013-11-21 Talk/Presentation.tex
Switched some sections around and removed some stuff.

@@
   \frametitle{Determining the Accuracy of an Algorithm}
-%	\begin{itemize}
-%		\item False positive (FP): intrusion detection system incorrectly identifies normal activity as being an attack. 
-%		\item False negative (FN): intrusion detection system fails to identify harmful activity. 
-%		\item True positive (TP): intrusion detection system correctly identifies activities to be attacks. 
-%		\item True negative (TN): intrusion detection system correctly
-%identifies activities to be normal.
-%		\item Detection rate (DR): number of intrusions detected by the system divided
-%by the total number of intrusions that happen.
-%	\end{itemize}
 \begin{table}
@@
 	\begin{itemize}
-		\item Detection rate (DR): number of intrusions detected by the system divided by the total number of intrusions that happen.
+		\item Detection rate (DR): 
 	\end{itemize}
@@
 \section[Using Genetic Algorithms]{Using Genetic Algorithms}
-\subsection{Algorithm Overview}
-\begin{frame}
-  \frametitle{Algorithm Overview}
-	\begin{itemize}
-		\item The system is divided into 2 phases: a precalculation phase and a detection phase.
-		\item In the precalculation phase, a set of chromosomes are created using training data.
-It takes network data as an input and outputs a set of chromosomes.
-Then this set of chromosomes is used in the detection phase.
-		\item In the detection phase, selection, crossover, and mutation occur, and then the type of the data (whether it is an attack or normal behavior) is predicted.
-	\end{itemize}
-\end{frame}
-
 \subsection{Experimental Design and Results}
@@
   \begin{table}
-\caption{Number of records}
+\caption{ }
 \begin{tabular}{lll}
@@
 %\end{table}
-	\begin{figure}
-  \includegraphics[width=0.95\textwidth]{../resultsForGA.JPG}
-  \caption{From [Hoque-Mukit-Bikas, 2012]}
-  	\end{figure}
+
+%	\begin{figure}
+%  \includegraphics[width=0.95\textwidth]{../resultsForGA.JPG}
+%  \caption{From [Hoque-Mukit-Bikas, 2012]}
+%  	\end{figure}
 
@@
   \item The use of genetic algorithms and fuzzy logic in intrusion detection are effective ways of detecting attacks.
-  \item The genetic algorithm had a high detection rate for denial of service attacks.
+%  \item The genetic algorithm had a high detection rate for denial of service attacks.
   \item The fuzzy genetic algorithm had a higher detection rate than a decision tree algorithm in most cases.

### 2013-11-23 paper.tex
Changed the title of one of the subsections.
Made some changes to the fuzzy genetic algorithm section, and got rid of some stuff in the genetic algorithm section.

@@
 \usepackage{algorithm}
-\usepackage{multirow}
+%\usepackage{multirow}
 
@@
 \begin{abstract}
+Intrusion detection systems provide one way of detecting attacks on systems, by monitoring network activities for malicious or abnormal behaviors.
 This paper describes two different ways of training an intrusion detection system about possible attacks to a system: genetic algorithms and fuzzy logic.
-I will describe how genetic algorithms are used in intrusion detection systems and compare the results to the winning entry of the KDD99 Classifier Learning Contest.
-I will then describe how fuzzy genetic algorithms are used and compare those results with a decision tree algorithm.
+The use of genetic algorithms and fuzzy logic in intrusion detection are effective ways of detecting attacks.
+I will describe an approach to using fuzzy genetic algorithms and compare those results with results obtained using a decision tree.
+I will then describe the results from using a genetic algorithm and compare those with the winning entry of the KDD99 Classifier Learning Contest.
+\mycomment{work on abstract}
+
 \end{abstract}
@@
 % A category with the (minimum) three required fields
-\category{}{Security and Privacy}{Intrusion Detection Systems}
+%\category{}{Security and Privacy}{Intrusion Detection Systems}
 
-\terms{Security}
+%\terms{Security}
 
-\keywords{Intrusion detection, genetic algorithm, fuzzy logic, KDD99, RLD09, computer security}
+\keywords{Intrusion detection, genetic algorithms, fuzzy logic, KDD99, RLD09, computer security}
 
 \section{Introduction}
-An attack on important and confidential data is a concern for many people, so it is important to have a way to detect and analyze these attacks.
-A way of detecting attacks is by using an intrusion detection system.
-An intrusion detection system is a device that monitors network activities for malicious or abnormal behaviors and then produces reports and alerts~\cite{DBLP:journals/corr/abs-1204-1336}.
+ \mycomment{WORK ON INTRODUCTION} The CSci computer lab at the University of Minnesota, Morris gets large numbers of login attempts that are attempts at intrusion.
+An attacker could be trying to gain root access to the system, which would then give them the ability to do things like delete files of other users and change user passwords.
+If the university had an intrusion detection system, it would be possible to classify those attempts into legitimate and illegitimate attempts to login.
+Then it would be possible to block IP addresses that are generating large numbers of attacks.
+
+An attack on systems is a concern for many people, so it is important to have a way to detect and analyze these attacks.
+Intrusion detection systems provide one way of detecting attacks by monitoring network activities for malicious or abnormal behaviors and then producing reports and alerts~\cite{DBLP:journals/corr/abs-1204-1336}.
 There are various ways of training the intrusion detection system about possible threats.
-Two approaches that I will talk about in this paper are the use of genetic algorithms and fuzzy logic.
+Two approaches that I will talk about in this paper are genetic algorithms and fuzzy logic.
 
-In Section~\ref{sec:background} I will give some background on intrusion detection.
+In Section~\ref{sec:background} I will give some background on intrusion detection and genetic algorithms.
 I will describe the main types of networking attacks in Section~\ref{sec:typesAttacks}, and ways of detecting attacks in Section~\ref{sec:detectionMeth}.
 In Section~\ref{sec:dataSets} I will describe two data sets that are commonly used in intrusion detection.
-I will explain how \emph{rules} are used to differentiate between normal connections and attacks in Section~\ref{sec:rules}, and then give information on genetic algorithms in Section~\ref{sec:GA}.
 After that, I will describe the measures that are used in determining the accuracy of an intrusion detection system in Section~\ref{sec:PosNeg}.
-In Section~\ref{sec:genAlgImp} I will describe the use of a genetic algorithm in an intrusion detection system, and in Section~\ref{sec:fuzGenAlgImp} I will describe the use of a fuzzy genetic algorithm.
+I will then give information on genetic algorithms in Section~\ref{sec:GA}.
+In Section~\ref{sec:fuzGenAlgImp} I will describe the use of a fuzzy genetic algorithm in an intrusion detection system, and in Section~\ref{sec:genAlgImp} I will describe the results of using a genetic algorithm.
 Section~\ref{sec:conclusion} has some conclusions about the use of genetic algorithms and fuzzy genetic algorithms in intrusion detection systems.
@@
 \label{sec:typesAttacks}
-There are four main types of networking attacks that this paper will address: \emph{denial of service}, \emph{remote to user attacks}, \emph{user to root attacks}, and \emph{probing}.
+There are four main types of networking attacks that this paper will address: \emph{denial of service}, \emph{remote to user attacks}, \emph{user to root attacks}, and \emph{probe attacks}.
 Each attack that happens on a network can be placed into one of these categories.~\cite{DBLP:journals/corr/abs-1204-1336}
@@
 For example, many systems lock out a user from an account after a certain number of failed login attempts.
-An attacker would be able to use this to prevent legitimate users from logging in~\cite{DoSAttacks}.
-Remote to user (R2L) attacks happen when an attacker sends packets to a machine over the network in order to gain access to things a local user would have on the machine.
+An attacker would be able to use this to prevent legitimate users from logging in, by intentionally failing to log in enough times to lock the account~\cite{DoSAttacks}.
+Remote to user, also known as remote to local (R2L) attacks happen when an attacker sends packets to a machine over the network in order to gain access to things a local user would have on the machine.
 An example is when an attacker tries to gain access to a machine by guessing possible usernames and passwords.
 User to root (U2R) attacks happen when an attacker starts out with access on the machine and then tries to gain root access to the system.
-For example, if a program expects a user to input their name, the programmer has to decide how many characters that name buffer will require.
-Assume the program allocates 20 characters for the name buffer.
-Suppose the user's name has 35 characters.
-The last 15 characters will overflow the buffer, which could then overwrite the instructions that are to be executed next.
-An attacker can cause commands to be executed by manipulating the data that overflows.
-Probing happens when an attacker examines a machine in order to collect information about weaknesses or vulnerabilities that in the future could be used to compromise the system.~\cite{DBLP:journals/corr/abs-1204-1336, typesOfAttacks}
+If an attacker has access to an account on the machine, then they could potentially run programs that take advantage of operating system weaknesses to gain root access.
+Probe attacks happen when an attacker examines a machine in order to collect information about weaknesses or vulnerabilities that in the future could be used to compromise the system.
+For example, the attacker could be trying to determine what version of a software is being run on that machine, and if that version has a known issue then that allows them to attempt to attack that.~\cite{DBLP:journals/corr/abs-1204-1336, typesOfAttacks}
 
-\subsection{Detection Methodologies}
+\subsection{Detection Methodologies and Rules}
 \label{sec:detectionMeth}
 There are two different ways of detecting attacks: \emph{signature-based detection} and \emph{anomaly-based detection}.
-
 A signature is a pattern that corresponds to a known attack.
-Signature-based detection compares well-known patterns of attacks that are already in the intrusion detection system against captured events in order to identify a possible attack.
-It is a simple and effective way to detect known attacks.
+Signature-based detection compares well-known patterns of attacks that are already known to the IDS against captured events in order to identify a possible attack.
+It is a simple and effective way to detect known attacks, but is ineffective against new kinds of unknown attacks.
 Signature-based detection is also called \emph{knowledge-based detection} or \emph{misuse detection}.~\cite{Liao201316}
@@
 Anomaly-based detection looks for patterns of activity that are rare and uncommon.
-It is an effective way to detect new attacks.
+It is harder to do than signature-based detection, but it can be an effective way to detect new, unknown attacks.
 Anomaly-based detection is also called \emph{behavior-based detection}.~\cite{DBLP:journals/corr/abs-1204-1336}
+
+A commonly used approach for detecting intrusions is to use rules.
+Rules are represented by if-then statements in the following format: If (\emph{condition}) then (\emph{consequence}).
+The condition part of the rule is composed of one or more features, and the consequence of the rule says if it is an intrusion or not.
+For example: if (\emph{duration} = 1, and \emph{src\_bytes} = 0, and \emph{serror\_rate} = 50) then \emph{intrusion}.~\cite{Fries:2008:FAN:1388969.1389037}
 
@@
 Two different data sets are used in this paper to evaluate the performance of intrusion detection systems: KDD99 and RLD09.
-The KDD99 data set is a benchmark data set that was generated by simulating a military network environment in 1999.
-It has long been a standard data set for intrusion detection.
-The data was processed into five million records.
-A record is a sequence of TCP packets, between which data flows to and from a source IP address to a target IP address.
-The data in the set is classified as either normal or attack activity.
+KDD99 is a benchmark data set that was generated by simulating a military network environment in 1999, and it has long been a standard data set for intrusion detection.
+The data was processed into five million records, where a record is a sequence of TCP packets, between which data flows to and from a source IP address to a target IP address.
+Each record in the data set is classified as either normal or attack activity.
 KDD99 uses 41 \emph{features}, which are properties of a record that are used to describe the activity and help to distinguish normal connections from attacks.~\cite{KDD99}
 
-In research conducted by P. Jongsuebsuk,
-
-N. Wattanapongsakorn, and
-C.
-Charnsripinyo~\cite{6496342, 6559603}, which will be discussed in Section~\ref{sec:fuzGenAlgImp}, eight out of the 41 KDD99 features~\cite{KDD99} were used.
-They decided on these 8 features based on research done in~\cite{Fries:2008:FAN:1388969.1389037}.
-The features are:
+In the research discussed in Section~\ref{sec:fuzGenAlgImp}~\cite{6496342, 6559603}, eight of the 41 KDD99 features were used based on research done in~\cite{Fries:2008:FAN:1388969.1389037}.
+These eight features are:
 \begin{enumerate}
-  \item duration: the length of the normal or attack activity in seconds.
+  \item duration: the length of the record in seconds.
   \item src\_bytes: the number of bytes sent from source to destination. Source is the user who may or may not be an attacker, and destination is the server being potentially attacked.
-  \item num\_failed\_logins: the number of failed login attempts.
-  \item root\_shell: returns 1 if root shell is obtained, which means that the user is able to log in as root.
-This gives them the ability to do things like add accounts and change user passwords.
-If root shell is not obtained, 0 is returned.
+  \item num\_failed\_logins: the number of failed login attempts in this record.
+  \item root\_shell: returns 1 if root shell is obtained, which means that the user is able to log in as root. If root shell is not obtained, 0 is returned.
   \item num\_access\_files: the number of operations on access control files.
@@
 An example would be (Alice, delete), which would give Alice permission to delete the file.~\cite{accessControl}
-  \item srv\_count: the number of connections to the same service as the current connection in the past two seconds.
+  \item srv\_count: the number of connections in other records.
   \item serror\_rate: the percentage of connections that have ``SYN" errors.
@@
 The connection is established when the client sends an ACK back to the server.
-A SYN error is a failure that happens early in this process.~\cite{TCP}
-  \item same\_srv\_rate: the percentage of connections to the same service.
+A SYN error is a failure to get an ACK back.~\cite{TCP}
+  \item same\_srv\_rate: the percentage of connections to the same service. \mycomment{Have example of service}
 \end{enumerate}
 
-The KDD99 data set is 14 years old, and newer attack types are not included in it because of its age. Because of this, P. Jongsuebsuk, N. Wattanapongsakorn, and
-C.
-Charnsripinyo~\cite{6496342, 6559603} created their own data set, RLD09, to use in their experiments.
+The KDD99 data set is 14 years old, and newer attack types are not included in it because of its age.
+Because of this, the authors of~\cite{6496342, 6559603} created their own data set, RLD09, to use in their experiments.
 To create the data set, the authors captured network data from a university in Bangkok, Thailand.
-RLD09 has 17 different types of attacks that can be divided into denial of service attacks and probe attacks.
-It also has normal network activity.
-RLD09 uses 12 features, which include the number of packets, source ports, and destination ports.
-For further information, see~\cite{6496342, 6559603}.
-RLD09 only has 12 features because the number of features affect the speed and memory consumption of the intrusion detection system.
+As well as normal network activity, RLD09 has 17 different types of attacks which can be classified as either denial of service attacks or probe attacks.
+RLD09 uses 12 features.
+For further information about the features, see~\cite{6496342, 6559603}.
 
-\subsection{Rules}
-\label{sec:rules}
-A commonly used approach for detecting intrusions and to differentiate between normal connections and attacks is to use rules.
-Rules are represented by if-then statements in the following format: If (\emph{condition}) then (\emph{consequence}).
-The condition part of the rule is composed of one or more features, and the consequence of the rule says if it is an intrusion or not.
-For example: if (\emph{duration} = 1, and \emph{src\_bytes} = 0, and \emph{serror\_rate} = 50) then \emph{intrusion}.~\cite{Fries:2008:FAN:1388969.1389037}
+In the research discussed in Section~\ref{sec:genAlgImp}, the authors of ~\cite{DBLP:journals/corr/abs-1204-1336} used only the numerical features of the KDD99 data set (34 out of 41 total features).
 
-\subsection{Genetic Algorithm}
-\label{sec:GA}
-Genetic algorithms are a search technique used to find solutions to problems.
-Operations analogous to biological \emph{mutation}, \emph{selection}, and \emph{crossover} are used to evolve and improve rules.
-
-Mutation is where random bits in an \emph{individual}, or possible solution, are changed.
-Selection is where individuals that have a better \emph{fitness} are chosen over the other individuals.
-The fitness function determines the quality of a particular individual.
-Crossover is where two individuals swap one of their characteristics with the other to form two new individuals.
-
-The solution to a problem is represented as a chromosome.
-First, a randomly generated population of chromosomes is created.
-According to the characteristics of the problem, the positions of each chromosome are encoded as bits, characters, or numbers.
-Then mutation, selection, and crossover are applied to each generation and eventually the best solution is found.~\cite{DBLP:journals/corr/abs-1204-6416}
-
-\subsection{False Positives, False Negatives, True Positives, True Negatives}
+\subsection{Determining the Accuracy of an Algorithm}
 \label{sec:PosNeg}
@@
 
-A \emph{false positive} (FP) happens when an intrusion detection system incorrectly identifies normal activity as being an attack.
-A \emph{false negative} (FN) happens when an intrusion detection system fails to identify harmful activity.
-A \emph{true positive} (TP) happens when an intrusion detection system correctly identifies activities to be attacks.
-A \emph{true negative} (TN) happens when an intrusion detection system correctly identifies activities to be normal.
+The \emph{false negative} (FN) rate is the percentage that attacks are misclassified from the total number of attack records.
+The \emph{false positive} (FP) rate is the percentage that normal records are classified as attacks from the total number of normal records.
+The \emph{true negative} (TN) rate is the percentage that normal records are classified correctly from the total number of normal records.
+The \emph{true positive} (TP) rate is the percentage that attacks are classified correctly from the total number of attack records.~\cite{6496342, bc1_ecindm}
 
-The \emph{detection rate} (DR) of an intrusion detection system is the number of intrusions detected by the system divided by the total number of intrusions.~\cite{ids}
+%A \emph{false positive} (FP) happens when an IDS incorrectly identifies normal activity as being an attack.
+A \emph{false negative} (FN) happens when an IDS fails to identify harmful activity.
+A \emph{true positive} (TP) happens when an IDS correctly identifies activities to be attacks.
+A \emph{true negative} (TN) happens when an IDS correctly identifies activities to be normal.~\cite{ids}
 
-\section{Using Genetic Algorithms}
-\label{sec:genAlgImp}
+\subsection{Genetic Algorithms}
+\label{sec:GA}
+\mycomment{Work on this section} Genetic algorithms (GA) are a search technique used to find solutions to problems.
+Operations analogous to biological \emph{mutation}, \emph{selection}, and \emph{crossover} are used to evolve and improve rules.
 
-\subsection{Algorithm Overview}
-In research conducted by M.
-S.
-Hoque, M.
-A.
-Mukit, and M.
-A.
-N.
-Bikas~\cite{DBLP:journals/corr/abs-1204-1336}, a genetic algorithm was used to make their intrusion detection system.
-The system is divided into two phases: a precalculation phase and a detection phase.
-In the precalculation phase, a set of chromosomes are created using training data.
-Then this set of chromosomes is used in the detection phase for comparison.
-Algorithm~\ref{alg:precalc} is used in the precalculation phase.
-It initializes chromosomes for comparison.
-It takes the network data as an input and outputs a set of chromosomes.
+Possible solutions to problems can be represented in a variety of problem dependent ways, such as bit strings.
+First, a randomly generated population of rules is created.
+Then mutation, selection, and crossover are applied to each generation until an acceptable solution is found or some time limit is exceeded.
 
-\begin{algorithm}
-\caption{Major steps in precalculation}
-\label{alg:precalc}
-\begin{algorithmic}
-\STATE{range = 0.125}
-\FOR{each training data}
-  \IF{it has neighboring chromosome within range}
-    \STATE{Merge it with the nearest chromosome}
-  \ELSE \STATE{Create new chromosome with it}
-  \ENDIF
-\ENDFOR
-\end{algorithmic}
-\end{algorithm}
-
-In the detection phase, the precalculated set of chromosomes is used to find the fitness of each chromosome in the population.
-Selection, crossover, and mutation occur, and then the type of the data (whether it is an attack or normal behavior) is predicted.
-Algorithm~\ref{alg:detection} is used in the detection phase.
-It takes the precalculated set of chromosomes as an input and outputs the type of data.
-
-\begin{algorithm}
-\caption{Major steps in detection}
-\label{alg:detection}
-\begin{algorithmic}
-\STATE{Initialize the population}
-\WHILE{number of generation is not reached}
-  \FOR{each chromosome in the population}
-    \FOR{each precalculated chromosome}
-      \STATE{Find fitness // Fitness function is the standard deviation equation with distance}
-    \ENDFOR
-    \STATE{Assign optimal fitness as the fitness of that
-
-    chromosome}
-  \ENDFOR
-  \STATE{Remove some chromosomes with worse fitness}
-  \STATE{Apply crossover to the selected pair of chromosomes of the population // crossover rate~=~0.15}
-  \STATE{Apply mutation to each chromosome of the
-
-  population~//~mutation rate~=~0.35}
-\ENDWHILE
-\end{algorithmic}
-\end{algorithm}
-
-\subsection{Experimental Design and Results}
-The authors of~\cite{DBLP:journals/corr/abs-1204-1336} used the KDD99 data set.
-They used only the numerical features of the KDD99 data set (34 out of 41 total features).
-For the training data set, the KDD99 10\% version file was used.
-This is a 10\% subset of the full KDD99 data set.
-For the test data set, the KDD99 corrected version file was used.
-This file corrects various typos that were in other files.~\cite{KDD99} The training set has a total of 494,021 records, and 396,741 of them are attacks.
-The test set has a total of 311,029 records, and 250,436 of them are attacks.
-Table~\ref{tab:numberOfRecords} shows the number of records of normal and attack activity in the training and test sets.
-
-\begin{table}
-\center
-\caption{Number of records}
-\vspace{0.20cm}
-\begin{tabular}{lll}
-  & Training & Testing \\ 
-Normal & 97,280 & 60,593\\
-Probe  & 4,107  & 4,166\\
-DoS	   & 391,458 & 229,853\\
-U2R    & 52      & 228\\
-R2L    & 1,124   & 16,189\\
-Total  & 494,021 & 311,029\\
-\end{tabular}
-\center
-\label{tab:numberOfRecords}
-\end{table}
-
-\begin{table*}
-\center
-\caption{Results for Genetic Algorithm Experiment}
-\vspace{0.20cm}
-\begin{tabular}{cccccccc}
-& & \multicolumn{5}{ c }{Predicted label} \\ 
-& & Normal & Probe & DoS & U2R & R2L & \% Correct \\ \cline{3-8}
-\multicolumn{1}{ c }{\multirow{2}{*}{Actual class} } &
-\multicolumn{1}{ c| }{Normal} & 42138 & 1421 & 15835 & 486 & 713 & 69.5\\
-\multicolumn{1}{ c  }{}                        &
-\multicolumn{1}{ c| }{Probe} & 398 & 2963 & 654 & 2 & 149 & 71.1   \\
-\multicolumn{1}{ c  }{\multirow{2}{*}{} } &
-\multicolumn{1}{ c| }{Dos} & 921 & 432 & 228489 & 1 & 10 & 99.4\\
-\multicolumn{1}{ c  }{}                        &
-\multicolumn{1}{ c| }{U2R} & 146 & 21 & 8 & 43 & 10 & 18.9\\
-\multicolumn{1}{ c  }{}                        &
-\multicolumn{1}{ c| }{R2L} & 11191 & 578 & 3398 & 141 & 881 & 5.4\\
-
-\multicolumn{1}{ c  }{}                        &
-\multicolumn{1}{ c| }{} &  &  &  &  &  & \\
-
-\multicolumn{1}{ c  }{\% Correct}                        &
-\multicolumn{1}{ c| }{} & 76.9 & 54.7 & 92.0 & 6.4 & 50.0 \\
-\end{tabular}
-\center
-\label{tab:genAlgResults}
-\end{table*}
-
-\begin{table*}
-\center
-\caption{Results for Genetic Algorithm Experiment}
-\vspace{0.20cm}
-\begin{tabular}{cccc}
-%\cline{3-8}
-& & \multicolumn{2}{ c }{Predicted label} \\ 
-& & Normal & Intrusion\\ \cline{3-4}
-\multicolumn{1}{ c }{\multirow{2}{*}{Actual class} } &
-\multicolumn{1}{ c| }{Normal} & True Negative (42,138) & False Positive (18,455)\\
-\multicolumn{1}{ c  }{}                        &
-\multicolumn{1}{ c| }{Probe} & False Negative (12,528) & True Positive (237,908)\\
-\end{tabular}
-\center
-\label{tab:genAlgResults2}
-\end{table*}
-
-In the detection phase, an initial population is created and then is compared with each chromosome that was created in the precalculation phase.
-Crossover and mutation happen in the detection population in order to create a new population.
-The results from running the genetic algorithm are shown in Tables~\ref{tab:genAlgResults} and~\ref{tab:genAlgResults2}.
-The detection rate was 94.9\%.
-The false positive rate was 30.5\%.
-The false negative rate was 5\%.
-The true negative rate was 69.5\%.
-
-The authors of~\cite{DBLP:journals/corr/abs-1204-1336} compared their results with the winning entry of the KDD99 Classifier Learning Contest~\cite{KDD99Contest}.
-The winning entry used a decision tree algorithm, which is another common algorithm that is used.
-For further information on decision trees, see~\cite{decisionTree}.
-The authors of~\cite{DBLP:journals/corr/abs-1204-1336} found that they had a better detection rate for denial of service and user to root attacks than the winning entry.
-For the winning entry, the detection rate of denial of service was 97.1\%, and for user to root it was 13.2\%.
-In~\cite{DBLP:journals/corr/abs-1204-1336} the detection rate of denial of service was 99.4\% and for user to root it was 18.9\%.
+Mutation is where random bits in an \emph{individual}, or possible solution, are changed.
+Selection is where individuals that have a better \emph{fitness} are chosen over the other individuals.
+The fitness function determines the quality of a particular individual.
+For example, in intrusion detection systems the fitness measures the performance of a rule.
+Crossover is where two individuals swap sequences of bits to form two new individuals.~\cite{DBLP:journals/corr/abs-1204-6416}
 
@@
 \label{sec:fuzGenAlgImp}
-The focus of~\cite{6496342, 6559603} is on detecting new or unknown types of attacks in a network.
-The intrusion detection system used is able to identify normal network activity as well as attacks using a fuzzy genetic algorithm.
-This kind of algorithm is able to learn new attacks, and has a high detection rate.
+The IDS used in~\cite{6496342, 6559603} is able to identify normal network activity as well as attacks using a fuzzy genetic algorithm. This kind of algorithm is able to learn new attacks.
 
@@
 Attacks on systems do not always have a fixed pattern, so fuzzy logic is used to detect patterns that have a behavior that is between normal and unusual.
-Fuzzy logic rules are similar to the rules described in Section~\ref{sec:rules}, except that \emph{consequence} is a certainty factor.
+Fuzzy logic rules are similar to the rules described in Section~\ref{sec:detectionMeth}, except that \emph{consequence} is a certainty factor.
 For example, if (\emph{duration} = 6) then (\emph{probability of it being an attack is 50\%}).
+\mycomment{certainty vs probability}
 
-In order to measure the probability that a record is an attack, a trapezoidal shape was used in~\cite{6496342, 6559603}.
-This is shown in Figure~\ref{fig:trapFigure}.
+In order to measure the probability that a record is an attack, a trapezoidal shape was used in~\cite{6496342, 6559603}; this is shown in Figure~\ref{fig:trapFigure}.
 The trapezoidal shape has four parameters: $a, b, c$, and $d$.
@@
 \psfig{file=TrapezoidFigure.pdf, width=3.3in}
-\caption{Used to measure the probability that a record is an attack}
+\caption{Illustration of fuzzy rule used to measure the probability that a record is an attack. Also see Algorithm~\ref{alg:fuzAlg}. This is used in~\cite{6496342, 6559603}.}
 \label{fig:trapFigure}
@@
 
-\begin{algorithm}
-\caption{Fuzzy Algorithm}
+\begin{algorithm}[t]
+\caption{Fuzzy Algorithm that is based on an algorithm used in~\cite{6496342, 6559603}, but a few corrections have been made. Also see Figure~\ref{fig:trapFigure}.}
 \label{alg:fuzAlg}
 \begin{algorithmic}
-\IF{data value is between $b$ and $c$}
+\IF{$b \leq \textrm{data} \leq c$}
   \STATE{prob = 1.0}
-\ELSIF{data value is between $a$ and $b$}
+\ELSIF{$a < \textrm{data} < b$}
   \STATE{
@@
   }
-\ELSIF{data value is between $c$ and $d$}
+\ELSIF{$c < \textrm{data} < d$}
   \STATE{
@@
 
-An example of using the fuzzy shape and algorithm is: suppose that the feature is duration, and suppose it is 6 seconds.
+An example of using the fuzzy logic approach is: suppose that the feature is duration, and suppose it is 6 seconds, so then data = 6.
 Suppose that $a$ = 1, $b$ = 3, $c$ = 5, $d$ = 7.
-Then the probability of an attack is equal to
+Because data (6) is between $c$ and $d$, then the probability of an attack is equal to
 
@@
 
-The four parameters a, b, c, and d are encoded into blocks of binary strings, where each block is a feature with values between 0.0 and 7.0.
+The four parameters $a, b, c$, and $d$ are encoded into blocks of binary strings, where each block is a feature with values between 0 and 7.
 See Figure~\ref{fig:fuzEncodingForFeature} for an example of a block.
-A rule has 12 blocks of features, and at the end of the string is the type of attack.
-An example of this is Figure~\ref{fig:rule}.
+A rule has one block for each of 12 features followed at the end by a marker indicating the type of attack.
+An example of this is shown in Figure~\ref{fig:rule}.
+The authors of~\cite{6496342, 6559603} compute the probability for each of the 12 blocks, and if the sum of those is greater than a threshold, then it will be declared as an attack.
 
@@
 \centering
-\caption{Fuzzy encoding for a feature}
+\caption{Fuzzy encoding for a feature that is used in~\cite{6496342, 6559603}.}
 \vspace{0.20cm}
@@
 \centering
-\caption{A rule with 12 blocks of features}
+\caption{A rule with 12 blocks of features used in~\cite{6496342, 6559603}.}
 \vspace{0.20cm}
@@
 \subsection{Algorithm Overview}
-The algorithm that is used in~\cite{6496342, 6559603} first randomly generates rules.
-Then the rules are improved in the training phase, which can be seen in Algorithm~\ref{alg:fuzGenAlg}.
-After that, the rules are used to classify the data in the testing phase.
-Algorithm~\ref{alg:fuzGenAlg} describes the fuzzy genetic algorithm that is used.
+The algorithm in~\cite{6496342, 6559603} first randomly generates rules.
+Then the rules are improved in the training phase, which can be seen in Algorithm~\ref{alg:fuzGenAlg}, which describes the fuzzy genetic algorithm that is used.
 One record (either an attack or normal activity) is passed into a rule.
@@
 The parameters of each block measure the probability of an attack using the trapezoidal fuzzy rule shape.
-The probabilities of each block are then compared with a threshold to determine if the record represents an attack or normal behavior.
+The sum of the probabilities from each block are then compared with a threshold to determine if the record represents an attack or normal behavior.
 
-\begin{algorithm}
-\caption{Fuzzy Genetic Algorithm}
+\begin{algorithm}[t]
+\caption{Fuzzy Genetic Algorithm that is based on an algorithm used in~\cite{6496342, 6559603}, but a few corrections have been made.}
 \label{alg:fuzGenAlg}
 \begin{algorithmic}
+\FOR{each rule}
+
   \FOR{each record}
-
-  \FOR{each rule}
     \FOR{each feature}
@@
   \STATE{find $A$, $B$, $\alpha$, and $\beta$}
-  \STATE{// $A$ is total number of attack records.
-$B$ is total number of normal records.
-$\alpha$ is total number of attack records correctly identified as attack.
-$\beta$ is total number of normal records incorrectly classified as attack.}
+  \STATE{// $A$ is the total number of attack records.
+$B$ is the total number of normal records.
+$\alpha$ is the total number of attack records correctly identified as attack.
+$\beta$ is the total number of normal records incorrectly classified as attack.}
 \ENDFOR
@@
 \STATE{preserve\_best()}
-\STATE{crossover() // single-point crossover}
-\STATE{mutation() // mutation rate~=~0.30}
+\STATE{crossover()}
+\STATE{mutation()}
 \end{algorithmic}
@@
 
-The fitness function to be maximized is:
+The fitness function to be maximized in Algorithm~\ref{alg:fuzGenAlg} is:
 \begin{equation*}
@@
 \end{equation*}
-$A$ is total number of attack records.
-$B$ is total number of normal records.
-$\alpha$ is total number of attack records correctly identified as attack.
-$\beta$ is total number of normal records incorrectly classified as attack.
+where $A$ is the total number of attack records, $B$ is the total number of normal records, $\alpha$ is the total number of attack records correctly identified as attack, and $\beta$ is the total number of normal records incorrectly classified as attack.
 
 A population size of 10 was used for each generation.
-An individual in the population represents a possible detection rule.
-The two best individuals from a present generation are preserved for the next generation.
-The other individuals in the new generation come from mutation and single-point crossover.
+The two best individuals from the present generation are preserved for the next generation.
+The other individuals in the new generation come from mutation and crossover.
 
@@
 \subsubsection{Experiments using Only RLD09}
-The experiments using only RLD09 that the authors of~\cite{6496342, 6559603} performed used a total of 16,000 records of normal activity and 10,500 records of attack activity.
+The experiments using only RLD09 that the authors of~\cite{6496342} used a total of 16,000 records of normal activity and 10,500 records of attack activity.
 Of the attack records, 4,000 of them were denial of service attacks and 6,500 were probe attacks.
 
-In the first experiment, the fuzzy genetic algorithm was used to create denial of service and probe detection rules and then the rules were verified with known attack types.
+In the first experiment, the fuzzy genetic algorithm was used to create separate denial of service and probe detection rules.
+Both of the denial of service and probe rules were then used together in the testing process to identify attacks from the testing data set; this is shown in Algorithm~\ref{alg:identifyAttacks}.
 10,000 records were used for the training set and all 26,500 records were used for the testing set.
-The two steps in the training process were to find a denial of service rule, and find a probe rule.
-Both of these rules were then used together in the testing process to identify attacks from the testing data set.
+(The authors of~\cite{6496342} really should not have reused the training data in the testing set because it inflates their results.)
 
+\begin{algorithm}[t]
+\caption{Algorithm that was used to identify attacks in~\cite{6496342}.}
+\label{alg:identifyAttacks}
+\begin{algorithmic}
+\IF{dos\_rule = yes or probe\_rule = yes} 
+	\STATE{This record is an attack;}
+\ELSE \STATE{This record is normal;}
+\ENDIF
+\end{algorithmic}
+\end{algorithm}
+
+\mycomment{Detection rate} To evaluate the accuracy of the fuzzy genetic algorithm, detection rate (DR) was used.
+Detection rate is the percentage of normal and attack activity correctly classified from the total number of data records.
 The detection rate of denial of service attacks in training was 91.64\% and the detection rate of probe attacks in training was 94.79\%.
@@
 \center
-\caption{Results from Experiment 1, using only RLD09}
+\caption{Results from Experiment 1, using only RLD09~\cite{6496342}}
 \vspace{0.20cm}
@@
 
-In the second experiment,~\cite{6496342, 6559603} pulled some types of attacks out of the training set and kept them for unknown data testing.
+In the second experiment,~\cite{6496342} pulled some types of attacks out of the training set and kept them for unknown data testing.
 This was to test that the fuzzy genetic algorithm could detect unknown attacks.
@@
 These three attacks were then used for the testing data set.
-Table~\ref{tab:fuzGenExp2} shows some of the results from the fuzzy genetic algorithm and a decision tree algorithm, which is another common way of addressing these problems.
+Table~\ref{tab:fuzGenExp2} shows the results from the fuzzy genetic algorithm and a decision tree algorithm, which is another common algorithm for classification problems.
 For further information on decision trees, see~\cite{decisionTree}, and for further information on the types of denial of service and probe attacks, see~\cite{typesOfAttacks}.
-When compared with the decision tree algorithm, the fuzzy genetic algorithm has a higher detection rate in all cases except 3 and 5.
+When compared with the decision tree algorithm, the fuzzy genetic algorithm has a higher detection rate in all cases except 5 and 7, however, no information on statistical significance was given.
 
 \begin{table}
-\caption{Unknown Attack Experiment, using only RLD09}
+\caption{Unknown Attack Experiment, using only RLD09~\cite{6496342}}
 \vspace{0.20cm}
@@
 
-2 & UDP Flood (DoS) & Avg = & Avg =\\
+2 & HTTP Flood (DoS) & Avg = & Avg =\\
+  & IP Scan (Probe)		& 88.4 & 95.30\\
+  & Null Scan (Probe)	&		  &\\ \hline
+
+3 & Smurf (DoS) & Avg = & Avg =\\
+  & Port Scan (Probe)		& 97.65 & 99.15\\
+  & Connect Scan (Probe)	&		  &\\ \hline
+
+4 & UDP Flood (DoS) & Avg = & Avg =\\
   & Host Scan (Probe) & 46.65 & 99.80\\
@@
 
-3 & Jping (DoS)    & Avg =          & Avg =\\
+5 & Jping (DoS)    & Avg =          & Avg =\\
   & Syn Scan (Probe) & 99.70 & 98.75\\
@@
 
-4 & UDP Flood (DoS) & Avg = & Avg =\\
+6 & UDP Flood (DoS) & Avg = & Avg =\\
   & RCP Scan (Probe)  & 70.35 & 98.15\\
@@
 
-5 & Http Flood (DoS) & Avg =          & Avg =\\
+7 & Http Flood (DoS) & Avg =          & Avg =\\
   & RCP Scan (Probe)  & 99.94 & 97.50\\
@@
 \subsubsection{Experiments using Both RLD09 and KDD99}
-The authors of~\cite{6496342, 6559603} also ran experiments that used both the RLD09 data set and the KDD99 data set in order to compare how the fuzzy genetic algorithm would perform on both.
-They used the KDD99 10\% version file~\cite{KDD99} for both the training dataset and testing dataset.
+The authors of~\cite{6559603} also ran experiments that used both the RLD09 data set and the KDD99 data set in order to compare how the fuzzy genetic algorithm would perform on both.
+They used a subset of the KDD99 data set for both the training data set and testing data set.
 
 The first experiment used the fuzzy genetic algorithm to classify normal activity and attacks from both data sets.
-The authors of~\cite{6496342, 6559603} first trained and tested the fuzzy genetic algorithm (Algorithm~\ref{alg:fuzGenAlg}) with the KDD99 data set.
+The authors of~\cite{6559603} first trained and tested the fuzzy genetic algorithm (Algorithm~\ref{alg:fuzGenAlg}) with the KDD99 data set.
 There were 6 different types of denial of service attacks and 4 different types of probe attacks.
@@
 \begin{table}
-\caption{KDD99 and RLD09 Results}
+\caption{KDD99 and RLD09 results from the first experiment in~\cite{6559603}.}
 \vspace{0.20cm}
@@
 Ten tests were run, and Table~\ref{tab:kddAttacks} shows the accuracy of detecting some of the cases.
-The results showed that the detection rate of most of the cases were greater than 93\%.
+The results showed that the detection rate of eight out of ten cases were greater than 93\%.
 There were only two cases that had low detection rates, one of which is case 1 in Table~\ref{tab:kddAttacks}.
+The two low detection rates were 16.56\% and 15.58\%, both of which were denial of service attacks.
 
 \begin{table}
-\caption{Results for KDD99 with Certain Attacks}
+\caption{Results for KDD99 with Certain Attacks~\cite{6559603}}
 \vspace{0.20cm}
@@
 Again, there were only two test cases that had low detection rates, one of which is case 2 in Table~\ref{tab:rldAttacks}.
+The two low detection rates were 89.59\% (denial of service), and 86.89\% (probe).
 
 \begin{table}
-\caption{Results for RLD99 with Certain Attacks}
+\caption{Results for RLD09 with Certain Attacks~\cite{6559603}}
 \vspace{0.20cm}
@@
 \end{table}
+
+\section{Using Genetic Algorithms}
+\label{sec:genAlgImp}
+
+\subsection{Algorithm Overview}
+The authors of~\cite{DBLP:journals/corr/abs-1204-1336} used a genetic algorithm to develop an IDS.
+
+\subsection{Experimental Design and Results}
+The authors of~\cite{DBLP:journals/corr/abs-1204-1336} used the KDD99 data set.
+Standard subsets of the KDD99 data set were used for training and testing.
+The training set had a total of 494,021 records, 396,741 of which were attacks.
+The test set had a total of 311,029 records, 250,436 of which were attacks.
+Table~\ref{tab:numberOfRecords} shows the distribution of each type of attack, as well as normal activity, that were in the training and test sets.
+
+\begin{table}
+\center
+\caption{Training and testing records used in~\cite{DBLP:journals/corr/abs-1204-1336}}
+\vspace{0.20cm}
+\begin{tabular}{lll}
+  & Training & Testing \\ 
+Normal & 97,280 & 60,593\\ \hline
+DoS	   & 391,458 & 229,853\\
+R2L    & 1,124   & 16,189\\
+U2R    & 52      & 228\\
+Probe  & 4,107  & 4,166\\ \hline
+Total Attacks & 396,741 & 250,436 \\
+Total Records  & 494,021 & 311,029\\
+\end{tabular}
+\center
+\label{tab:numberOfRecords}
+\end{table}
+
+The results from running the genetic algorithm are shown in Table~\ref{tab:genAlgResults}.
+Denial of service attacks had the highest detection rate at 99.4\%.
+A high detection rate on denial of service attacks isn't surprising because there were a lot of denial of service records used compared with the other types of attacks.
+The accuracy of the genetic algorithm is shown in Table~\ref{tab:genAlgResults2}.
+The true negative rate was 69.5\%.
+The false positive rate was 30.5\%.
+The false negative rate was 5\%.
+The true positive rate was 94.9\%.
+
+\begin{table*}
+\center
+\caption{Results for GA Experiment in~\cite{DBLP:journals/corr/abs-1204-1336}}
+\vspace{0.20cm}
+\begin{tabular}{cc|cccccc}
+& & & & Predicted & & & \\
+& Actual & Normal & Probe & DoS & U2R & R2L & \% Correct \\ \cline{3-8}
+& Normal & 42138 & 1421 & 15835 & 486 & 713 & 69.5\\
+& Probe & 398 & 2963 & 654 & 2 & 149 & 71.1   \\
+& Dos & 921 & 432 & 228489 & 1 & 10 & 99.4\\
+& U2R & 146 & 21 & 8 & 43 & 10 & 18.9\\
+& R2L & 11191 & 578 & 3398 & 141 & 881 & 5.4\\
+&& &&&&& \\
+& \% Correct &  76.9 & 54.7 & 92.0 & 6.4 & 50.0 & \\
+\end{tabular}
+\center
+\label{tab:genAlgResults}
+\end{table*}
+
+%\begin{table}
+%\center
+%\caption{Results for GA Experiment in~\cite{DBLP:journals/corr/abs-1204-1336}}
+%\vspace{0.20cm}
+%\begin{tabular}{c|cccc}
+%& & Predicted & & \\
+%Actual & Normal & Probe & DoS & \% Correct \\ \cline{2-5}
+%Normal & 42138 & 1421 & 15835 & 69.5\\
+%Probe & 398 & 2963 & 654 & 71.1   \\
+%Dos & 921 & 432 & 228489 & 99.4\\
+%&&&& \\
+%\% Correct &  76.9 & 54.7 & 92.0 &\\
+%\end{tabular}
+%\center
+%\label{tab:genAlgResults}
+%\end{table}
+
+\begin{table}
+\center
+\caption{Accuracy of the GA in~\cite{DBLP:journals/corr/abs-1204-1336}}
+\vspace{0.20cm}
+\begin{tabular}{l|ll}
+& & Predicted\\
+Actual & Normal & Attack \\ \cline{2-3}
+Normal & TN: 42,138. 69.5\% & FP: 18,455. 30.5\% \\
+Attack & FN: 12,528. 5\% & TP: 237,908. 94.9\%  \\
+\end{tabular}
+\center
+\label{tab:genAlgResults2}
+\end{table}
+
+The authors of~\cite{DBLP:journals/corr/abs-1204-1336} compared their results (Table~\ref{tab:genAlgResults}) with the winning entry of the KDD99 Classifier Learning Contest~\cite{KDD99Contest}, which can be seen in Table~\ref{tab:winningEntry}.
+The winning entry used a decision tree algorithm.
+The winning entry had a higher detection rate for normal activity, probe attacks, and remote to user attacks.
+The authors of~\cite{DBLP:journals/corr/abs-1204-1336} found that they had a better detection rate for denial of service and user to root attacks than the winning entry.
+For the winning entry, the detection rate of denial of service was 97.1\%, and for user to root it was 13.2\%.
+In~\cite{DBLP:journals/corr/abs-1204-1336} the detection rate of denial of service was 99.4\% and for user to root it was 18.9\%.
+The authors of~\cite{DBLP:journals/corr/abs-1204-1336} provide no information about the statistical significance of these improvements.
+
+\begin{table*}
+\center
+\caption{Results for the Winning Entry of the KDD99 Classifier Learning Contest}
+\vspace{0.20cm}
+\begin{tabular}{c|cccccc}
+& & & & Predicted & & \\
+Actual & Normal & Probe & DoS & U2R & R2L & \% Correct \\ \cline{2-7}
+Normal &  60262 & 243 & 78 & 4 & 6 & 99.5\\
+Probe &  511 & 3471 & 184 & 0 & 0 & 83.3 \\
+Dos &  5299 & 1328 & 223226 & 0 & 0 & 97.1\\
+U2R &  168 & 20 & 0 & 30 & 10 & 13.2\\
+R2L &  14527 & 294 & 0 & 8 & 1360 & 8.4\\
+& &&&&& \\
+\% Correct &  74.6 & 64.8 & 99.9 & 71.4 & 98.8 &\\
+\end{tabular}
+\center
+\label{tab:winningEntry}
+\end{table*}
 

### 2013-11-23 paper.tex
Added some stuff to the results section.

@@
 \section{Introduction}
- \mycomment{WORK ON INTRODUCTION} The CSci computer lab at the University of Minnesota, Morris gets large numbers of login attempts that are attempts at intrusion.
+ \mycomment{work on introduction} The CSci computer lab at the University of Minnesota, Morris gets large numbers of login attempts that are attempts at intrusion.
 An attacker could be trying to gain root access to the system, which would then give them the ability to do things like delete files of other users and change user passwords.
@@
 A SYN error is a failure to get an ACK back.~\cite{TCP}
-  \item same\_srv\_rate: the percentage of connections to the same service. \mycomment{Have example of service}
+  \item same\_srv\_rate: the percentage of connections to the same service.
 \end{enumerate}
@@
 \center
-\caption{Results from Experiment 1, using only RLD09~\cite{6496342}}
+\caption{Results from Experiment 1, using only RLD09~\cite{6496342}.}
 \vspace{0.20cm}
@@
 \begin{table}
-\caption{Unknown Attack Experiment, using only RLD09~\cite{6496342}}
+\caption{Unknown Attack Experiment, using only RLD09~\cite{6496342}.}
 \vspace{0.20cm}
@@
 
-7 & Http Flood (DoS) & Avg =          & Avg =\\
+7 & HTTP Flood (DoS) & Avg =          & Avg =\\
   & RCP Scan (Probe)  & 99.94 & 97.50\\
@@
 \begin{table}
-\caption{Results for KDD99 with Certain Attacks~\cite{6559603}}
+\caption{Results for KDD99 with Certain Attacks. 10 tests were run in total, 5 are shown here.~\cite{6559603}}
 \vspace{0.20cm}
@@
 2 & Smurf & DoS & 0.76 & 0.10 & 99.73\\
-3 & Portsweep & Probe & 6.40 & 0.00 & 93.66\\
-4 & Satan & Probe & 0.74 & 3.75 & 99.22\\
+3 & Neptune & DoS & 0.15 & 0.34 & 99.75\\
+4 & Portsweep & Probe & 6.40 & 0.00 & 93.66\\
+5 & Satan & Probe & 0.74 & 3.75 & 99.22\\
 \end{tabular}
@@
 17 tests were run, and Table~\ref{tab:rldAttacks} shows the accuracy of detecting some of the cases.
-The results showed that the detection rate of a majority of the cases were greater than 97.5\%.
-Again, there were only two test cases that had low detection rates, one of which is case 2 in Table~\ref{tab:rldAttacks}.
+The results showed that the detection rate of 15 out of 17 cases were greater than 97\%.
+Again, there were only two test cases that had low detection rates, cases 3 and 5 in Table~\ref{tab:rldAttacks}.
 The two low detection rates were 89.59\% (denial of service), and 86.89\% (probe).
@@
 \begin{table}
-\caption{Results for RLD09 with Certain Attacks~\cite{6559603}}
+\caption{Results for RLD09 with Certain Attacks. 17 tests were run in total, 6 are shown here.~\cite{6559603}}
 \vspace{0.20cm}
@@
 Test & Attack & Type & FP (\%) & FN (\%) & DR (\%)\\ \hline
-1 & Smurf & DoS & 0.02 & 0 & 99.98\\
-2 & UDP Flood & DoS & 11.06 & 0 & 89.59\\
-3 & Ackscan & Probe & 0.03 & 0 & 99.97\\
-4 & Synscan & Probe & 0.65 & 4.2 & 99.24\\
+1 & HTTP Flood & DoS & 0.36 & 3.5 & 99.46\\
+2 & Smurf & DoS & 0.02 & 0 & 99.98\\
+3 & UDP Flood & DoS & 11.06 & 0 & 89.59\\
+4 & Fin Scan & Probe & 2.58 & 0 & 97.50\\
+5 & IP Scan & Probe & 13.01 & 16.4 & 86.89\\
+6 & Syn Scan & Probe & 0.65 & 4.2 & 99.24\\
 \end{tabular}
@@
 
-\subsection{Algorithm Overview}
 The authors of~\cite{DBLP:journals/corr/abs-1204-1336} used a genetic algorithm to develop an IDS.
+They didn't explain their research that well and left a lot of things out, so I will just be describing their results.
 
@@
 \center
-\caption{Training and testing records used in~\cite{DBLP:journals/corr/abs-1204-1336}}
+\caption{Training and testing records used in~\cite{DBLP:journals/corr/abs-1204-1336}.}
 \vspace{0.20cm}
@@
 \center
-\caption{Results for GA Experiment in~\cite{DBLP:journals/corr/abs-1204-1336}}
+\caption{Results for GA Experiment in~\cite{DBLP:journals/corr/abs-1204-1336}.}
 \vspace{0.20cm}
@@
 \center
-\caption{Accuracy of the GA in~\cite{DBLP:journals/corr/abs-1204-1336}}
+\caption{Accuracy of the GA in~\cite{DBLP:journals/corr/abs-1204-1336}.}
 \vspace{0.20cm}
@@
 The authors of~\cite{DBLP:journals/corr/abs-1204-1336} provide no information about the statistical significance of these improvements.
+\mycomment{compare~\cite{DBLP:journals/corr/abs-1204-1336} with~\cite{6496342, 6559603}}
 
@@
 \center
-\caption{Results for the Winning Entry of the KDD99 Classifier Learning Contest}
+\caption{Results for the Winning Entry of the KDD99 Classifier Learning Contest.}
 \vspace{0.20cm}
@@
 
+\mycomment{fix references}
+
 %\section{Acknowledgments}

### 2013-11-25 paper.tex
Removed comments from bibliography file, changed format of Wikipedia sources.
Made sure that there were no widows or bad line breaks. Also fixed the references section for now. Also compared the fuzzy ga with the traditional ga.

@@
 \usepackage{algorithm}
-%\usepackage{multirow}
+
+\usepackage{natbib}
+\setlength{\bibsep}{3pt}
 
@@
 Intrusion detection systems provide one way of detecting attacks on systems, by monitoring network activities for malicious or abnormal behaviors.
-This paper describes two different ways of training an intrusion detection system about possible attacks to a system: genetic algorithms and fuzzy logic.
-The use of genetic algorithms and fuzzy logic in intrusion detection are effective ways of detecting attacks.
+This paper describes two ways of training an intrusion detection system to recognize possible attacks on a system: genetic algorithms and fuzzy logic.
 I will describe an approach to using fuzzy genetic algorithms and compare those results with results obtained using a decision tree.
-I will then describe the results from using a genetic algorithm and compare those with the winning entry of the KDD99 Classifier Learning Contest.
-\mycomment{work on abstract}
-
+I will then describe the results from using a traditional genetic algorithm and compare those with the winning entry of the KDD99 Classifier Learning Contest, as well as with the fuzzy genetic algorithm.
+These results show that the use of genetic algorithms and fuzzy logic in intrusion detection are effective ways of detecting attacks.
 \end{abstract}
@@
 \section{Introduction}
- \mycomment{work on introduction} The CSci computer lab at the University of Minnesota, Morris gets large numbers of login attempts that are attempts at intrusion.
+The CSci computer lab at the University of Minnesota, Morris gets large numbers of login attempts that are attempts at intrusion.
 An attacker could be trying to gain root access to the system, which would then give them the ability to do things like delete files of other users and change user passwords.
-If the university had an intrusion detection system, it would be possible to classify those attempts into legitimate and illegitimate attempts to login.
+If the university had an intrusion detection system (IDS), it would be possible to classify those attempts into legitimate and illegitimate attempts to login.
 Then it would be possible to block IP addresses that are generating large numbers of attacks.
@@
 An attack on systems is a concern for many people, so it is important to have a way to detect and analyze these attacks.
-Intrusion detection systems provide one way of detecting attacks by monitoring network activities for malicious or abnormal behaviors and then producing reports and alerts~\cite{DBLP:journals/corr/abs-1204-1336}.
-There are various ways of training the intrusion detection system about possible threats.
+Intrusion detection systems provide one way of detecting attacks by monitoring network activities for malicious or abnormal behaviors and then producing reports, alerts, and actions~\cite{DBLP:journals/corr/abs-1204-1336}.
+There are various ways of training an IDS about possible threats.
 Two approaches that I will talk about in this paper are genetic algorithms and fuzzy logic.
 
-In Section~\ref{sec:background} I will give some background on intrusion detection and genetic algorithms.
+In Section~\ref{sec:background} I will give some background on intrusion detection.
 I will describe the main types of networking attacks in Section~\ref{sec:typesAttacks}, and ways of detecting attacks in Section~\ref{sec:detectionMeth}.
 In Section~\ref{sec:dataSets} I will describe two data sets that are commonly used in intrusion detection.
-After that, I will describe the measures that are used in determining the accuracy of an intrusion detection system in Section~\ref{sec:PosNeg}.
-I will then give information on genetic algorithms in Section~\ref{sec:GA}.
-In Section~\ref{sec:fuzGenAlgImp} I will describe the use of a fuzzy genetic algorithm in an intrusion detection system, and in Section~\ref{sec:genAlgImp} I will describe the results of using a genetic algorithm.
-Section~\ref{sec:conclusion} has some conclusions about the use of genetic algorithms and fuzzy genetic algorithms in intrusion detection systems.
+After that,  in Section~\ref{sec:PosNeg} I will describe the measures that are used in determining the accuracy of an IDS.
+I will then provide an overview on genetic algorithms in Section~\ref{sec:GA}.
+In Section~\ref{sec:fuzGenAlgImp} I will describe the use of a fuzzy genetic algorithm in an IDS, and in Section~\ref{sec:genAlgImp} I will describe the results of using a traditional genetic algorithm.
+Section~\ref{sec:conclusion} provides some conclusions about the use of genetic algorithms and fuzzy genetic algorithms in intrusion detection systems.
 
@@
 \begin{enumerate}
+\setlength{\itemsep}{1pt}
   \item duration: the length of the record in seconds.
@@
   \item num\_failed\_logins: the number of failed login attempts in this record.
-  \item root\_shell: returns 1 if root shell is obtained, which means that the user is able to log in as root. If root shell is not obtained, 0 is returned.
+  \item root\_shell: returns 1 if root shell is obtained, else it returns 0.
   \item num\_access\_files: the number of operations on access control files.
@@
 RLD09 uses 12 features.
-For further information about the features, see~\cite{6496342, 6559603}.
+For further information about the features, see~\cite{6496342}.
 
-In the research discussed in Section~\ref{sec:genAlgImp}, the authors of ~\cite{DBLP:journals/corr/abs-1204-1336} used only the numerical features of the KDD99 data set (34 out of 41 total features).
+In the research discussed in Section~\ref{sec:genAlgImp}, the authors of ~\cite{DBLP:journals/corr/abs-1204-1336} used only the numerical features of KDD99 (34 out of 41).
+
+In a machine learning experiment, a common technique is to divide the data set into two subsets, a \emph{training set} and a \emph{testing set}.
+The given algorithm is then trained on the training set to look for patterns.
+These patterns are then verified using the test set.~\cite{bc1_ecindm}
 
@@
 \label{sec:PosNeg}
-In a machine learning experiment, a common technique is to divide the data set into two subsets, a \emph{training set} and a \emph{testing set}.
-The given algorithm is then trained on the training set to look for patterns.
-These patterns are then verified using the test set.~\cite{bc1_ecindm} Four different measures are then used to determine the accuracy of the algorithm in question.
-
+Four measures are used to determine the accuracy of an algorithm.
 The \emph{false negative} (FN) rate is the percentage that attacks are misclassified from the total number of attack records.
@@
 
-%A \emph{false positive} (FP) happens when an IDS incorrectly identifies normal activity as being an attack.
-A \emph{false negative} (FN) happens when an IDS fails to identify harmful activity.
-A \emph{true positive} (TP) happens when an IDS correctly identifies activities to be attacks.
-A \emph{true negative} (TN) happens when an IDS correctly identifies activities to be normal.~\cite{ids}
-
-\subsection{Genetic Algorithms}
+\section{Genetic Algorithms}
 \label{sec:GA}
-\mycomment{Work on this section} Genetic algorithms (GA) are a search technique used to find solutions to problems.
-Operations analogous to biological \emph{mutation}, \emph{selection}, and \emph{crossover} are used to evolve and improve rules.
+Genetic algorithms (GAs) are a search technique used to find solutions to problems.
+Operations analogous to biological \emph{mutation}, \emph{selection}, and \emph{crossover} are used to evolve and improve solutions.
 
 Possible solutions to problems can be represented in a variety of problem dependent ways, such as bit strings.
-First, a randomly generated population of rules is created.
+For example, in an IDS rules are represented as bit strings, and then they are used to classify an event as an attack or normal traffic, see Section~\ref{sec:fuzzyRulesAndAlgorithm} for more information.
+First, a randomly generated population of potential solutions is created.
 Then mutation, selection, and crossover are applied to each generation until an acceptable solution is found or some time limit is exceeded.
@@
 Mutation is where random bits in an \emph{individual}, or possible solution, are changed.
-Selection is where individuals that have a better \emph{fitness} are chosen over the other individuals.
-The fitness function determines the quality of a particular individual.
-For example, in intrusion detection systems the fitness measures the performance of a rule.
-Crossover is where two individuals swap sequences of bits to form two new individuals.~\cite{DBLP:journals/corr/abs-1204-6416}
+Crossover is where two individuals swap sequences of bits to form two new individuals.
+For example, in an IDS mutation takes the bits of a rule and changes them.
+Crossover takes two rules and creates new rules by swapping the bits of the old rules.
+
+Selection is where individuals that have better \emph{fitness} are chosen to be parents.
+The fitness of an individual is specified by the \emph{fitness function}, which determines the quality of a particular individual.
+For example, in an IDS the fitness measures the performance of a rule.
+Selection combined with a fitness function directs the search towards an effective solution.~\cite{DBLP:journals/corr/abs-1204-6416}
 
@@
 \subsection{Fuzzy Logic Rules and Algorithm}
+\label{sec:fuzzyRulesAndAlgorithm}
 Attacks on systems do not always have a fixed pattern, so fuzzy logic is used to detect patterns that have a behavior that is between normal and unusual.
 Fuzzy logic rules are similar to the rules described in Section~\ref{sec:detectionMeth}, except that \emph{consequence} is a certainty factor.
-For example, if (\emph{duration} = 6) then (\emph{probability of it being an attack is 50\%}).
-\mycomment{certainty vs probability}
+For example, if (\emph{duration} = 6) then (\emph{the degree of certainty of it being an attack is 0.5}).
 
-In order to measure the probability that a record is an attack, a trapezoidal shape was used in~\cite{6496342, 6559603}; this is shown in Figure~\ref{fig:trapFigure}.
+To find the degree of certainty of a record being an attack, a trapezoidal shape was used in~\cite{6496342, 6559603}; this is shown in Figure~\ref{fig:trapFigure}.
 The trapezoidal shape has four parameters: $a, b, c$, and $d$.
-Algorithm~\ref{alg:fuzAlg} calculates the probability of a record being an attack.
+Algorithm~\ref{alg:fuzAlg} calculates the certainty of a record being an attack.
 
 \begin{figure}
+\caption{Illustration of fuzzy rule used to find the degree of certainty of a record being an attack. Also see Algorithm~\ref{alg:fuzAlg}. This is used in~\cite{6496342, 6559603}.}
 \centering
-\psfig{file=TrapezoidFigure.pdf, width=3.3in}
-\caption{Illustration of fuzzy rule used to measure the probability that a record is an attack. Also see Algorithm~\ref{alg:fuzAlg}. This is used in~\cite{6496342, 6559603}.}
+\psfig{file=trapFigExample.pdf, width=3.3in}
 \label{fig:trapFigure}
@@
 \IF{$b \leq \textrm{data} \leq c$}
-  \STATE{prob = 1.0}
+  \STATE{certainty = 1.0}
 \ELSIF{$a < \textrm{data} < b$}
   \STATE{
-    prob = $(\textrm{data}-a)/(b-a)$
+    certainty = $(\textrm{data}-a)/(b-a)$
   }
@@
   \STATE{
-    prob = $(d-\textrm{data})/(d-c)$
+    certainty = $(d-\textrm{data})/(d-c)$
   }
-\ELSE \STATE{prob = 0.0}
+\ELSE \STATE{certainty = 0.0}
 \ENDIF
@@
 Suppose that $a$ = 1, $b$ = 3, $c$ = 5, $d$ = 7.
-Because data (6) is between $c$ and $d$, then the probability of an attack is equal to
+Because data (6) is between $c$ and $d$, then the degree of certainty of it being an attack is equal to
 
@@
 
+\newpage
+
 The four parameters $a, b, c$, and $d$ are encoded into blocks of binary strings, where each block is a feature with values between 0 and 7. 
-See Figure~\ref{fig:fuzEncodingForFeature} for an example of a block.
+%See Figure~\ref{fig:fuzEncodingForFeature} for an example of a block. 
 A rule has one block for each of 12 features followed at the end by a marker indicating the type of attack.
 An example of this is shown in Figure~\ref{fig:rule}.
-The authors of~\cite{6496342, 6559603} compute the probability for each of the 12 blocks, and if the sum of those is greater than a threshold, then it will be declared as an attack.
+The authors of~\cite{6496342, 6559603} compute the degree of certainty for each of the 12 blocks, and if the sum of those is greater than a threshold, then it will be declared as an attack.
 
-\begin{figure}
-\centering
-\caption{Fuzzy encoding for a feature that is used in~\cite{6496342, 6559603}.}
-\vspace{0.20cm}
-\begin{tabular}{|cccc|} \hline
-010 & 011 & 100 & 101\\
-a=2 & b=3 & c=4 & d=5\\
-\hline\end{tabular}
-\label{fig:fuzEncodingForFeature}
-\end{figure}
+%\begin{figure}
+%\centering
+%\caption{Fuzzy encoding for a feature that is used in~\cite{6496342, 6559603}.}
+%\vspace{0.20cm}
+%\begin{tabular}{|cccc|} \hline
+%010 & 011 & 100 & 101\\
+%a=2 & b=3 & c=4 & d=5\\
+%\hline\end{tabular}
+%\label{fig:fuzEncodingForFeature}
+%\end{figure}
 
@@
 Each feature in a record is matched to one block of the rule.
-The parameters of each block measure the probability of an attack using the trapezoidal fuzzy rule shape.
-The sum of the probabilities from each block are then compared with a threshold to determine if the record represents an attack or normal behavior.
+The parameters of each block measure the degree of certainty of an attack using the trapezoidal fuzzy rule shape.
+The sum of the degrees of certainty from each block are then compared with a threshold to determine if the record represents an attack or normal behavior.
 
 \begin{algorithm}[t]
-\caption{Fuzzy Genetic Algorithm that is based on an algorithm used in~\cite{6496342, 6559603}, but a few corrections have been made.}
+\caption{Fuzzy GA that is based on an algorithm used in~\cite{6496342, 6559603}, but a few corrections have been made.}
 \label{alg:fuzGenAlg}
@@
     \FOR{each feature}
-      \STATE{prob = fuzzy(); // Algorithm~\ref{alg:fuzAlg}}
-      \STATE{totalprob = totalprob + prob;}
+      \STATE{certainty = fuzzy(); // Algorithm~\ref{alg:fuzAlg}}
+      \STATE{total = total + certainty;}
     \ENDFOR
 
-    \IF{totalprob > threshold} 
+    \IF{total > threshold} 
       \STATE{class is attack;}
@@
 
+\newpage
+
 \subsection{Experimental Design and Results}
+\label{sec:experimentalDesignResults}
 A variety of experiments were run in~\cite{6496342, 6559603}. Two experiments used just RLD09, and three experiments used KDD99 and RLD09 together.
@@
 \subsubsection{Experiments using Only RLD09}
-The experiments using only RLD09 that the authors of~\cite{6496342} used a total of 16,000 records of normal activity and 10,500 records of attack activity.
-Of the attack records, 4,000 of them were denial of service attacks and 6,500 were probe attacks.
+The experiments using only RLD09 that~\cite{6496342} performed used a total of 16,000 records of normal activity and 10,500 records of attack activity.
+Of the attack records, 4,000 were denial of service attacks and 6,500 were probe attacks.
 
@@
 \begin{algorithm}[t]
-\caption{Algorithm that was used to identify attacks in~\cite{6496342}.}
+\caption{This algorithm was used to identify attacks and normal activity in~\cite{6496342}.}
 \label{alg:identifyAttacks}
@@
 
-\mycomment{Detection rate} To evaluate the accuracy of the fuzzy genetic algorithm, detection rate (DR) was used.
-Detection rate is the percentage of normal and attack activity correctly classified from the total number of data records.
-The detection rate of denial of service attacks in training was 91.64\% and the detection rate of probe attacks in training was 94.79\%.
+To evaluate the accuracy of the fuzzy genetic algorithm, detection rate (DR) was used.
+Detection rate is defined in~\cite{6496342, 6559603} as the percentage of normal and attack activity correctly classified from the total number of data records.
+The detection rate of DoS attacks in training was 91.64\% and the detection rate of probe attacks in training was 94.79\%.
 The detection rate of the testing data set increased to 97.92\%.
-Results from this experiment are shown in Table~\ref{tab:fuzGenExp1}.
+Results from this experiment are in Table~\ref{tab:fuzGenExp1}.
 
@@
 \begin{table}
-\caption{Unknown Attack Experiment, using only RLD09~\cite{6496342}.}
+\caption{Unknown attack experiment, using only the RLD09 data set~\cite{6496342}.}
 \vspace{0.20cm}
 \begin{tabular}{llll}
-Test & Unknown & Decision & Fuzzy\\
-Case & Attacks & Tree DR (\%)  & Genetic DR (\%)\\ \hline
+Test & Unknown & Decision & Fuzzy GA\\
+Case & Attacks & Tree DR (\%)  & DR (\%)\\ \hline
 
@@
 \vspace{0.20cm}
-\begin{tabular}{cccccc}
-Data set & Attack & Normal & FP (\%) & FN (\%) & DR (\%)\\ \hline
+\begin{tabular}{llllll}
+Data set & Attack & Normal & FP(\%) & FN(\%) & DR(\%)\\ \hline
 KDD99 & 160,117 & 39,337 & 0.13 & 1.55 & 98.72\\
@@
 \begin{tabular}{cccccc}
-Test & Attack & Type & FP (\%) & FN (\%) & DR (\%)\\ \hline
+Test & Attack & Type & FP(\%) & FN(\%) & DR(\%)\\ \hline
 1 & Back & DoS & 85.33 & 0.00 & 16.56\\
@@
 \begin{tabular}{cccccc}
-Test & Attack & Type & FP (\%) & FN (\%) & DR (\%)\\ \hline
+Test & Attack & Type & FP(\%) & FN(\%) & DR(\%)\\ \hline
 1 & HTTP Flood & DoS & 0.36 & 3.5 & 99.46\\
@@
 
-\subsection{Experimental Design and Results}
-The authors of~\cite{DBLP:journals/corr/abs-1204-1336} used the KDD99 data set.
-Standard subsets of the KDD99 data set were used for training and testing.
+The KDD99 data set was used in the experiments.
+Standard subsets of set were used for training and testing.
 The training set had a total of 494,021 records, 396,741 of which were attacks.
@@
 The results from running the genetic algorithm are shown in Table~\ref{tab:genAlgResults}.
+Detection rate is defined in~\cite{DBLP:journals/corr/abs-1204-1336} as the ratio between the number of correctly detected intrusions and the total number of intrusions.
 Denial of service attacks had the highest detection rate at 99.4\%.
@@
 \vspace{0.20cm}
-\begin{tabular}{cc|cccccc}
-& & & & Predicted & & & \\
-& Actual & Normal & Probe & DoS & U2R & R2L & \% Correct \\ \cline{3-8}
-& Normal & 42138 & 1421 & 15835 & 486 & 713 & 69.5\\
-& Probe & 398 & 2963 & 654 & 2 & 149 & 71.1   \\
-& Dos & 921 & 432 & 228489 & 1 & 10 & 99.4\\
-& U2R & 146 & 21 & 8 & 43 & 10 & 18.9\\
-& R2L & 11191 & 578 & 3398 & 141 & 881 & 5.4\\
-&& &&&&& \\
-& \% Correct &  76.9 & 54.7 & 92.0 & 6.4 & 50.0 & \\
+\begin{tabular}{l|cccccc}
+& & & Predicted & & & \\
+Actual & Normal & Probe & DoS & U2R & R2L & \% Correct \\ \hline
+Normal & 42138 & 1421 & 15835 & 486 & 713 & 69.5\\
+Probe & 398 & 2963 & 654 & 2 & 149 & 71.1   \\
+Dos & 921 & 432 & 228489 & 1 & 10 & 99.4\\
+U2R & 146 & 21 & 8 & 43 & 10 & 18.9\\
+R2L & 11191 & 578 & 3398 & 141 & 881 & 5.4\\
+& &&&&& \\
+\% Correct &  76.9 & 54.7 & 92.0 & 6.4 & 50.0 & \\
 \end{tabular}
@@
 
-%\begin{table}
-%\center
-%\caption{Results for GA Experiment in~\cite{DBLP:journals/corr/abs-1204-1336}}
-%\vspace{0.20cm}
-%\begin{tabular}{c|cccc}
-%& & Predicted & & \\
-%Actual & Normal & Probe & DoS & \% Correct \\ \cline{2-5}
-%Normal & 42138 & 1421 & 15835 & 69.5\\
-%Probe & 398 & 2963 & 654 & 71.1   \\
-%Dos & 921 & 432 & 228489 & 99.4\\
-%&&&& \\
-%\% Correct &  76.9 & 54.7 & 92.0 &\\
-%\end{tabular}
-%\center
-%\label{tab:genAlgResults}
-%\end{table}
-
 \begin{table}
@@
 \vspace{0.20cm}
+Predicted
 \begin{tabular}{l|ll}
-& & Predicted\\
-Actual & Normal & Attack \\ \cline{2-3}
+Actual & Normal & Attack \\ \hline
 Normal & TN: 42,138. 69.5\% & FP: 18,455. 30.5\% \\
@@
 The authors of~\cite{DBLP:journals/corr/abs-1204-1336} provide no information about the statistical significance of these improvements.
-\mycomment{compare~\cite{DBLP:journals/corr/abs-1204-1336} with~\cite{6496342, 6559603}}
 
@@
 \vspace{0.20cm}
-\begin{tabular}{c|cccccc}
-& & & & Predicted & & \\
-Actual & Normal & Probe & DoS & U2R & R2L & \% Correct \\ \cline{2-7}
+\begin{tabular}{l|cccccc}
+& & & Predicted & & &\\
+Actual & Normal & Probe & DoS & U2R & R2L & \% Correct \\ \hline
 Normal &  60262 & 243 & 78 & 4 & 6 & 99.5\\
@@
 
+The detection rate (as defined in Section~\ref{sec:experimentalDesignResults} and~\cite{6559603}) of the traditional genetic algorithm in~\cite{DBLP:journals/corr/abs-1204-1336} was 90\%, which means that it correctly classified 90\% of the test records.
+The detection rate of the fuzzy genetic algorithm in~\cite{6559603} was 99\%.
+I did this comparison and do not know the statistical significance.
+The system in~\cite{6559603} is not the same as the system in~\cite{DBLP:journals/corr/abs-1204-1336} with fuzzy logic added to it, but it can be said that the system in~\cite{6559603} is better than~\cite{DBLP:journals/corr/abs-1204-1336} however it was built.
+
 \section{Conclusions}
 \label{sec:conclusion}
-This paper showed that the use of genetic algorithms and fuzzy logic in intrusion detection are effective ways of detecting attacks.
-The genetic algorithm that was used in~\cite{DBLP:journals/corr/abs-1204-1336} had a high detection rate for denial of service attacks.
+The fuzzy genetic algorithm that was used in~\cite{6496342, 6559603} had a higher detection rate than a decision tree algorithm in most cases, and it was good at detecting unknown attacks.
+It had a higher detection rate than the traditional genetic algorithm that was used in~\cite{DBLP:journals/corr/abs-1204-1336}.
+The genetic algorithm in~\cite{DBLP:journals/corr/abs-1204-1336} had a high detection rate for denial of service attacks.
 When compared with the winning entry of the KDD99 Classifier Learning Contest, it was shown to have a better detection rate for both denial of service and user to root attacks.
-The fuzzy genetic algorithm that was used in~\cite{6496342, 6559603} had a higher detection rate than a decision tree algorithm in most cases.
-It was also shown that fuzzy genetic algorithms are good at detecting unknown attacks.
-
-\mycomment{fix references}
+This paper showed that the use of genetic algorithms and fuzzy genetic algorithms in intrusion detection are effective ways of detecting attacks.
 

### 2013-11-26 paper.tex
Changed a few small things based on feedback from Nic.

@@
 For further information about the features, see~\cite{6496342}.
-
-In the research discussed in Section~\ref{sec:genAlgImp}, the authors of ~\cite{DBLP:journals/corr/abs-1204-1336} used only the numerical features of KDD99 (34 out of 41).
+In the research discussed in Section~\ref{sec:genAlgImp}, the authors of~\cite{DBLP:journals/corr/abs-1204-1336} used only the numerical features of KDD99 (34 out of 41).
 
@@
 The given algorithm is then trained on the training set to look for patterns.
-These patterns are then verified using the test set.~\cite{bc1_ecindm}
+These patterns are then verified using the test set.~\cite{bc1_ecindm} This approach is used in~\cite{6496342, 6559603} and~\cite{DBLP:journals/corr/abs-1204-1336}.
 
@@
 Possible solutions to problems can be represented in a variety of problem dependent ways, such as bit strings.
-For example, in an IDS rules are represented as bit strings, and then they are used to classify an event as an attack or normal traffic, see Section~\ref{sec:fuzzyRulesAndAlgorithm} for more information.
+For example, IDS rules can be represented as bit strings; see Section~\ref{sec:fuzzyRulesAndAlgorithm} for more information.
 First, a randomly generated population of potential solutions is created.
-Then mutation, selection, and crossover are applied to each generation until an acceptable solution is found or some time limit is exceeded.
+Then mutation, crossover, and selection are applied to each generation until an acceptable solution is found or some time limit is exceeded.
 
-Mutation is where random bits in an \emph{individual}, or possible solution, are changed.
+Mutation is where random bits in an \emph{individual}, or possible solution, are randomly changed.
 Crossover is where two individuals swap sequences of bits to form two new individuals.
-For example, in an IDS mutation takes the bits of a rule and changes them.
+For example, in an IDS mutation takes the bits of a rule and changes them to form a slightly different rule.
 Crossover takes two rules and creates new rules by swapping the bits of the old rules.
@@
 The fitness of an individual is specified by the \emph{fitness function}, which determines the quality of a particular individual.
-For example, in an IDS the fitness measures the performance of a rule.
+For example, in an IDS the fitness measures how well a rule classifies records as either attacks or normal activity.
 Selection combined with a fitness function directs the search towards an effective solution.~\cite{DBLP:journals/corr/abs-1204-6416}
@@
 \label{sec:fuzGenAlgImp}
-The IDS used in~\cite{6496342, 6559603} is able to identify normal network activity as well as attacks using a fuzzy genetic algorithm. This kind of algorithm is able to learn new attacks.
+The IDS used in~\cite{6496342, 6559603} is able to identify normal network activity as well as attacks using a fuzzy genetic algorithm.
+This kind of algorithm is able to learn to recognize new attacks.
 
@@
 \end{equation*}
-
-\newpage
 
@@
 
-The detection rate (as defined in Section~\ref{sec:experimentalDesignResults} and~\cite{6559603}) of the traditional genetic algorithm in~\cite{DBLP:journals/corr/abs-1204-1336} was 90\%, which means that it correctly classified 90\% of the test records.
-The detection rate of the fuzzy genetic algorithm in~\cite{6559603} was 99\%.
-I did this comparison and do not know the statistical significance.
-The system in~\cite{6559603} is not the same as the system in~\cite{DBLP:journals/corr/abs-1204-1336} with fuzzy logic added to it, but it can be said that the system in~\cite{6559603} is better than~\cite{DBLP:journals/corr/abs-1204-1336} however it was built.
+The detection rate (as defined in Section~\ref{sec:experimentalDesignResults} and~\cite{6559603}) of the traditional genetic algorithm in~\cite{DBLP:journals/corr/abs-1204-1336} was 90\% (see Tables~\ref{tab:genAlgResults2} and~\ref{tab:numberOfRecords}).
+This means that it correctly classified 90\% of the test records.
+The detection rate of the fuzzy genetic algorithm in~\cite{6559603} was 99\% (see Table~\ref{tab:bothSetsResults}).
+The fuzzy genetic algorithm in~\cite{6559603} is not the same as the traditional genetic algorithm in~\cite{DBLP:journals/corr/abs-1204-1336} with fuzzy logic added to it, but it can be said that the system in~\cite{6559603} is better than~\cite{DBLP:journals/corr/abs-1204-1336} however it was built.
 

### 2013-11-26 Talk/Presentation.tex
Changed a few small things and also removed the GA section for now.

@@
 \usepackage{algorithmic}
-\usepackage{multirow}
 
@@
   	\item 
-	\item 
-	\item 
+	\item have story here,
+	\item also explain root somewhere
 	\item 
@@
 
-\subsection{Detection Methodologies}
+\subsection{Detection Methodologies and Rules}
 \begin{frame}
@@
   \begin{itemize}
-  	\item Signature-based detection: compares well-known patterns of attacks that are already in the intrusion detection system against captured events in order to identify a possible attack.
+  	\item Signature-based detection: compares well-known patterns of attacks
+that are already known to the IDS against captured events in order to identify a possible attack.
+	\begin{itemize}
+		\item It is a simple and effective way to detect known attacks, but is ineffective against new kinds of unknown attacks.
+	\end{itemize}
   	\item Anomaly-based detection: looks for patterns of activity that are rare and uncommon.
+	\begin{itemize}
+		\item It is harder to do than signature-based detection, but it can be an effective way to detect new, unknown attacks.
+	\end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Rules}
+	\begin{itemize}
+		\item A commonly used approach for detecting intrusions is to use rules.
+		\item If-Then format: If (\emph{condition}) then (\emph{consequence}).
+		\begin{itemize}
+			\item The condition is composed of one or more features, and the consequence says if it is an intrusion or not.
+			\item If \emph{duration = 4} then \emph{intrusion}.
+		\end{itemize}				
 	\end{itemize}
@@
 	\begin{itemize}
-		\item duration: length of the normal or attack activity
+		\item duration: length of the record in seconds.
 in seconds.
@@
 	\begin{itemize}
-		\item RLD09 was created because KDD99 is 14 years old.
+		\item RLD09 was created because KDD99 is 14 years old, and newer attack types are not included in it because of its age.
 		\item Data was captured from a university in Bangkok, Thailand.
-		\item 17 different types of attacks (divided into denial of service and probe attacks), and 12 features.
+		\item 17 different types of attacks (divided into denial of service and probe attacks).
+
+\linespace
+\linespace
+\linespace
+\linespace
+		\item .............maybe have something like "the authors of [6] used only the numerical features of KDD99 (34 out of 41)." if I decide to explain that paper.
 	\end{itemize}
@@
 
-\subsection{Rules}
 \begin{frame}
-  \frametitle{Rules}
+  \frametitle{Training and Testing Sets}
 	\begin{itemize}
-		\item A commonly used approach for detecting intrusions and to differentiate between normal connections and attacks is to use rules.
-		\item If-Then format: If (\emph{condition}) then (\emph{consequence}).
+		\item A common technique is to divide the data set into two subsets, a \emph{training set} and a \emph{testing set}.
+        \item The given algorithm is then trained on the training set to look for patterns.
+        \item These patterns are then verified using the test set.
+	\end{itemize}
+\end{frame}
+
+\subsection{Determining the Accuracy of an Algorithm}
+\begin{frame}
+  \frametitle{Determining the Accuracy of an Algorithm}
+\centering{Predicted}
+\begin{table}
+Actual
+\begin{tabular}{l|ll}
+ & Not Attack & Attack \\ \hline
+Not Attack & True Negative (TN) & False Positive (FP) \\
+Attack & False Negative (FN) & True Positive (FP)  \\
+\end{tabular}
+\end{table}
+\linespace
 	\begin{itemize}
-			\item The condition is composed of one or more features, and the consequence says if it is an intrusion or not.
-			\item If \emph{duration = 4} then \emph{intrusion}.
-		\end{itemize}				
+		\item Detection rate (DR): the percentage of normal and attack activity correctly classified from the total number of data records.
+
+		\linespace
+		\linespace
+		\linespace
+		\linespace
+		\linespace
+		\item .........maybe talk about the other paper's DR?
 	\end{itemize}
@@
 	\begin{itemize}
-		\item Search technique used to find solutions to problems.
-		\item Mutation: random bits in an individual, or possible solution, are changed.
-		\item Selection: individuals that have a better fitness are chosen over the other individuals.
-		\item Fitness function: determines the quality of a particular individual.
-		\item Crossover: Two individuals swap one of their characteristics with the other to form two new individuals.
-	\end{itemize}
-\end{frame}
-
-\subsection{Determining the Accuracy of an Algorithm}
-\begin{frame}
-  \frametitle{Determining the Accuracy of an Algorithm}
-\begin{table}
-\begin{tabular}{cccc}
-& & \multicolumn{2}{ c }{Predicted} \\
-& & Not Attack & Attack \\ \cline{3-4}
-\multicolumn{1}{ c }{\multirow{2}{*}{Actual} } &
-\multicolumn{1}{ c| }{Not Attack} & True Negative (TN) & False Positive (FP)\\
-\multicolumn{1}{ c }{} &
-\multicolumn{1}{ c| }{Attack} & False Negative (FN) & True Positive (TP)\\
-\end{tabular}
-\end{table}
-\linespace
-	\begin{itemize}
-		\item Detection rate (DR): 
+		\item GAs are a search technique used to find solutions to problems. Operations analogous to biological mutation, selection, and crossover are used to evolve and improve solutions.
+        \item Possible solutions to problems can be represented in a variety of problem dependent ways, such as bit strings.
+        \item First, a randomly generated population of potential solutions is created.
+Then mutation, crossover, and selection are applied to each generation until an acceptable solution is found or some time limit is exceeded.
+        \item Mutation: random bits in an individual, or possible solution, are randomly changed.
+        \item Crossover: Two individuals swap sequences of bits to form two new individuals.
+        \item Selection: individuals that have better fitness are chosen to be parents.
+        \item The fitness of an individual is specified by the fitness function, which determines the quality of a particular individual.
 	\end{itemize}
@@
 		\begin{itemize}
-			\item If (\emph{duration} = 6) then (\emph{probability of it being an attack is 50\%}).
+			\item If (\emph{duration} = 6) then (\emph{the degree of certainty of it being an attack is 0.5}).
 		\end{itemize}
@@
 \begin{frame}
-  \frametitle{Measuring the Probability of a Record Being an Attack}
-	\begin{itemize}	
-	\item Feature: duration, 6 seconds.
-	\end{itemize}
+  \frametitle{Finding the Degree of Certainty of a Record Being an Attack}
+	%\begin{itemize}	
+	%\item Suppose that the feature is duration, and it is 6 seconds, so then data = 6.
+	%\end{itemize}
+	Suppose that the feature is duration, and it is 6 seconds. Then data=6.
 %  \begin{columns}
@@
 	\begin{itemize}
-	\item The four parameters are encoded into blocks.
-	\item Each block is a feature with values between 0.0 and 7.0.
+	\item The four parameters are encoded into blocks. Each block is a feature with values between 0 and 7.
 	\end{itemize}
@@
 	\begin{itemize}
-	\item A rule has 12 blocks of features, at the end is the type of attack.
+	\item A rule has one block for each of 12 features followed at the end by a marker indicating the type of attack.
 	\end{itemize}
@@
 
+	\begin{itemize}
+		\item The degree of certainty is computed for each of the 12 blocks, and if the sum of those is greater than a threshold, then it will be declared as an attack.
+	\end{itemize}
+
 \end{frame}
@@
 	\begin{itemize}
-		\item The algorithm generates rules, improves rules, then rules are used to classify the data.
+		\item The algorithm generates rules and improves rules.
 		\item One record is passed into a rule.
 		\item Each feature in a record is matched to one block of the rule.
-		\item The parameters of each block measure the probability of an attack using the trapezoidal fuzzy rule shape.
-		\item The probabilities of each block are then compared with a threshold to determine if the record represents an attack or normal behavior.
+		\item The parameters of each block measure the degree of certainty of an attack using the trapezoidal fuzzy rule shape.
+		\item The sum of the degrees of certainty from each block are then compared with a threshold to determine if the record represents an attack or normal behavior.
 	\end{itemize}
@@
 \begin{algorithmic}
+\FOR{each rule}
   \FOR{each record}
-  \FOR{each rule}
     \FOR{each feature}
@@
 \begin{column}{0.4\textwidth}
-	Fitness function:
+	The fitness function in the algorithm is:
 	\begin{equation*}
@@
 	\begin{itemize}
-		\item Fuzzy GA was used to create DoS and probe detection rules, then the rules were verified with known attack types.
-		\item Two steps in the training process: find a DoS rule, find a probe rule. Both of these rules were then used together in the testing process to identify attacks from the testing data set.
+		\item Fuzzy GA was used to create DoS and probe detection rules. Both of the DoS and probe rules were then used together in the testing process to identify attacks from the testing data set.
 		\item 10,000 records were used for the training set, 26,500 records were used for the test set.
 	\end{itemize}
+
+\linespace
+\begin{itemize}
+	\item The following algorithm was used to identify attacks and normal activity in this experiment.
+\end{itemize}
+
+\begin{algorithmic}
+\IF{dos\_rule = yes or probe\_rule = yes}
+	\STATE{This record is an attack;}
+\ELSE \STATE{This record is normal;}
+\ENDIF
+\end{algorithmic}
 \end{frame}
@@
 	\frametitle{Experiments Using Only RLD09}
-	Experiment 2 Results
+	Experiment 2 Results (7 tests were run in total, 5 are shown here.)
 
@@
 		\item 158,597 records of DoS attacks. 1,500 records of probe attacks.
+		\item 10 tests were run in total, 5 are shown here.
 	\end{itemize}
@@
 		\item 6,400 records of DoS attacks. 10,400 records of probe attacks.
+		\item 17 tests were run in total, 6 are shown here.
 	\end{itemize}
@@
 Test & Attack & Type & FP (\%) & FN (\%) & DR (\%)\\ \hline
-1 & Smurf & DoS & 0.02 & 0 & 99.98\\
-2 & UDP Flood & DoS & 11.06 & 0 & 89.59\\
-3 & Ackscan & Probe & 0.03 & 0 & 99.97\\
-4 & Synscan & Probe & 0.65 & 4.2 & 99.24\\
+1 & HTTP Flood & DoS & 0.36 & 3.5 & 99.46\\
+2 & Smurf & DoS & 0.02 & 0 & 99.98\\
+3 & UDP Flood & DoS & 11.06 & 0 & 89.59\\
+4 & Fin Scan & Probe & 2.58 & 0 & 97.50\\
+5 & IP Scan & Probe & 13.01 & 16.4 & 86.89\\
+6 & Syn Scan & Probe & 0.65 & 4.2 & 99.24\\
 \end{tabular}
 \end{table}
-
-\end{frame}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section[Using Genetic Algorithms]{Using Genetic Algorithms}
-\subsection{Experimental Design and Results}
-\begin{frame}
-  \frametitle{Experimental Design}
-  \begin{columns}
-
-  \begin{column}{0.5\textwidth}
-	\begin{itemize}
-		\item KDD99 data set.
-		\item Used only the numerical features of KDD99 (34 out of 41 total features).
-		\item Training set: 494,021 records (396,741 of them are attacks). Test set: 311,029 records (250,436 of them are attacks).
-	\end{itemize}
-  \end{column}
-
-  \begin{column}{0.5\textwidth}
-  \begin{table}
-\caption{ }
-\begin{tabular}{lll}
-  & Training & Testing \\
-Normal & 97,280 & 60,593\\
-Probe & 4,107 & 4,166\\
-DoS         & 391,458 & 229,853\\
-U2R & 52 & 228\\
-R2L & 1,124 & 16,189\\
-Total & 494,021 & 311,029\\
-\end{tabular}
-\end{table}
-  \end{column}
-  \end{columns}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Results}
-%\begin{table}
-%\begin{small}
-%\begin{tabular}{cccccccc}
-%& & \multicolumn{5}{ c }{Predicted} \\
-%& & Normal & Probe & DoS & U2R & R2L & \% Correct \\ \cline{3-8}
-%\multicolumn{1}{ c }{\multirow{2}{*}{Actual} } &
-%\multicolumn{1}{ c| }{Normal} & 42138 & 1421 & 15835 & 486 & 713 & 69.5\\
-%\multicolumn{1}{ c }{} &
-%\multicolumn{1}{ c| }{Probe} & 398 & 2963 & 654 & 2 & 149 & 71.1 \\
-%\multicolumn{1}{ c }{\multirow{2}{*}{} } &
-%\multicolumn{1}{ c| }{Dos} & 921 & 432 & 228489 & 1 & 10 & 99.4\\
-%\multicolumn{1}{ c }{} &
-%\multicolumn{1}{ c| }{U2R} & 146 & 21 & 8 & 43 & 10 & 18.9\\
-%\multicolumn{1}{ c }{} &
-%\multicolumn{1}{ c| }{R2L} & 11191 & 578 & 3398 & 141 & 881 & 5.4\\
-
-%\multicolumn{1}{ c }{} &
-%\multicolumn{1}{ c| }{} & & & & & & \\
-
-%\multicolumn{1}{ c }{\% Correct} &
-%\multicolumn{1}{ c| }{} & 76.9 & 54.7 & 92.0 & 6.4 & 50.0 \\
-%\end{tabular}
-%\end{small}
-%\end{table}
-
-%	\begin{figure}
-%  \includegraphics[width=0.95\textwidth]{../resultsForGA.JPG}
-%  \caption{From [Hoque-Mukit-Bikas, 2012]}
-%  	\end{figure}
 
@@
 \frametitle{Conclusions}
-
 	\begin{itemize}
-  \item The use of genetic algorithms and fuzzy logic in intrusion detection are effective ways of detecting attacks.
-%  \item The genetic algorithm had a high detection rate for denial of service attacks.
 		\item The fuzzy genetic algorithm had a higher detection rate than a decision tree algorithm in most cases.
 		\item Fuzzy genetic algorithms are good at detecting unknown attacks.
+		\item The use of fuzzy genetic algorithms in intrusion detection is an effective way of detecting attacks.
 	\end{itemize}
-
 \end{frame}

### 2013-11-26 Talk/Presentation.tex
Split the GA section into two slides, changed some things on the detection methodologies slide, also added some things to The Big Picture slide.

@@
   \frametitle{The Big Picture}
-
-  \begin{columns}
-  \begin{column}{0.6\textwidth}
   \begin{itemize}
-  	\item 
-	\item have story here,
-	\item also explain root somewhere
-	\item 
-	\item 
+  	\item Computer lab gets large numbers of login attempts that are attempts at intrusion.
+	\item An attack could be trying to gain root access to the system.
+	\item With an intrusion detection system (IDS), it would be possible to classify those attempts into legitimate and illegitimate attempts to login.
+	\item Intrusion detection systems provide one way of detecting attacks by monitoring network activities for malicious or abnormal behaviors and then producing reports, alerts, and actions.
+	\item A way of training an IDS about possible threats is by using a fuzzy genetic algorithm.
   \end{itemize}
-  \end{column}
-  \begin{column}{0.4\textwidth}
-
-  \end{column}
-  \end{columns}
 \end{frame}
@@
 	\begin{itemize}
-		\item It is a simple and effective way to detect known attacks, but is ineffective against new kinds of unknown attacks.
+		\item Simple and effective way to detect known attacks, but is ineffective against new kinds of unknown attacks.
 	\end{itemize}
+
+\linespace
+\linespace
+
   	\item Anomaly-based detection: looks for patterns of activity that are rare and uncommon.
 	\begin{itemize}
-		\item It is harder to do than signature-based detection, but it can be an effective way to detect new, unknown attacks.
+		\item Harder to do than signature-based detection, but it can be an effective way to detect new, unknown attacks.
 	\end{itemize}
@@
 		\item Has long been a standard data set for intrusion detection.
-		\item Data was processed into 5 million records.
+
+\linespace
+
+		\item Data was processed into 5 million \emph{records}.
 			\begin{itemize}
@@
 		\item Data in the set is classified as normal or attack activity.
-		\item Uses 41 features, which are properties of a record that are used to describe the activity and help to distinguish normal connections from attacks.
+
+\linespace
+
+		\item Uses 41 \emph{features}, which are properties of a record that are used to describe the activity and help to distinguish normal connections from attacks.
 
@@
 	\begin{itemize}
-		\item RLD09 was created because KDD99 is 14 years old, and newer attack types are not included in it because of its age.
+		\item RLD09 was created because KDD99 is 14 years old, and newer attack types are not included in KDD99 because of its age.
 		\item Data was captured from a university in Bangkok, Thailand.
 		\item 17 different types of attacks (divided into denial of service and probe attacks).
-
-\linespace
-\linespace
-\linespace
-\linespace
-		\item .............maybe have something like "the authors of [6] used only the numerical features of KDD99 (34 out of 41)." if I decide to explain that paper.
 	\end{itemize}
@@
 		\item Detection rate (DR): the percentage of normal and attack activity correctly classified from the total number of data records.
-
-		\linespace
-		\linespace
-		\linespace
-		\linespace
-		\linespace
-		\item .........maybe talk about the other paper's DR?
 	\end{itemize}
@@
 	\begin{itemize}
-		\item GAs are a search technique used to find solutions to problems. Operations analogous to biological mutation, selection, and crossover are used to evolve and improve solutions.
+		\item GAs are a search technique used to find solutions to problems.
         \item Possible solutions to problems can be represented in a variety of problem dependent ways, such as bit strings.
@@
 Then mutation, crossover, and selection are applied to each generation until an acceptable solution is found or some time limit is exceeded.
+	\end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Genetic Algorithms}
+	\begin{itemize}
         \item Mutation: random bits in an individual, or possible solution, are randomly changed.
-        \item Crossover: Two individuals swap sequences of bits to form two new individuals.
+        \item Crossover: two individuals swap sequences of bits to form two new individuals.
         \item Selection: individuals that have better fitness are chosen to be parents.
@@
 	\begin{itemize}
-		\item The algorithm generates rules and improves rules.
 		\item One record is passed into a rule.
@@
 	\begin{itemize}
-		\item A variety of experiments were run. Two experiments used just RLD09. Three experiments used both the RLD09 and KDD99 in order to compare how the fuzzy GA would perform on both.
+		\item A variety of experiments were run. Two experiments used just RLD09. Three experiments used both RLD09 and KDD99 in order to compare how the fuzzy GA would perform on both.
 	\end{itemize}
@@
 	\begin{itemize}
-		\item Fuzzy GA was used to create DoS and probe detection rules. Both of the DoS and probe rules were then used together in the testing process to identify attacks from the testing data set.
-		\item 10,000 records were used for the training set, 26,500 records were used for the test set.
+		\item Fuzzy GA was used to create DoS and probe detection rules. Both the DoS and probe rules were then used together in the testing process to identify attacks from the testing data set.
+		\item 10,000 records in the training set, 26,500 records in the test set.
 	\end{itemize}
@@
 
-	\bibitem{DBLP:journals/corr/abs-1204-1336}
-Mohammad Sazzadul Hoque, Md. Abdul Mukit, Md. Abu Naser Bikas.
-\newblock An Implementation of Intrusion Detection System Using Genetic Algorithm.
-\newblock \emph{CoRR}, abs/1204.1336, 2012.
+%	\bibitem{DBLP:journals/corr/abs-1204-1336}
+%Mohammad Sazzadul Hoque, Md. Abdul Mukit, Md. Abu Naser Bikas.
+%\newblock An Implementation of Intrusion Detection System Using Genetic Algorithm.
+%\newblock \emph{CoRR}, abs/1204.1336, 2012.
 	\end{small}

### 2013-11-27 paper.tex
Made a few minor changes, like getting rid of unnecessary commas.

@@
 \begin{abstract}
-Intrusion detection systems provide one way of detecting attacks on systems, by monitoring network activities for malicious or abnormal behaviors.
+Intrusion detection systems provide one way of detecting attacks on systems by monitoring network activities for malicious or abnormal behaviors.
 This paper describes two ways of training an intrusion detection system to recognize possible attacks on a system: genetic algorithms and fuzzy logic.
@@
 \label{sec:typesAttacks}
-There are four main types of networking attacks that this paper will address: \emph{denial of service}, \emph{remote to user attacks}, \emph{user to root attacks}, and \emph{probe attacks}.
+There are four main types of networking attacks that this paper will address: \emph{denial of service attacks}, \emph{remote to user attacks}, \emph{user to root attacks}, and \emph{probe attacks}.
 Each attack that happens on a network can be placed into one of these categories.~\cite{DBLP:journals/corr/abs-1204-1336}
@@
 User to root (U2R) attacks happen when an attacker starts out with access on the machine and then tries to gain root access to the system.
-If an attacker has access to an account on the machine, then they could potentially run programs that take advantage of operating system weaknesses to gain root access.
+If an attacker has access to an account on the machine, they could potentially run programs that take advantage of operating system weaknesses to gain root access.
 Probe attacks happen when an attacker examines a machine in order to collect information about weaknesses or vulnerabilities that in the future could be used to compromise the system.
@@
 Fuzzy logic rules are similar to the rules described in Section~\ref{sec:detectionMeth}, except that \emph{consequence} is a certainty factor.
-For example, if (\emph{duration} = 6) then (\emph{the degree of certainty of it being an attack is 0.5}).
+For example, if (\emph{duration} = 6) then (\emph{the degree of certainty of the record being an attack is 0.5}).
 
@@
 An example of using the fuzzy logic approach is: suppose that the feature is duration, and suppose it is 6 seconds, so then data = 6.
-Suppose that $a$ = 1, $b$ = 3, $c$ = 5, $d$ = 7.
+Suppose that $a$ = 1, $b$ = 3, $c$ = 5, $d$~=~7.
 Because data (6) is between $c$ and $d$, then the degree of certainty of it being an attack is equal to

### 2013-12-02 Talk/Presentation.tex
Changed some things on some of the Experiments/results slides. Also added a few examples to some of the slides.

@@
 
-  	\linespace
-  	\linespace
+  	\begin{itemize}
+  		\item Many systems lock out a user from an account after a certain number of failed login attempts.
+An attacker would be able to use this to prevent legitimate users from logging in, by intentionally failing to log in enough times to lock the account.
+  	\end{itemize}
 
-  	\item Remote to User (R2L): attacker tries to gain access to things a local user would have on the machine.
+%  	\linespace
+%  	\linespace
 
-  	\linespace
-  	\linespace
+%  	\item Remote to User (R2L): attacker tries to gain access to things a local user would have on the machine.
 
-  	\item User to Root (U2R): attacker starts out with access on the machine and then tries to gain root access to the system.
+%  	\linespace
+%  	\linespace
+
+%  	\item User to Root (U2R): attacker starts out with access on the machine and then tries to gain root access to the system.
 
@@
   	\item Probe: attacker examines a machine in order to collect information about weaknesses or vulnerabilities that in the future could be used to compromise the system.
+
+  	\begin{itemize}
+  		\item An attack could be trying to determine what version of a software is being run on that machine, and if that version has a known issue then that allows them to attempt to attack that.
+  	\end{itemize}
   \end{itemize}
@@
 			\end{itemize}
-		\item Data in the set is classified as normal or attack activity.
+		\item Each record is classified as either normal or attack activity.
 
@@
 
-		\item Uses 41 \emph{features}, which are properties of a record that are used to describe the activity and help to distinguish normal connections from attacks.
+		\item KDD99 uses 41 \emph{features}, which are properties of a record that are used to describe the activity and help to distinguish normal connections from attacks.
 
@@
 		\item duration: length of the record in seconds.
-in seconds.
 		%\item src\_bytes: number of bytes sent from source to destination.
@@
 		\item Data was captured from a university in Bangkok, Thailand.
-		\item 17 different types of attacks (divided into denial of service and probe attacks).
+		\item As well as normal network activity, RLD09 has 17 different types of attacks (divided into denial of service and probe attacks).
 	\end{itemize}
@@
 	\begin{itemize}
-		\item A common technique is to divide the data set into two subsets, a \emph{training set} and a \emph{testing set}.
+		\item In a machine learning experiment, a common technique is to divide the data set into two subsets, a \emph{training set} and a \emph{testing set}.
         \item The given algorithm is then trained on the training set to look for patterns.
@@
         \item Possible solutions to problems can be represented in a variety of problem dependent ways, such as bit strings.
+
+        \begin{itemize}
+        	\item IDS rules can be represented as bit strings.
+        \end{itemize}
+
+        \linespace
         \item First, a randomly generated population of potential solutions is created.
@@
         \item Crossover: two individuals swap sequences of bits to form two new individuals.
+
+        \linespace
+        \linespace
+
+       	\item In an IDS: mutation takes the bits of a rule and changes them to form a slightly different rule. Crossover takes 2 rules and creates new rules by swapping the bits of the old rules.
+
+	\end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Genetic Algorithms}
+	\begin{itemize}
         \item Selection: individuals that have better fitness are chosen to be parents.
         \item The fitness of an individual is specified by the fitness function, which determines the quality of a particular individual.
+
+        \linespace
+        \linespace
+
+       	\item In an IDS: fitness measures how well a rule classifies records as either attacks or normal activity.
+Selection combined with a fitness function directs the search towards an effective solution.
 	\end{itemize}
@@
 		\begin{itemize}
-			\item If (\emph{duration} = 6) then (\emph{the degree of certainty of it being an attack is 0.5}).
+			\item If (\emph{duration} = 6) then (\emph{the degree of certainty of the record being an attack is 0.5}).
 		\end{itemize}
@@
 	\frametitle{Algorithm}
-\begin{columns}
-\begin{column}{0.6\textwidth}
-\begin{small}
-\begin{algorithmic}
-\FOR{each rule}
-  \FOR{each record}
-    \FOR{each feature}
-      \STATE{prob = fuzzy();}
-      \STATE{totalprob = totalprob + prob;}
-    \ENDFOR    
-    \IF{totalprob > threshold} 
-      \STATE{class is attack;}
-      \ELSE \STATE{class is normal;}
-    \ENDIF
-  \ENDFOR
-  \STATE{find $A$, $B$, $\alpha$, and $\beta$  %// $A$: \# of attack records.
-$B$: \# of normal records.
-$\alpha$: \# of attack records correctly identified as attack.
-$\beta$: \# of normal records incorrectly classified as attack.}
+%\begin{columns}
+%\begin{column}{0.6\textwidth}
+%\begin{small}
+%\begin{algorithmic}
+%\FOR{each rule}
+%  \FOR{each record}
+%    \FOR{each feature}
+%      \STATE{prob = fuzzy();}
+%      \STATE{totalprob = totalprob + prob;}
+%    \ENDFOR    
+%    \IF{totalprob > threshold} 
+%      \STATE{class is attack;}
+%      \ELSE \STATE{class is normal;}
+%    \ENDIF
+%  \ENDFOR
+%  \STATE{find $A$, $B$, $\alpha$, and $\beta$} 
+%\ENDFOR
+%\STATE{calculate fitness}
+%\STATE{crossover(), mutation()}
 
-  } 
-\ENDFOR
-\STATE{calculate fitness}
-\STATE{crossover(), mutation()}
+%\end{algorithmic}
+%\end{small}
+%\end{column}
 
-\end{algorithmic}
-\end{small}
-\end{column}
-
-\begin{column}{0.4\textwidth}
+%\begin{column}{0.4\textwidth}
 	The fitness function in the algorithm is:
@@
 
+	$\alpha$: \# of attack records correctly identified as attack.
+
 	$A$: \# of attack records.
+
+	$\beta$: \# of normal records incorrectly classified as attack.
 
@@
 
-	$\alpha$: \# of attack records correctly identified as attack.
-
-	$\beta$: \# of normal records incorrectly classified as attack.
-
-\end{column}
-\end{columns}
+%\end{column}
+%\end{columns}
 \end{frame}
@@
 
-\linespace
 \begin{itemize}
@@
 		\item Attacks were pulled out of the training set and kept for unknown data testing. This was to test that the fuzzy GA could detect unknown attacks.
-		\item Used fuzzy GA and a decision tree algorithm.
-		\item For each test case there were 13 attack types plus normal activity that were in the training data set. Three attack types were used for the unknown testing data set.
+		\item Used fuzzy GA and a decision tree algorithm, which is another common algorithm for classification problems.
+		\item 7 tests were run, and for each test case there were 13 attack types plus normal activity that were in the training data set. Three attack types were used for the unknown testing data set.
 	\end{itemize}
@@
 	\frametitle{Experiments Using Both RLD09 and KDD99}
+Three experiments used both RLD09 and KDD99.
+
+\linespace
+\linespace
+
 Experiment 1
@@
 1 & Back & DoS & 85.33 & 0.00 & 16.56\\
-2 & Smurf & DoS & 0.76 & 0.10 & 99.73\\
-3 & Neptune & DoS & 0.15 & 0.34 & 99.75\\
+2 & PoD & DoS & 84.66 & 0.00 & 15.58\\
+3 & Smurf & DoS & 0.76 & 0.10 & 99.73\\
 4 & Portsweep & Probe & 6.40 & 0.00 & 93.66\\
@@
 \end{table}
+
+\begin{itemize}
+	\item 8 test cases had DR greater than 93\%. Only 2 cases had low DR.
+\end{itemize}
 
@@
 \end{table}
+
+\begin{itemize}
+	\item 15 cases had DR greater than 97\%. 2 cases had low DR, (cases 3 and 5).
+\end{itemize}
 

### 2013-12-02 Talk/Presentation.tex
Changed the theme, changed the determining the accuracy of an algorithm slide and the encoding of features and rules slide.

@@
 {
-  \usetheme{CambridgeUS}
+  \usetheme{Frankfurt}
   \setbeamercovered{transparent}
+  %\usecolortheme{whale}
 }
@@
   \frametitle{Determining the Accuracy of an Algorithm}
-\centering{Predicted}
+%\centering{Predicted}
 \begin{table}
-Actual
+%Actual
+Predicted
 \begin{tabular}{l|ll}
- & Not Attack & Attack \\ \hline
+Actual   & Not Attack & Attack \\ \hline
 Not Attack & True Negative (TN) & False Positive (FP) \\
@@
 	\item The four parameters are encoded into blocks. Each block is a feature with values between 0 and 7.
-	\end{itemize}
+%	\end{itemize}
 
-\begin{figure}
-\begin{tabular}{|cccc|} \hline
-010 & 011 & 100 & 101\\
-a=2 & b=3 & c=4 & d=5\\
-\hline\end{tabular}
-\end{figure}
+%\begin{figure}
+%\begin{tabular}{|cccc|} \hline
+%010 & 011 & 100 & 101\\
+%a=2 & b=3 & c=4 & d=5\\
+%\hline\end{tabular}
+%\end{figure}
 
-	\begin{itemize}
+%	\begin{itemize}
 	\item A rule has one block for each of 12 features followed at the end by a marker indicating the type of attack.
@@
 \hline\end{tabular}
-\caption{Based on [Jongsuebsuk-Wattanapongsakorn-Charnsripinyo, 2013]}
+\caption{Based on [Jongsuebsuk \emph{et al.}, 2013]}
 \end{small}

### 2013-12-03 Talk/Presentation.tex
Put the fuzzy rule stuff before the GA stuff. Also got rid of some text on some of the slides.

@@
   	\item Computer lab gets large numbers of login attempts that are attempts at intrusion.
-	\item An attack could be trying to gain root access to the system.
-	\item With an intrusion detection system (IDS), it would be possible to classify those attempts into legitimate and illegitimate attempts to login.
+	\begin{itemize}
+		\item Trying to gain root access to the system.
+		\item Delete files of other users and change user passwords.
+	\end{itemize}
+	\linespace
+	\item With an intrusion detection system (IDS) -> classify those attempts.
 	\item Intrusion detection systems provide one way of detecting attacks by monitoring network activities for malicious or abnormal behaviors and then producing reports, alerts, and actions.
-	\item A way of training an IDS about possible threats is by using a fuzzy genetic algorithm.
+	\item Training an IDS: use a fuzzy genetic algorithm.
   \end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section[Background]{Background}
+\section[Intrusion Detection]{Intrusion Detection}
 \subsection{Types of Networking Attacks}
@@
   \begin{itemize}
-  	\item Denial of Service (DoS): attacker makes a machine inaccessible to a user by making it too busy to serve legitimate requests.
+  	\item Denial of Service (DoS):
+
+  	makes machine inaccessible to user by making it too busy to serve legitimate requests.
 
   	\begin{itemize}
-  		\item Many systems lock out a user from an account after a certain number of failed login attempts.
-An attacker would be able to use this to prevent legitimate users from logging in, by intentionally failing to log in enough times to lock the account.
+  		\item  Systems lock out user from account after certain number of failed login attempts.
+Attacker would be able to use this to prevent users from logging in, by failing to log in enough times to lock the account.
   	\end{itemize}
-
-%  	\linespace
-%  	\linespace
-
-%  	\item Remote to User (R2L): attacker tries to gain access to things a local user would have on the machine.
-
-%  	\linespace
-%  	\linespace
-
-%  	\item User to Root (U2R): attacker starts out with access on the machine and then tries to gain root access to the system.
 
@@
 
-  	\item Probe: attacker examines a machine in order to collect information about weaknesses or vulnerabilities that in the future could be used to compromise the system.
+  	\item Probe:
+
+  	examines machine in order to collect information about weaknesses that in the future could be used to compromise the system.
 
   	\begin{itemize}
-  		\item An attack could be trying to determine what version of a software is being run on that machine, and if that version has a known issue then that allows them to attempt to attack that.
+  		\item Attacker could be trying to determine what version of a software is being run on that machine, and if that version has a known issue then that allows them to attempt to attack that.
   	\end{itemize}
@@
 
-\subsection{Detection Methodologies and Rules}
+\subsection{Detection Methodologies}
 \begin{frame}
@@
 	\begin{itemize}
-		\item Simple and effective way to detect known attacks, but is ineffective against new kinds of unknown attacks.
+		\item Simple and effective way to detect known attacks,
+
+		ineffective against new kinds of unknown attacks.
 	\end{itemize}
@@
 	\begin{itemize}
-		\item Harder to do than signature-based detection, but it can be an effective way to detect new, unknown attacks.
-	\end{itemize}
-  \end{itemize}
-\end{frame}
+		\item Harder to do than signature-based detection,
 
-\begin{frame}
-  \frametitle{Rules}
-	\begin{itemize}
-		\item A commonly used approach for detecting intrusions is to use rules.
-		\item If-Then format: If (\emph{condition}) then (\emph{consequence}).
-		\begin{itemize}
-			\item The condition is composed of one or more features, and the consequence says if it is an intrusion or not.
-			\item If \emph{duration = 4} then \emph{intrusion}.
+		can be an effective way to detect new, unknown attacks.
 	\end{itemize}
@@
 		\item Each record is classified as either normal or attack activity.
-
-\linespace
-
-		\item KDD99 uses 41 \emph{features}, which are properties of a record that are used to describe the activity and help to distinguish normal connections from attacks.
-
 	\end{itemize}
@@
 \begin{frame}
-  \frametitle{Some Features of KDD99}
+  \frametitle{Features of KDD99}
 	\begin{itemize}
+		\item KDD99 uses 41 \emph{features}, which are properties of a record that are used to describe the activity and help to distinguish normal connections from attacks.
+
+		\linespace
+		\linespace
+
 		\item duration: length of the record in seconds.
-		%\item src\_bytes: number of bytes sent from source to destination.
 		\item num\_failed\_logins: number of failed login attempts.
 		\item root\_shell: returns 1 if root shell is obtained, else returns 0.
-		%\item num\_access\_files: number of operations on access control files.
-		%\item srv\_count: number of connections to the same service as the current connection in the past two seconds.
-		%\item serror\_rate: percentage of connections that have ``SYN" errors.
-		%\item same\_srv\_rate: percentage of connections to the same service.
 	\end{itemize}
@@
 	\begin{itemize}
-		\item RLD09 was created because KDD99 is 14 years old, and newer attack types are not included in KDD99 because of its age.
+		\item RLD09 was created because KDD99 is 14 years old,
+
+		newer attack types are not in KDD99 because of its age.
 		\item Data was captured from a university in Bangkok, Thailand.
-		\item As well as normal network activity, RLD09 has 17 different types of attacks (divided into denial of service and probe attacks).
+		\item As well as normal network activity, has 17 different types of attacks (divided into denial of service and probe attacks).
+	\end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Rules}
+	\begin{itemize}
+		\item A commonly used approach for detecting intrusions is to use rules.
+		\item If-Then format: If (\emph{condition}) then (\emph{consequence}).
+		\begin{itemize}
+			\item The condition is composed of one or more features, and the consequence says if it is an intrusion or not.
+			\item If \emph{duration = 4} then \emph{intrusion}.
+		\end{itemize}				
 	\end{itemize}
@@
 	\begin{itemize}
-		\item In a machine learning experiment, a common technique is to divide the data set into two subsets, a \emph{training set} and a \emph{testing set}.
+		\item The authors divided the data set into two subsets, a \emph{training set} and a \emph{test set}.
         \item The given algorithm is then trained on the training set to look for patterns.
@@
 Not Attack & True Negative (TN) & False Positive (FP) \\
-Attack & False Negative (FN) & True Positive (FP)  \\
+Attack & False Negative (FN) & True Positive (TP)  \\
 \end{tabular}
@@
 \linespace
-	\begin{itemize}
-		\item Detection rate (DR): the percentage of normal and attack activity correctly classified from the total number of data records.
-	\end{itemize}
-\end{frame}
-
-\subsection{Genetic Algorithms}
-\begin{frame}
-  \frametitle{Genetic Algorithms}
-	\begin{itemize}
-		\item GAs are a search technique used to find solutions to problems.
-        \item Possible solutions to problems can be represented in a variety of problem dependent ways, such as bit strings.
-
-        \begin{itemize}
-        	\item IDS rules can be represented as bit strings.
-        \end{itemize}
-
 \linespace
-        \item First, a randomly generated population of potential solutions is created.
-Then mutation, crossover, and selection are applied to each generation until an acceptable solution is found or some time limit is exceeded.
-	\end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Genetic Algorithms}
-	\begin{itemize}
-        \item Mutation: random bits in an individual, or possible solution, are randomly changed.
-        \item Crossover: two individuals swap sequences of bits to form two new individuals.
-
-        \linespace
-        \linespace
-
-       	\item In an IDS: mutation takes the bits of a rule and changes them to form a slightly different rule. Crossover takes 2 rules and creates new rules by swapping the bits of the old rules.
-
-	\end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Genetic Algorithms}
-	\begin{itemize}
-        \item Selection: individuals that have better fitness are chosen to be parents.
-        \item The fitness of an individual is specified by the fitness function, which determines the quality of a particular individual.
-
-        \linespace
-        \linespace
-
-       	\item In an IDS: fitness measures how well a rule classifies records as either attacks or normal activity.
-Selection combined with a fitness function directs the search towards an effective solution.
-	\end{itemize}
+\begin{center}
+Detection rate (DR): percentage of normal and attack activity correctly classified from the total number of data records.
+\end{center}
 \end{frame}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section[Using Fuzzy Genetic Algorithms]{Using Fuzzy Genetic Algorithms}
-
-\subsection{Fuzzy Algorithm}
-
+\section[Fuzzy Algorithm]{Fuzzy Algorithm}
+\subsection{Fuzzy Logic}
 \begin{frame}
@@
 	\begin{itemize}
-		\item Attacks on systems do not always have a fixed pattern, so fuzzy logic is used to detect patterns that have a behavior that is between normal and unusual.
+		\item Fuzzy logic is used in intrusion detection systems to find the degree of certainty of a record being an attack.
 		\item Fuzzy logic rules are similar to the rules described before, except that
@@
 
+\subsection{Finding the Degree of Certainty}
 \begin{frame}
   \frametitle{Finding the Degree of Certainty of a Record Being an Attack}
-	%\begin{itemize}	
-	%\item Suppose that the feature is duration, and it is 6 seconds, so then data = 6.
-	%\end{itemize}
 	Suppose that the feature is duration, and it is 6 seconds. Then data=6.
-%  \begin{columns}
-%  \begin{column}{0.6\textwidth}
   \includegraphics[width=0.95\textwidth]{../trapFigExample.pdf}
-%  \end{column}
-
-%  \begin{column}{0.4\textwidth}
-%  Example:
-%  \begin{itemize}
-%  	\item Feature: duration (length of the activity in seconds).
-%  	\item a=1, b=3, c=5, d=7
-%  	\item The length of the activity is 6 seconds (between c and d).
-%  	\linespace
-%  	\item prob = $\frac{d-\textrm{data}}{d-c}$ = $\frac{7-6}{7-5}$ = 0.5
-%  \end{itemize}
-
-%  \end{column}
-%  \end{columns}
 \end{frame}
 
+\subsection{Encoding of Features and Rules}
 \begin{frame}
@@
 	\item The four parameters are encoded into blocks. Each block is a feature with values between 0 and 7.
-%	\end{itemize}
-
-%\begin{figure}
-%\begin{tabular}{|cccc|} \hline
-%010 & 011 & 100 & 101\\
-%a=2 & b=3 & c=4 & d=5\\
-%\hline\end{tabular}
-%\end{figure}
-
-%	\begin{itemize}
 	\item A rule has one block for each of 12 features followed at the end by a marker indicating the type of attack.
@@
 \begin{tabular}{|cccc|c|cccc|c|} \hline
-010 & 011 & 100 & 101   & ...... & 010 & 011 & 101 & 111   & DoS\\
-a=2 & b=3 & c=4 & d=5   & ...... & a=2 & b=3 & c=5 & d=7   &\\ 
+010 & 011 & 100 & 101   & ... & 010 & 011 & 101 & 111   & Attack\\
+a=2 & b=3 & c=4 & d=5   & ... & a=2 & b=3 & c=5 & d=7   &\\ 
     &     & Block 1&    &        &     & Block 12& &       & Type\\
@@
 \end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section[Genetic Algorithms]{Genetic Algorithms}
+\subsection{GA Overview}
+\begin{frame}
+  \frametitle{Genetic Algorithms}
+	\begin{itemize}
+		\item GAs: search technique used to find solutions to problems.
+        \item Possible solutions to problems can be represented in a variety of problem dependent ways, such as bit strings.
 
-\subsection{Algorithm Overview}
+        \begin{itemize}
+        	\item IDS rules can be represented as bit strings.
+        \end{itemize}
 
+        \linespace
+        \item First, randomly generated population of potential solutions is created.
+Mutation, crossover, selection are applied to each generation until acceptable solution is found or some time limit is exceeded.
+	\end{itemize}
+\end{frame}
+
+\subsection{Mutation and Crossover}
 \begin{frame}
-	\frametitle{Algorithm Overview}
+  \frametitle{Mutation and Crossover}
 	\begin{itemize}
-		\item One record is passed into a rule.
-		\item Each feature in a record is matched to one block of the rule.
-		\item The parameters of each block measure the degree of certainty of an attack using the trapezoidal fuzzy rule shape.
-		\item The sum of the degrees of certainty from each block are then compared with a threshold to determine if the record represents an attack or normal behavior.
+        \item Mutation: random bits in an individual, or possible solution, are randomly changed.
+        \begin{itemize}
+        	\item Takes bits of rule and changes them to form slightly different rule.
+
+        	\begin{figure}
+			\begin{tabular}{|cccc|} \hline
+			010 & 011 & 100 & \textbf{101}\\
+			a=2 & b=3 & c=4 & \textbf{d=5}\\
+			\hline\end{tabular}
+			$\rightarrow$
+			\quad
+			\begin{tabular}{|cccc|} \hline
+			010 & 011 & 100 & \textbf{111}\\
+			a=2 & b=3 & c=4 & \textbf{d=7}\\
+			\hline\end{tabular}			
+			\end{figure}
+        \end{itemize}
+
+        \linespace
+        \linespace
+        \linespace
+
+        \item Crossover: two individuals swap sequences of bits to form two new individuals.
+        \begin{itemize}
+        	\item Take 2 rules and creates new rules by swapping bits of old rules.
+
+        	\begin{figure}
+			\begin{tabular}{|cccc|} \hline
+			001 & 011 & 101 & 111\\
+			a=1 & b=3 & c=5 & d=7\\
+			\hline\end{tabular}
+			\quad
+			\begin{tabular}{|cccc|} \hline
+			010 & 100 & 110 & 111\\
+			a=2 & b=4 & c=6 & d=7\\
+			\hline\end{tabular}			
+			\end{figure}
+        \end{itemize}
+	\end{itemize}
+\end{frame}
+
+\subsection{Selection and Fitness}
+\begin{frame}
+  \frametitle{Selection and Fitness}
+	\begin{itemize}
+        \item Selection: individuals that have better fitness are chosen to be parents.
+        \item The fitness of an individual is specified by the fitness function, which determines the quality of a particular individual.
+
+        \linespace
+        \linespace
+
+       	\item In an IDS: fitness measures how well a rule classifies records as either attacks or normal activity. Selection combined with fitness function directs search towards effective solution.
 	\end{itemize}
@@
 \begin{frame}
-	\frametitle{Algorithm}
-%\begin{columns}
-%\begin{column}{0.6\textwidth}
-%\begin{small}
-%\begin{algorithmic}
-%\FOR{each rule}
-%  \FOR{each record}
-%    \FOR{each feature}
-%      \STATE{prob = fuzzy();}
-%      \STATE{totalprob = totalprob + prob;}
-%    \ENDFOR    
-%    \IF{totalprob > threshold} 
-%      \STATE{class is attack;}
-%      \ELSE \STATE{class is normal;}
-%    \ENDIF
-%  \ENDFOR
-%  \STATE{find $A$, $B$, $\alpha$, and $\beta$} 
-%\ENDFOR
-%\STATE{calculate fitness}
-%\STATE{crossover(), mutation()}
-
-%\end{algorithmic}
-%\end{small}
-%\end{column}
-
-%\begin{column}{0.4\textwidth}
+	\frametitle{Fitness function}
 	The fitness function in the algorithm is:
@@
 	$B$: \# of normal records.
-
-%\end{column}
-%\end{columns}
 \end{frame}
-
-\subsection{Experimental Design and Results}
-
-\begin{frame}
-	\frametitle{Experiments}
-	\begin{itemize}
-		\item A variety of experiments were run. Two experiments used just RLD09. Three experiments used both RLD09 and KDD99 in order to compare how the fuzzy GA would perform on both.
-	\end{itemize}
-\end{frame}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section[Experiments and Results]{Experiments and Results}
+\subsection{Two Experiments using Only RLD09}
 \begin{frame}
@@
 	\begin{itemize}
-		\item Fuzzy GA was used to create DoS and probe detection rules. Both the DoS and probe rules were then used together in the testing process to identify attacks from the testing data set.
-		\item 10,000 records in the training set, 26,500 records in the test set.
+		\item Fuzzy GA was used to create DoS and probe detection rules.
+
+		Both rules were then used together in testing process to identify attacks from testing data set.
+
+		\linespace
+		\linespace
+
+		\item Training set: 10,000 records.
+
+		Test set: 26,500 records.
+
+		\linespace
+		\linespace
+
+		\item If the record is either a DoS rule or a Probe rule, it is classified as an attack; else it is normal.
 \end{itemize}
-
-\begin{itemize}
-	\item The following algorithm was used to identify attacks and normal activity in this experiment.
-\end{itemize}
-
-\begin{algorithmic}
-\IF{dos\_rule = yes or probe\_rule = yes}
-	\STATE{This record is an attack;}
-\ELSE \STATE{This record is normal;}
-\ENDIF
-\end{algorithmic}
 \end{frame}
@@
 	\begin{itemize}
-		\item Attacks were pulled out of the training set and kept for unknown data testing. This was to test that the fuzzy GA could detect unknown attacks.
+		\item Attacks pulled out of training set and kept for unknown data testing, to test that fuzzy GA could detect unknown attacks.
 		\item Used fuzzy GA and a decision tree algorithm, which is another common algorithm for classification problems.
-		\item 7 tests were run, and for each test case there were 13 attack types plus normal activity that were in the training data set. Three attack types were used for the unknown testing data set.
+		\item 7 tests were run.
+
+		For each test case there were 13 attack types plus normal activity that were in the training data set.
+
+		3 attack types were used for the unknown testing data set.
 	\end{itemize}
@@
 \end{table}
-
 \end{frame}
 
+\subsection{Three Experiments using Both RLD09 and KDD99}
 \begin{frame}
@@
 
-Experiment 1
-	\begin{itemize}
-		\item Used fuzzy GA to classify normal activity and attacks from KDD99 and RLD09.
-	\end{itemize}
+Experiment 1 - Used fuzzy GA to classify normal activity and attacks from KDD99 and RLD09.
 
@@
 		\item Used the fuzzy GA to classify types of attacks in KDD99.
-		\item 158,597 records of DoS attacks. 1,500 records of probe attacks.
 		\item 10 tests were run in total, 5 are shown here.
@@
 \begin{itemize}
-	\item 8 test cases had DR greater than 93\%. Only 2 cases had low DR.
+	\item 8 test cases had DR greater than 93\%. Only 2 cases had low DR, (cases 1 and 2).
 \end{itemize}
@@
 		\item Used the fuzzy GA to classify types of attacks in RLD09.
-		\item 6,400 records of DoS attacks. 10,400 records of probe attacks.
 		\item 17 tests were run in total, 6 are shown here.
@@
 	\begin{center}
-	{\huge Questions?}
+	{\Large Questions?}
 	\end{center}
+\end{frame}
+
+\begin{frame}
+	\frametitle{Probe Attacks}
+	\begin{itemize}
+		\item Advance Port Scan
+		\item Ack Scan
+		\item Xmas Tree
+		\item Host Scan
+		\item UDP Scan
+		\item Syn Scan
+		\item Fin Scan
+		\item RCP Scan
+		\item Portsweep
+		\item Satan
+		\item IP Scan
+	\end{itemize}
+\end{frame}
+
+\begin{frame}
+	\frametitle{DoS Attacks}
+	\begin{itemize}
+		\item UDP Flood
+		\item Jping
+		\item HTTP Flood
+		\item Back
+		\item PoD
+		\item Smurf
+	\end{itemize}
 \end{frame}
@@
 \newblock In \emph{2013 10th International Conference on Electrical Engineering/Electronics, Computer, Telecommunications and Information Technology (ECTI-CON)}, pages 1-6, 2013.
-
-%	\bibitem{DBLP:journals/corr/abs-1204-1336}
-%Mohammad Sazzadul Hoque, Md. Abdul Mukit, Md. Abu Naser Bikas.
-%\newblock An Implementation of Intrusion Detection System Using Genetic Algorithm.
-%\newblock \emph{CoRR}, abs/1204.1336, 2012.
 	\end{small}

### 2013-12-03 Talk/Presentation.tex
Reduced the amount of text on some of the slides.

@@
 		\item Trying to gain root access to the system.
-		\item Delete files of other users and change user passwords.
+		\item Delete files of users and change user passwords.
 	\end{itemize}
 	\linespace
-	\item With an intrusion detection system (IDS) -> classify those attempts.
-	\item Intrusion detection systems provide one way of detecting attacks by monitoring network activities for malicious or abnormal behaviors and then producing reports, alerts, and actions.
+	\item With intrusion detection system (IDS): classify attempts.
+	\item Intrusion detection systems provide way of detecting attacks by monitoring network activities for malicious or abnormal behaviors then producing reports, alerts, actions.
 	\item Training an IDS: use a fuzzy genetic algorithm.
@@
   		\item  Systems lock out user from account after certain number of failed login attempts.
-Attacker would be able to use this to prevent users from logging in, by failing to log in enough times to lock the account.
+
+  		Use this to prevent users from logging in, by failing to log in enough times to lock account.
   	\end{itemize}
@@
 
-  	examines machine in order to collect information about weaknesses that in the future could be used to compromise the system.
+  	examines machine in order to collect info about weaknesses that in future could be used to compromise system.
 
   	\begin{itemize}
-  		\item Attacker could be trying to determine what version of a software is being run on that machine, and if that version has a known issue then that allows them to attempt to attack that.
+  		\item Trying to determine what version of software is being run, if version has known issue it allows attacker to attempt to attack that.
   	\end{itemize}
@@
   	\item Signature-based detection: compares well-known patterns of attacks
-that are already known to the IDS against captured events in order to identify a possible attack.
+that are already known to IDS against captured events in order to identify possible attack.
 	\begin{itemize}
@@
 			\end{itemize}
-		\item Each record is classified as either normal or attack activity.
+		\item Each record: classified as either normal or attack.
 	\end{itemize}
@@
 	\begin{itemize}
-		\item KDD99 uses 41 \emph{features}, which are properties of a record that are used to describe the activity and help to distinguish normal connections from attacks.
+		\item KDD99 uses 41 \emph{features} - properties of record that are used to describe activity, help to distinguish normal connections from attacks.
 
@@
 	\begin{itemize}
-		\item RLD09 was created because KDD99 is 14 years old,
+		\item RLD09: created because KDD99 is 14 years old,
 
-		newer attack types are not in KDD99 because of its age.
-		\item Data was captured from a university in Bangkok, Thailand.
-		\item As well as normal network activity, has 17 different types of attacks (divided into denial of service and probe attacks).
+		newer attack types not in KDD99 because of age.
+		\item Data was captured from university in Bangkok, Thailand.
+		\item Has normal network activity, and 17 different types of attacks (divided into denial of service and probe attacks).
 	\end{itemize}
@@
 		\begin{itemize}
-			\item The condition is composed of one or more features, and the consequence says if it is an intrusion or not.
+			\item Condition: one or more features
+			\item Consequence: says if it is an intrusion or not.
+			\linespace
 			\item If \emph{duration = 4} then \emph{intrusion}.
@@
 	\begin{itemize}
-		\item Fuzzy logic is used in intrusion detection systems to find the degree of certainty of a record being an attack.
-		\item Fuzzy logic rules are similar to the rules described before, except that
-consequence is a certainty factor. 
+		\item Fuzzy logic: used in intrusion detection systems to find degree of certainty of record being attack.
+		\item Fuzzy logic rules: similar to rules described before, except that
+consequence is certainty factor. 
 		\begin{itemize}
@@
 	\begin{itemize}
-	\item The four parameters are encoded into blocks. Each block is a feature with values between 0 and 7.
-	\item A rule has one block for each of 12 features followed at the end by a marker indicating the type of attack.
+	\item Four parameters are encoded into blocks. Each block is feature with values 0-7.
+	\item Rule has 1 block for each of 12 features followed at end by marker indicating type of attack.
 	\end{itemize}
@@
 	\begin{itemize}
-		\item The degree of certainty is computed for each of the 12 blocks, and if the sum of those is greater than a threshold, then it will be declared as an attack.
+		\item Degree of certainty is computed for each of 12 blocks, if sum of those is greater than threshold, declared as attack.
 	\end{itemize}
@@
 		\item GAs: search technique used to find solutions to problems.
-        \item Possible solutions to problems can be represented in a variety of problem dependent ways, such as bit strings.
+        \item Possible solutions to problems: represented in a variety of problem dependent ways.
 
@@
         \linespace
-        \item First, randomly generated population of potential solutions is created.
-Mutation, crossover, selection are applied to each generation until acceptable solution is found or some time limit is exceeded.
+        \item Randomly generated population of potential solutions is created. Mutation, crossover, selection applied to each generation until acceptable solution is found or time limit is exceeded.
 	\end{itemize}
@@
         \item Selection: individuals that have better fitness are chosen to be parents.
-        \item The fitness of an individual is specified by the fitness function, which determines the quality of a particular individual.
+        \item Fitness of individual is specified by fitness function, which determines quality of particular individual.
 
@@
 
-		\item If the record is either a DoS rule or a Probe rule, it is classified as an attack; else it is normal.
+		\item If record is either DoS rule or Probe rule, it is classified as attack; else normal.
 \end{itemize}
@@
 	\begin{itemize}
-		\item Attacks pulled out of training set and kept for unknown data testing, to test that fuzzy GA could detect unknown attacks.
+		\item Attacks pulled out of training set, kept for unknown data testing, to test that fuzzy GA could detect unknown attacks.
 		\item Used fuzzy GA and a decision tree algorithm, which is another common algorithm for classification problems.
@@
 
-		For each test case there were 13 attack types plus normal activity that were in the training data set.
+		Each test case: 13 attack types plus normal activity that were in training set.
 
-		3 attack types were used for the unknown testing data set.
+		3 attack types used for unknown testing data set.
 	\end{itemize}

### 2013-12-04 Talk/Presentation.tex
Changed the trapezoid example, added another slide about trapezoids, added stuff to the training/test set slide, added stuff about Back and PoD.

@@
 % The text in square brackets is the short version of the date if you need that.
-\date[December 2013] % (optional)
-{December 2013 \\ UMM CSci Senior Seminar Conference}
+\date[December 7, 2013] % (optional)
+{December 7, 2013 \\ UMM CSci Senior Seminar Conference}
 
@@
   \begin{itemize}
-  	\item Denial of Service (DoS):
-
-  	makes machine inaccessible to user by making it too busy to serve legitimate requests.
+  	\item Denial of Service (DoS) - makes machine inaccessible to user by making it too busy to serve legitimate requests.
 
   	\begin{itemize}
-  		\item  Systems lock out user from account after certain number of failed login attempts.
-
-  		Use this to prevent users from logging in, by failing to log in enough times to lock account.
+  		\item Systems lockout user from account after failed login attempts. 
+  		\item Use this to prevent users from logging in, by failing to log in enough times to lock account.
   	\end{itemize}
@@
   	\linespace
+  	\linespace
 
-  	\item Probe:
-
-  	examines machine in order to collect info about weaknesses that in future could be used to compromise system.
+  	\item Probe - examines machine to collect info about weaknesses, could be used to compromise system.
 
   	\begin{itemize}
-  		\item Trying to determine what version of software is being run, if version has known issue it allows attacker to attempt to attack that.
+  		\item Trying to determine version of software is being run, if it has known issue it allows attacker to attempt to attack that.
   	\end{itemize}
@@
 
+\linespace
 \linespace
@@
 	\begin{itemize}
-		\item The authors divided the data set into two subsets, a \emph{training set} and a \emph{test set}.
-        \item The given algorithm is then trained on the training set to look for patterns.
+		\item Divide data set into 2 subsets: \emph{training set} and \emph{test set}.
+        \item The given algorithm is trained on the training set to look for patterns.
         \item These patterns are then verified using the test set.
+        \linespace
+        \linespace
+		\item Algorithm used runs risk of memorizing the data in training set, so important to keep some data separate, as unseen data for testing.        
 	\end{itemize}
@@
 		\item Fuzzy logic: used in intrusion detection systems to find degree of certainty of record being attack.
+			\begin{itemize}
+				\item If it's not clear if activity is an attack, fuzzy logic says where on a spectrum it is, how certain it is of being an attack.
+			\end{itemize}
+		\linespace
+		\linespace
+		\linespace
 		\item Fuzzy logic rules: similar to rules described before, except that
@@
 		\begin{itemize}
-			\item If (\emph{duration} = 6) then (\emph{the degree of certainty of the record being an attack is 0.5}).
+			\item If (\emph{duration} = 5.4) then (\emph{the degree of certainty of the record being an attack is 0.8}).
 		\end{itemize}
@@
 \begin{frame}
+  \frametitle{Trapezoidal Shape}
+	\begin{itemize}
+		\item Used to decide how certain a record is of being an attack.
+		\item Described with 4 numbers, used to determine what the trapezoid looks like.
+		\begin{center}
+		  \includegraphics[width=0.85\textwidth]{../trapFigTemplate.pdf}		
+		\end{center}
+		\item Certain in middle, not as certain in triangle areas.
+	\end{itemize}
+\end{frame}
+
+\begin{frame}
   \frametitle{Finding the Degree of Certainty of a Record Being an Attack}
-	Suppose that the feature is duration, and it is 6 seconds. Then data=6.
+	Suppose that the feature is duration, and it is 5.4 seconds. Then data=5.4.
   \includegraphics[width=0.95\textwidth]{../trapFigExample.pdf}
@@
 \end{table}
+
+\begin{center}
+Testing has higher DR; it's more likely to classify attacks because it can match either attack type.
+\end{center}
 \end{frame}
@@
 \begin{itemize}
-	\item 15 cases had DR greater than 97\%. 2 cases had low DR, (cases 3 and 5).
+	\item 15 cases had DR greater than 97\%. 2 cases had lower DR, (cases 3 and 5).
 \end{itemize}
@@
 \begin{frame}
+	\frametitle{DoS Attacks}
+	\begin{itemize}
+		\item Back: attack against the Apache web server.
+Attacker submits requests with URL's containing many front slashes.
+As server tries to process these requests it will slow down, becomes unable to process other requests.
+		\linespace
+		\linespace
+		\item PoD: involves sending a malformed/malicious ping to computer.
+Historically, many systems could not handle a ping packet larger than the maximum IPv4 packet size (65,535 bytes).
+Sending a ping of this size could crash the target computer.
+	\end{itemize}
+\end{frame}
+
+\begin{frame}
+	\frametitle{DoS Attacks}
+	\begin{itemize}
+		\item UDP Flood
+		\item Jping
+		\item HTTP Flood
+		\item Smurf
+	\end{itemize}
+\end{frame}
+
+\begin{frame}
 	\frametitle{Probe Attacks}
@@
 		\item IP Scan
-	\end{itemize}
-\end{frame}
-
-\begin{frame}
-	\frametitle{DoS Attacks}
-	\begin{itemize}
-		\item UDP Flood
-		\item Jping
-		\item HTTP Flood
-		\item Back
-		\item PoD
-		\item Smurf
 	\end{itemize}

### 2013-12-06 Talk/Presentation.tex
Added more to the mutation and crossover slides, also made a few small changes.

@@
 	\begin{itemize}
-		\item Trying to gain root access to the system.
-		\item Delete files of users and change user passwords.
+		\item Trying to gain root access to the system
+		%\item Delete files of users and change user passwords.
 	\end{itemize}
@@
 	\item With intrusion detection system (IDS): classify attempts.
-	\item Intrusion detection systems provide way of detecting attacks by monitoring network activities for malicious or abnormal behaviors then producing reports, alerts, actions.
+	\item Intrusion detection systems provide a way of detecting attacks by monitoring network activities for malicious or abnormal behaviors then producing reports, alerts, actions.
 	\item Training an IDS: use a fuzzy genetic algorithm.
@@
   	\item Signature-based detection: compares well-known patterns of attacks
-that are already known to IDS against captured events in order to identify possible attack.
+that are already known to IDS against captured events in order to identify possible attacks.
 	\begin{itemize}
@@
 		\item Data was captured from university in Bangkok, Thailand.
-		\item Has normal network activity, and 17 different types of attacks (divided into denial of service and probe attacks).
+		\item Has normal network activity, and 17 different types of attacks
 	\end{itemize}
@@
 	\begin{itemize}
+		\item Algorithm used runs risk of memorizing the data in training set, so important to keep some data separate, as unseen data for testing.
+        \linespace
+        \linespace
 		\item Divide data set into 2 subsets: \emph{training set} and \emph{test set}.
@@
         \item These patterns are then verified using the test set.
-        \linespace
-        \linespace
-		\item Algorithm used runs risk of memorizing the data in training set, so important to keep some data separate, as unseen data for testing.        
 	\end{itemize}
@@
   \frametitle{Determining the Accuracy of an Algorithm}
-%\centering{Predicted}
 \begin{table}
-%Actual
 Predicted
@@
 			\begin{itemize}
-				\item If it's not clear if activity is an attack, fuzzy logic says where on a spectrum it is, how certain it is of being an attack.
+				\item If it's not clear if activity is an attack, fuzzy logic says where on a spectrum it is and how certain it is of being an attack.
 			\end{itemize}
@@
 		\begin{itemize}
-			\item If (\emph{duration} = 5.4) then (\emph{the degree of certainty of the record being an attack is 0.8}).
+			\item If (\emph{duration} = 6.2) then (\emph{the degree of certainty of the record being an attack is 0.8}).
 		\end{itemize}
@@
 		\item Used to decide how certain a record is of being an attack.
-		\item Described with 4 numbers, used to determine what the trapezoid looks like.
+		\item Described with 4 numbers that are used to determine what the trapezoid looks like.
 		\begin{center}
@@
   \frametitle{Finding the Degree of Certainty of a Record Being an Attack}
-	Suppose that the feature is duration, and it is 5.4 seconds. Then data=5.4.
+	Suppose that the feature is duration, and it is 6.2 seconds. Then data=6.2.
   \includegraphics[width=0.95\textwidth]{../trapFigExample.pdf}
@@
         \begin{itemize}
-        	\item IDS rules can be represented as bit strings.
+        	\item IDS rules are represented as bit strings.
         \end{itemize}
@@
         \linespace
-        \item Randomly generated population of potential solutions is created. Mutation, crossover, selection applied to each generation until acceptable solution is found or time limit is exceeded.
+        \item A randomly generated population of potential solutions is created. Mutation, crossover, selection applied to each generation until acceptable solution is found or time limit is exceeded.
 	\end{itemize}
@@
 \begin{frame}
-  \frametitle{Mutation and Crossover}
-	\begin{itemize}
-        \item Mutation: random bits in an individual, or possible solution, are randomly changed.
-        \begin{itemize}
-        	\item Takes bits of rule and changes them to form slightly different rule.
-
+  \frametitle{Mutation}
+Mutation: random bits in an individual, or possible solution, are randomly changed. Mutation takes bits of rule and changes them to form slightly different rule.  
 \begin{figure}
+\begin{small}
 	\begin{tabular}{|cccc|} \hline
@@
 	\hline\end{tabular}
+\end{small}
 \end{figure}
+	\includegraphics[width=0.50\textwidth]{../Talk/MutCrossTrapezoids/mut2345.pdf}
+	\includegraphics[width=0.55\textwidth]{../Talk/MutCrossTrapezoids/mut2347.pdf}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Crossover}
+  \begin{itemize}
+  	\item Crossover: two individuals swap sequences of bits to form two new individuals.
+  	\item Crossover takes 2 rules and creates new rules by swapping bits of old rules.
   \end{itemize}
+\end{frame}
 
-        \linespace
-        \linespace
-        \linespace
-
-        \item Crossover: two individuals swap sequences of bits to form two new individuals.
-        \begin{itemize}
-        	\item Take 2 rules and creates new rules by swapping bits of old rules.
-
+\begin{frame}
+  \frametitle{Crossover}
 \begin{figure}
+\begin{small}
+\textbf{Before}
 	\begin{tabular}{|cccc|} \hline
-			001 & 011 & 101 & 111\\
-			a=1 & b=3 & c=5 & d=7\\
+	001 & \textbf{011} & 101 & 111\\
+	a=1 & \textbf{b=3} & c=5 & d=7\\
 	\hline\end{tabular}
@@
 	\begin{tabular}{|cccc|} \hline
-			010 & 100 & 110 & 111\\
-			a=2 & b=4 & c=6 & d=7\\
+	010 & \textbf{100} & 110 & 111\\
+	a=2 & \textbf{b=4} & c=6 & d=7\\
 	\hline\end{tabular}
+\end{small}
 \end{figure}
-        \end{itemize}
-	\end{itemize}
+	\includegraphics[width=0.50\textwidth]{../Talk/MutCrossTrapezoids/cross1357.pdf}
+	\includegraphics[width=0.55\textwidth]{../Talk/MutCrossTrapezoids/cross2467.pdf}
+
+\begin{figure}
+\begin{small}
+\textbf{After}
+	\begin{tabular}{|cccc|} \hline
+	001 & \textbf{100} & 101 & 111\\
+	a=1 & \textbf{b=4} & c=5 & d=7\\
+	\hline\end{tabular}
+	\quad
+	\begin{tabular}{|cccc|} \hline
+	010 & \textbf{011} & 110 & 111\\
+	a=2 & \textbf{b=3} & c=6 & d=7\\
+	\hline\end{tabular}
+\end{small}
+\end{figure}
+	\includegraphics[width=0.50\textwidth]{../Talk/MutCrossTrapezoids/cross1457.pdf}
+	\includegraphics[width=0.55\textwidth]{../Talk/MutCrossTrapezoids/cross2367.pdf}
 \end{frame}
@@
 
-		\item If record is either DoS rule or Probe rule, it is classified as attack; else normal.
+		\item If record is classified as either DoS rule or Probe rule, it is classified as attack; else normal.
 \end{itemize}
@@
 
-\begin{center}
-Testing has higher DR; it's more likely to classify attacks because it can match either attack type.
-\end{center}
+\linespace
+Testing has:
+\begin{itemize}
+	\item Higher DR: IDS is more likely to classify attacks because it can match either attack type.
+	\item Lower FN: IDS is less likely to predict that it's not an attack when it really is.
+\end{itemize}
 \end{frame}
@@
 		\item Attacks pulled out of training set, kept for unknown data testing, to test that fuzzy GA could detect unknown attacks.
-		\item Used fuzzy GA and a decision tree algorithm, which is another common algorithm for classification problems.
-		\item 7 tests were run.
-
-		Each test case: 13 attack types plus normal activity that were in training set.
+		\item Fuzzy GA and decision tree algorithm, which is another common algorithm for classification problems.
+		\item 7 tests were run. Each test case: 13 attack types plus normal activity that were in training set.
 
 		3 attack types used for unknown testing data set.
+		\item Anomaly-based detection
 	\end{itemize}
@@
 		\item Fuzzy genetic algorithms are good at detecting unknown attacks.
-		\item The use of fuzzy genetic algorithms in intrusion detection is an effective way of detecting attacks.
+		\item The use of fuzzy genetic algorithms in intrusion detection is an effective way of detecting attacks - DR in experiments was in the high 90s.
 	\end{itemize}
@@
 	\end{center}
-\end{frame}
-
-\begin{frame}
-	\frametitle{DoS Attacks}
-	\begin{itemize}
-		\item Back: attack against the Apache web server.
-Attacker submits requests with URL's containing many front slashes.
-As server tries to process these requests it will slow down, becomes unable to process other requests.
-		\linespace
-		\linespace
-		\item PoD: involves sending a malformed/malicious ping to computer.
-Historically, many systems could not handle a ping packet larger than the maximum IPv4 packet size (65,535 bytes).
-Sending a ping of this size could crash the target computer.
-	\end{itemize}
-\end{frame}
-
-\begin{frame}
-	\frametitle{DoS Attacks}
-	\begin{itemize}
-		\item UDP Flood
-		\item Jping
-		\item HTTP Flood
-		\item Smurf
-	\end{itemize}
-\end{frame}
-
-\begin{frame}
-	\frametitle{Probe Attacks}
-	\begin{itemize}
-		\item Advance Port Scan
-		\item Ack Scan
-		\item Xmas Tree
-		\item Host Scan
-		\item UDP Scan
-		\item Syn Scan
-		\item Fin Scan
-		\item RCP Scan
-		\item Portsweep
-		\item Satan
-		\item IP Scan
-	\end{itemize}
 \end{frame}
@@
 
+\begin{frame}
+	\frametitle{DoS Attacks}
+	\begin{itemize}
+		\item Back: Attacker submits requests with URL's containing many front slashes. As server tries to process these requests it will slow down, becomes unable to process other requests.
+		\linespace
+		\linespace
+		\item PoD: involves sending a malformed/malicious ping to computer.
+Historically, many systems could not handle a ping packet larger than the maximum IPv4 packet size (65,535 bytes).
+Sending a ping of this size could crash the target computer.
+	\end{itemize}
+\end{frame}
+
+\begin{frame}
+	\frametitle{Features of KDD99}
+	\begin{itemize}
+	\begin{small}
+  \item src\_bytes: number of bytes sent from source to destination. Source is user who may or may not be attacker, destination is server being potentially attacked.
+
+  \linespace
+  \linespace
+
+  \item serror\_rate: percentage of connections that have ``SYN" errors.
+When client attempts to connect to server, it first sends a SYN (synchronize) message to server.
+The server then acknowledges the request by sending a SYN-ACK to client.
+The connection is established when client sends an ACK back to server.
+A SYN error is a failure to get an ACK back.
+  \end{small}
+	\end{itemize}
+\end{frame}
+
 \end{document}

### 2013-12-06 Talk/Presentation.tex
Added another slide on encoding of features and rules, added to the mutation slide, added stuff to the fitness function slide, added more to the conclusions side.

@@
 \usepackage{algorithmic}
+
+\newcommand{\tab}[1]{\hspace{.2\textwidth}\rlap{#1}}
 
@@
 	\begin{itemize}
-		\item RLD09: created because KDD99 is 14 years old,
+		\item RLD09: created because KDD99 was 10 years old,
 
@@
 			\linespace
-			\item If \emph{duration = 4} then \emph{intrusion}.
+			\item If \emph{duration $\leq$ 4} then \emph{intrusion}.
 		\end{itemize}				
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section[Fuzzy Algorithm]{Fuzzy Algorithm}
+\section[Fuzzy Classification]{Fuzzy Classification}
 \subsection{Fuzzy Logic}
@@
 \begin{tabular}{|cccc|c|cccc|c|} \hline
-010 & 011 & 100 & 101   & ... & 010 & 011 & 101 & 111   & Attack\\
-a=2 & b=3 & c=4 & d=5   & ... & a=2 & b=3 & c=5 & d=7   &\\ 
+010 & 011 & 100 & 101   & ... & 001 & 100 & 101 & 111   & Attack\\
+a=2 & b=3 & c=4 & d=5   & ... & a=1 & b=4 & c=5 & d=7   &\\  
     &     & Block 1&    &        &     & Block 12& &       & Type\\
@@
 \end{figure}
+\end{frame}
 
-	\begin{itemize}
-		\item Degree of certainty is computed for each of 12 blocks, if sum of those is greater than threshold, declared as attack.
-	\end{itemize}
+\begin{frame}
+	\frametitle{Encoding of Features and Rules}
+\begin{figure}
+\begin{small}
+\begin{tabular}{|cccc|c|cccc|c|} \hline
+010 & 011 & 100 & 101   & ... & 001 & 100 & 101 & 111   & Attack\\
+a=2 & b=3 & c=4 & d=5   & ... & a=1 & b=4 & c=5 & d=7   &\\ 
+    &     & Block 1&    &        &     & Block 12& &       & Type\\
+\hline\end{tabular}
+\end{small}
+\end{figure}
+\tab{$\downarrow$} \tab{} \tab{} \tab{$\downarrow$}
+  \includegraphics[width=0.50\textwidth]{../Talk/MutCrossTrapezoids/mut2345.pdf}
+  \includegraphics[width=0.58\textwidth]{../Talk/MutCrossTrapezoids/cross1457.pdf}
 
+Degree of certainty is computed for each of 12 blocks, if sum of those is greater than a threshold, declared as attack.
 \end{frame}
@@
 Mutation: random bits in an individual, or possible solution, are randomly changed. Mutation takes bits of rule and changes them to form slightly different rule.
+
 \begin{figure}
@@
 \end{figure}
+
 	\includegraphics[width=0.50\textwidth]{../Talk/MutCrossTrapezoids/mut2345.pdf}
 	\includegraphics[width=0.55\textwidth]{../Talk/MutCrossTrapezoids/mut2347.pdf}
+
+After mutation: transition from c to d is more gradual.
 \end{frame}
@@
   	\item Crossover: two individuals swap sequences of bits to form two new individuals.
-  	\item Crossover takes 2 rules and creates new rules by swapping bits of old rules.
+  	\item In IDS: crossover takes 2 rules and creates new rules by swapping bits of old rules.
   \end{itemize}
@@
 	\frametitle{Fitness function}
-	The fitness function in the algorithm is:
+	The fitness function to be maximized is:
 	\begin{equation*}
@@
 	$B$: \# of normal records.
+
+\linespace
+\linespace
+\linespace
+\linespace
+Best possible value of $\beta$ is 0.
+It's good if $\alpha$ = $A$.
+
+Best possible fitness value is 1.
 \end{frame}
@@
 
-		Both rules were then used together in testing process to identify attacks from testing data set.
+		\item Both rules were then used together in testing process to identify attacks from testing data set.
+
+		\linespace
+		\linespace
+
+		\item If record is classified as either DoS rule or Probe rule, it is classified as attack; else normal.
 
@@
 		Test set: 26,500 records.
-
-		\linespace
-		\linespace
-
-		\item If record is classified as either DoS rule or Probe rule, it is classified as attack; else normal.
 \end{itemize}
@@
 		\item Fuzzy genetic algorithms are good at detecting unknown attacks.
-		\item The use of fuzzy genetic algorithms in intrusion detection is an effective way of detecting attacks - DR in experiments was in the high 90s.
+		\begin{itemize}
+			\item Fuzzy GA DR: 99.8\%, Decision Tree DR: 46.7\%
+		\end{itemize}
+
+		\item The use of fuzzy genetic algorithms in intrusion detection is an effective way of detecting attacks - DR in all experiments was in the high 90s.
 	\end{itemize}

### 2013-12-06 Talk/Presentation.tex
Added more to the big picture slide.

@@
 		\item Trying to gain root access to the system
-		%\item Delete files of users and change user passwords.
+		\item Delete files of users and change user passwords.
 	\end{itemize}
 	\linespace
+	\linespace
+	\linespace
+
+	\item An attempt is an attack if there are more than \emph{n} number of attempts within \emph{t} time interval.
+	\item Number of login attempts (for failed passwords or for users that don't exist) over period of 6 days (12/1 - 12/6):
+	\begin{itemize}
+		\item normal lab box (avenger): 1,834 attacks
+		\item normal box, more recently added than avenger (kenshiro): 1,887 attacks
+	\end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{The Big Picture}
+  \begin{itemize}
 	\item With intrusion detection system (IDS): classify attempts.

### 2013-12-07 Talk/Presentation.tex
Made a few changes.

@@
   	\begin{itemize}
-  		\item Trying to determine version of software is being run, if it has known issue it allows attacker to attempt to attack that.
+  		\item Trying to determine what version of software is being run, if that version has known issue, it allows attacker to attempt to attack that.
   	\end{itemize}
@@
 	\begin{itemize}
-		\item Fuzzy logic: used in intrusion detection systems to find degree of certainty of record being attack.
+		\item Fuzzy logic: used in intrusion detection systems to find degree of certainty of a record being attack.
 			\begin{itemize}
-				\item If it's not clear if activity is an attack, fuzzy logic says where on a spectrum it is and how certain it is of being an attack.
+				\item If it's not clear if the activity is an attack, fuzzy logic says where on a spectrum it is and how certain it is of being an attack.
 			\end{itemize}
@@
 	\begin{itemize}
-	\item Four parameters are encoded into blocks. Each block is feature with values 0-7.
-	\item Rule has 1 block for each of 12 features followed at end by marker indicating type of attack.
+	\item Four parameters are encoded into blocks. Each block is a feature with values 0-7.
+	\item A rule has 1 block for each of 12 features followed at end by a marker indicating type of attack.
 	\end{itemize}
@@
 
-Degree of certainty is computed for each of 12 blocks, if sum of those is greater than a threshold, declared as attack.
+Degree of certainty is computed for each of the 12 blocks, if sum of those is greater than a threshold, declared as an attack.
 \end{frame}
@@
         \linespace
-        \item A randomly generated population of potential solutions is created. Mutation, crossover, selection applied to each generation until acceptable solution is found or time limit is exceeded.
+        \item A randomly generated population of potential solutions is created.
+Mutation, crossover, selection are applied to each generation until acceptable solution is found or time limit is exceeded.
 	\end{itemize}
@@
 \begin{itemize}
+	\item Both DoS and Probe. If it sees either type of attack, it just classifies it as an attack.
 	\item Higher DR: IDS is more likely to classify attacks because it can match either attack type.
